---
title: "combed4nDIABLO"
output: html_document
date: "2025-09-03"
---

## load packages

```{r }

library(CCA)
library(ggplot2)
library(ggtext)
library(ggrepel)
library(vegan)
#BiocManager::install("sRDA")
library(sRDA)
library(stringr)
library(tibble)
#devtools::install_github("mixOmicsTeam/mixOmics")
library(mixOmics)
library(ggraph)
library(dplyr)
library(data.table)
library(tidyr)
library(stringr)
library(tibble)
#BiocManager::install("biomaRt")
library(biomaRt)
library(readr)
library(purrr)
library(igraph)
#install.packages("ggExtra")
library(ggExtra)


# (to get multi-block correlation- and dim reduction plots, a loadings plot, and 2 network plots)

```

## curating/pre-processing

```{r, data preparation }

set.seed(123)
### See 3nMicrobeCecumSerum.Rmd for blocks of code used to generate these objects
# diff abundant metabolites and microbes
#HsMmSb_MouseRNAseq.Rmd blocks 1-2 is where this file was first created
result.metabomat_microbe <- data.frame(data.table::fread("clippedMicro.csv", sep = ","))
rownames(result.metabomat_microbe) <- result.metabomat_microbe$V1
result.metabomat_microbe$V1 <- NULL
result.metabomat_microbe$samples <- NULL
result.metabomat_microbe <- t(result.metabomat_microbe)
###### FILTERRRRRRRR
# result.metabomat_microbe <- result.metabomat_microbe[, grep("Bactero|Lachnospi|Oscillo|Sphingo|Clostrid",
#                                                             colnames(result.metabomat_microbe),
#                                                             value = TRUE)]
### we can go back and quickly re-run the corncob models in cecal_shotgun.Rmd, then subset data.clr columns by the mutual taxa from pw compare corncob models against macaque-donor
# colFilter <- c(fdr_corncob_humac,fdr_corncob_sqmac)
# #colFilter <- intersect(fdr_corncob_humac,fdr_corncob_sqmac)
# colFilter <- gsub("\\|", ".", colFilter)
# filtered_mat <- result.metabomat_microbe[rownames(result.metabomat_microbe) %in% colFilter, ]
# result.metabomat_microbe <- filtered_mat
# ### OR we can try this microbiome mixMC pre-processing
# data.offset <- result.metabomat_microbe + 1
# sum(which(data.offset == 0))
# #data.filter <- t(result.filter$data.filter)
# data.filter <- t(data.offset)
# # #lib size not an issue for our samples; so we can move to PCA
# # ###### double check that low count removal/pseudocount is sufficient to get diablo to take
# pca.result <- pca(data.filter, logratio = 'CLR')
# sampd <- data.frame(sample = rownames(pca.result$X),
#             group = substr(rownames(pca.result$X), start = 1, stop = 3))
# # plotIndiv(pca.result,  # plot samples
# #           group = sampd$group,
# #           title = '',
# #           legend = TRUE)
# data.clr <- logratio.transfo(data.filter,
#                              logratio = 'CLR', offset = 0)
# plotIndiv(pca(data.clr),  # plot samples
#           group = sampd$group,
#           title = 'group-wise microbiome sample dist w/CLR',
#           legend = TRUE)
# 
# # ### may need to perform another filter on this if spls stdev remains zero and downstream work reveals no enrichment for taxa or for taxa that make little sense in bio-context
# result.metabomat_microbe <- data.clr

result.metabomat_liver <- data.frame(data.table::fread("clippedGene.csv", sep = ","))
rownames(result.metabomat_liver) <- result.metabomat_liver$V1
result.metabomat_liver$V1 <- NULL
result.metabomat_liver$samples <- NULL
colnames(result.metabomat_liver) <- c("sqm10","sqm12","hum22","mac4","mac3","hum24","sqm16","mac1",
  "sqm13","hum19","sqm17","mac5","sqm15","hum23","hum20",
  "hum21","mac6","mac8","hum25","sqm18","sqm9","mac2",
  "mac7","hum26","sqm14","sqm11")
#MetabolomicsFinalReport.Rmd block 1 is where this file was first created
## commenting out the total.score filtering after reading in metabolomic data given diablo is responsible for feature selection
## as well as interdataset corr in this analysis and slice_max which picks an annotated compound if it has the same variable_id by which has the highest total score...
## which is not a bad idea for feature selection to weed out unlikely annotations but we want to let diablo handle this 
result.metabomat_cecum <- readr::read_csv("clippedCecum.csv") #%>% filter(Total.score > 0.99)
result.metabomat_cecum <- result.metabomat_cecum %>%
  group_by(variable_id) %>%
  filter(Total.score > 0.9 & mz.match.score > 0.99) %>% 
  #slice_max(Total.score, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  column_to_rownames("Compound.name")
result.metabomat_cecum <- result.metabomat_cecum[, grep("Ben.Oli", colnames(result.metabomat_cecum))]
colnames(result.metabomat_cecum) <- c("mac1","mac2","mac3","mac4","mac5","mac6","mac7","mac8",
                                      "sqm9", "sqm10","sqm11", "sqm12","sqm13", "sqm14","sqm15", "sqm16","sqm17", "sqm18",
                                      "hum19","hum20","hum21","hum22","hum23","hum24","hum25","hum26")
# result.metabomat_cecum <- result.metabomat_cecum[grep("cholic|cholest|cholate|cholan|phospho|bili",
#                                                             rownames(result.metabomat_cecum),
#                                                             value = TRUE), ]
#MetabolomicsFinalReport_serum.Rmd block 1 is where this file was first created 
result.metabomat_serum <- readr::read_csv("clippedSerum.csv") #%>% filter(Total.score > 0.99)
result.metabomat_serum <- result.metabomat_serum %>% 
  group_by(variable_id) %>%
  filter(Total.score > 0.9 & mz.match.score > 0.99) %>% 
  #slice_max(Total.score, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  column_to_rownames("Compound.name")
result.metabomat_serum <- result.metabomat_serum[, grep("Ben.Ant", colnames(result.metabomat_serum))]
colnames(result.metabomat_serum) <- c("mac1","mac2","mac3","mac4","mac5","mac6","mac7","mac8",
                                     "sqm10","sqm11", "sqm12","sqm13", "sqm16","sqm17", "sqm18",
                                      "hum19","hum20","hum21","hum22","hum23","hum24")
## FIX: ENSEMBL GENE ID into GENE NAMES
mart <- useDataset("mmusculus_gene_ensembl", useMart("ensembl"))
opts <- listAttributes(mart)
bridge <- getBM(filters = "ensembl_gene_id",
               attributes = c("ensembl_gene_id","external_gene_name"),
               values = rownames(result.metabomat_liver),
               mart = mart)
result.metabomat_liver$ensembl_gene_id <- rownames(result.metabomat_liver)
result.metabomat_liver <- merge(bridge, result.metabomat_liver, by = "ensembl_gene_id")
result.metabomat_liver$ensembl_gene_id <- NULL
#remove rows that have an empty string in slot external_gene_name
result.metabomat_liver <- result.metabomat_liver[result.metabomat_liver$external_gene_name != "", ]
rownames(result.metabomat_liver) <- result.metabomat_liver$external_gene_name
result.metabomat_liver$external_gene_name <- NULL
# result.metabomat_liver <- result.metabomat_liver[grep("Cyp|Fgf|lotho|Shp|Hnf|Lrh|Nrf|Abcg|Slc10",
#                                                             rownames(result.metabomat_liver),
#                                                             value = TRUE), ]

### get liver RNAseq sig hits; Going to try without making subsets or doing much filtering first
### this may help reduce any unforseen overfitting problems or biasing input feature sets
### if things look weird on this go around without filtering we can come back and toy with some filtering maybe
# livermat_1 <- dds@assays@data@listData[["counts"]]
# #livermatsig_1 <- data.frame(subset(livermat_1, rownames(livermat_1) %in% rownames(res.05)))
# # to get only DE counts
# livermatsig_1 <- data.frame(subset(livermat_1, rownames(livermat_1) %in% rownames(resSig)))

#X1 <- t(result.metabomat_microbe)
X1 <- result.metabomat_microbe
# run the below ONLY ONCE BUT DO MAKE SURE YOU RUN IT
colnames(X1) <- sub(".*s__", "", colnames(X1))
X2 <- t(result.metabomat_cecum)
X3 <- t(result.metabomat_serum)
X4 <- t(result.metabomat_liver)

# order so rows in dataframes match one another
X1 <- X1[match(rownames(X3), rownames(X1)), ] %>% na.omit()
X2 <- X2[match(rownames(X1), rownames(X2)), ] %>% na.omit()
X3 <- X3[match(rownames(X1), rownames(X3)), ] %>% na.omit()
X4 <- X4[match(rownames(X3), rownames(X4)), ]

Y <- as.factor(gsub("[[:digit:]]+", "", rownames(X1)))

# X1 <- scale(log1p(X1 / rowSums(X1)))
# X2 <- scale(log1p(X2))
# X3 <- scale(log1p(X3))
# X4 <- scale(log1p(X4))

data_blocks <- list(microbe = X1,
                 cecum = X2,
                 serum = X3,
                 liver = X4)

##### pw comparison setup
##### sqm vs. mac
# data.smm <- lapply(data, function(m) {
#   m[Reduce(intersect,
#            lapply(data, function(mm) rownames(mm[grep("sqm|mac", rownames(mm)), , drop = FALSE]))
#            ),,drop = FALSE]
#   }
# )
# #### hum vs. mac
# data.hsm <- lapply(data, function(m) {
#   m[Reduce(intersect,
#            lapply(data, function(mm) rownames(mm[grep("hum|mac", rownames(mm)), , drop = FALSE]))
#            ),,drop = FALSE]
#   }
# )
#data_list <- list(data.smm,data.hsm)
#names(data_list) <- c("data.smm", "data.hsm")

#lapply(data_list$data.smm, dim)

data_list <- list(data_blocks)
names(data_list) <- c("data")
lapply(data_list$data, dim)


```


## model preparation for integration

```{r, design model}

# library(BiocParallel)
# 
# for (nm in names(data_list)) {
#   
#   data <- data_list[[nm]]
#   
#   ## initialize design matrix
#   design <- matrix(0.1, ncol = length(data), nrow = length(data), 
#                    dimnames = list(names(data), names(data)))
#   diag(design) <- 0
#   
#   ### intersect sample/rownames across blocks
#   tograb <- Reduce(intersect, list(rownames(data$microbe),
#                                    rownames(data$liver),
#                                    rownames(data$serum), 
#                                    rownames(data$cecum)))
#   data$serum   <- data$serum[tograb, ]
#   data$microbe <- data$microbe[tograb, ]
#   data$liver   <- data$liver[tograb, ]
#   data$cecum   <- data$cecum[tograb, ]
#   
#   ### aware of all three groups at one
#   Y <- as.factor(gsub("[[:digit:]]+", "", rownames(data$serum)))
#   ### aware of macaque-donor distinctness
#   #Y <- factor(ifelse(gsub("[[:digit:]]+", "", rownames(data$serum)) == "mac", "fat", "normal_wt"))
#   
#   # initialize first tester DIABLO model
#   basic.diablo.model <- block.splsda(X = data, Y = Y, ncomp = 2, design = design)
#   
#   # model fit evaluation
#   perf.diablo <- perf(basic.diablo.model, validation = 'loo', 
#                       folds = 10, nrepeat = 5, progressBar = TRUE,
#                       BPPARAM = MulticoreParam(workers = 4))
#   
#   ## select the optimal component
#   ncomp <- perf.diablo$choice.ncomp$WeightedVote["Overall.BER", "centroids.dist"]
#   #liver = c(5:9, seq(10, 18, 2), seq(20,30,5)))
#   # test grid for tuning, the following grid had lowest corr at 0.91 (cecum-serum), every other inter-dataset corr was HIGHER than 0.91 !!
#   #             list(microbe = 218, 
#   #                  cecum = 535,
#   #                  serum = 387,
#   #                  liver = 723)
#   # test grid for tuning, the following grid had lowest corr at 0.92 (cecum-serum), every other inter-dataset corr was HIGHER than 0.92 !!
#               # list(microbe = 20, 
#               #      cecum = 50,
#               #      serum = 50,
#               #      liver = 70)
#   test.keepX <- list(
#     microbe = c(10, 15, 20),
#     cecum   = c(10, 15, 20),
#     serum   = c(5, 10, 15),
#     liver   = c(5, 7, 10)
#   )
#   # tune model
#   tuned_model <- tune.block.splsda(
#     X = data, Y = Y,
#     ncomp = ncomp, test.keepX = test.keepX,
#     design = design, validation = 'loo',
#     folds = 10, nrepeat = 5, dist = "centroids.dist"#,
#     #BPPARAM = MulticoreParam(workers = 4)
#   )
#   
#   assign(paste0("tune.TCGA_", nm), tuned_model)
#   assign(paste0("choice.keepX_", nm), tuned_model$choice.keepX)
#   
#   # build final model with optimal keepX
#   final_model <- block.splsda(
#     X = data, Y = Y, ncomp = 2, 
#     keepX = tuned_model$choice.keepX,
#     design = design
#   )
#   
#   assign(paste0("final.diablo.model_", nm), final_model)
# }
# 
# # tune.TCGA = tune.block.splsda(X = data, Y = Y, 
# #                                   ncomp = ncomp, test.keepX = test.keepX, 
# #                                   design = design, validation = 'loo',
# #                                   folds = 5, nrepeat = 1, dist = "centroids.dist"
# #                                   #BPPARAM = BiocParallel::MulticoreParam(workers = parallel::detectCores()-1)
# #                               )
# 
# wv_row <- perf.diablo$choice.ncomp$WeightedVote["Overall.BER", ]
# 
# result <- if(length(unique(wv_row)) == 1) {
#   "centroids.dist"          #centroids.dist if all are tied
# } else {
#   names(wv_row)[which.max(wv_row)]  #otherwise winner dist metric used
# }
# 
# #write_rds(final.diablo.model_data, "recmodel.Rds")

```


## dimenstionality reduction, interdataset corr., and loadings

```{r}

final.diablo.model_data <- readRDS("recmodel.Rds")
#PAPER 1
#c("#1b9e77", "#a6761d", "#666666"))


# Capitalize names

levels(final.diablo.model_data$Y) <- tools::toTitleCase(levels(final.diablo.model_data$Y))

# Define corrected block names
corrected_names <- c("Microbiome", "Cecal Metabolome", "Serum Metabolome", "Liver Transcriptome")

# Update block labels
final.diablo.model_data$names$blocks <- corrected_names

# Update names of colnames list
names(final.diablo.model_data$names$colnames)[1:4] <- corrected_names

# Update names of all top-level slots that use block names

slots_to_update <- c("X", "variates", "loadings", "explained_variance", "ind")

for (slot in slots_to_update) {
  if (!is.null(final.diablo.model_data[[slot]])) {
    current_names <- names(final.diablo.model_data[[slot]])
    n <- length(current_names)
    
    # Only rename if there are at least 4 elements
    if (n >= 4) {
      names(final.diablo.model_data[[slot]])[1:4] <- corrected_names
    }
  }
}

names(final.diablo.model_data$ncomp)[1:4] <- corrected_names

#S5A

jpeg("plotDIAB.jpeg", units="in", height = 10, width= 15, res=300)

#set plotting parameters

par(cex = 14,       # global scaling (affects most text)
    cex.lab = 1,   # axis label size
    font.lab = 2,    # bold axis labels
    cex.axis = 1.3, # tick label size
    ps=21
    ) 



plotDiablo(final.diablo.model_data,
           ncomp = 1,
           col.per.group = c("#1b9e77", "#a6761d", "#666666"),
          size.title = 22,
          size.subtitle = 22,
          size.axis = 1,
          size.legend = 28,
          size.legend.title = 22,
          font= 2
)


dev.off()


#S5B

jpeg("plotINDIV.jpeg", units="in", height = 10, width= 15, res=300)


plot9 <- plotIndiv(final.diablo.model_data,
          ind.names = FALSE,
          legend = TRUE,
          title =,
  subtitle = c("Microbiome", "Cecal Metabolome", "Serum Metabolome", "Liver Transcriptome"),
          col = c("#1B9E77", "#A6761D", "#666666"),
          legend.title = "Species",       #  legend
          legend.font = 2,    # Bold legend
          cex = 4,              # Increase axis and label text
          size.title = 24,
          size.subtitle = 18,
          size.xlabel = 22,
          size.ylabel = 22,
          size.axis = 19,
          size.legend = 22,
          size.legend.title = 22,
          X.label = "Variate 1",
          Y.label = "Variate 2",
          font = 2                # Bold axis labels
)
 
plot(plot9$graph)+ removeGrid()

dev.off()

## forgoing
# plotArrow(final.diablo.model_data, ind.names = FALSE, legend = TRUE, 
#           title = 'DIABLO', col = c("#1b9e77", "#a6761d", "#666666"))
# plotVar(final.diablo.model_data, var.names = FALSE, 
#         style = 'graphics', legend = TRUE,
#         pch = c(16, 17, 15, 18), cex = c(2,2,2,2), 
#         col = c('darkorchid', 'brown1', 'lightgreen', 'blue'))
## PAPER 1

#S5B

#replace _ with "

cleaned_names <- colnames(final.diablo.model_data$X$Microbiome) %>%
  stringr::str_replace_all("_", " ")

colnames(final.diablo.model_data$X$Microbiome) <- cleaned_names
rownames(final.diablo.model_data$loadings$Microbiome) <- cleaned_names


jpeg("loadings4nclip.jpeg", units = "in", height = 15, width = 25, res = 300)

#apply preplotting functions


layout(
  matrix(c(1, 5, 2,
           3, 5, 4), nrow = 2, byrow = TRUE),
  widths = c(1, 0.3, 1)  
)

par(oma = c(0, 42.4, 1, 0))
par(mar = c(5, 65, 4, 2), lwd = 2, font.axis = 2, font.lab = 2, font = 2)

#plot
plotLoadings(
  final.diablo.model_data,
  comp = 1,
  contrib = 'max',
  method = 'median',
  legend.color = c("#1b9e77", "#a6761d", "#666666"),
  subtitle = c("Microbiome", "Cecal Metabolome", "Serum Metabolome", "Liver Transcriptome"),
  title = NULL,
  legend = FALSE,
  cex.subtitle = 3,
  size.name = 2,
  size.title = 10,
  size.subtitle = 4,
  size.xlabel = 4,
  size.ylabel = 4,
  lwd=2,
  size.axis = 2,
  font = 2,
)


par(mar = c(0, 0, 0, 0))
plot(1, type = "n", axes = FALSE, xlab = "", ylab = "")

legend(
  "center",
  legend = levels(final.diablo.model_data$Y),
  fill = c("#1b9e77", "#a6761d", "#666666"),
  title = "Species",
  cex = 2,
  text.font = 2
)

#  # Top-left panel → Microbiome
# par(mfg = c(1, 1))
# usr <- par("usr")
# text(x = mean(usr[1:2]), y = usr[4] - 0.035 * diff(usr[3:4]), "Microbiome", font = 2, cex = 3.4)
# 
# # Top-right panel → Cecal Metabolome
# par(mfg = c(1, 2))
# usr <- par("usr")
# text(x = mean(usr[1:2]), y = usr[4] - 0.035 * diff(usr[3:4]), "Cecal Metabolome", font = 2, cex = 3.4)
# 
# 
# # Liver Metabolome → Panel 4
# par(mfg = c(2, 1))
# usr <- par("usr")
# text(x = mean(usr[1:2]), y = usr[4] - 0.03 * diff(usr[3:4]), "Liver Transcriptome", font = 2, cex = 3.4)
# 
# # Serum Transcriptome → Panel 3
# par(mfg = c(1, 3))
# usr <- par("usr")
# text(x = mean(usr[1:2]), y = usr[4] - 0.03 * diff(usr[3:4]), "Serum Metabolome", font = 2, cex = 3.4)
# 


dev.off()



```


## initial diablo network 

```{r}

########### preliminary network projection that diablo produces from highly correlated features it identifies across blocks
jpeg("network4nclip.jpeg", units="in", width=40, height=30, res=300)
diablo_net <- network(final.diablo.model_data, blocks = c(1,2,3,4),
        color.node = c('darkorchid','brown1', 'lightgreen', 'blue'), cutoff = 0.75,
        size.node = 0.05, graph.scale = 0.1
        )
dev.off()


#### but that's hard to visualize and it's not clear how those ideas fit into proposed models from independent analyses 
############################## hypertrophy (main mechanism)  ###################################
# B. theta’s incomplete carb ferment, causing increase incompletely fermented carb/carboxylic acid intermediates
# These can disrupt cysteine catabolism (Cdo1 suggested in serum; Csad in transcriptome; consistent Fmo1 expr across groups – this is final conv. to taurine)
# Incr. hypotaurine leading to potential spillover
# increasing adipocyte uptake (Slc6a6)/Incr. Slc6a6 associated with obesity
# Lit. says incr. hypotaurine:taurine ratio depletes B-cat for Wnt signaling (altered Wnt5a methylation bc hypotaurine?)
# depleted B-cat reduces Wnt complex/signal forcing stem cells toward adipocyte fate via paths like Ppar (suggestion of both Wnt and Ppar in serum data)

############################## hyperplasia (complementary/supportive)  ###################################
# B. theta’s incomplete carb ferment, causing increase incompletely fermented carb/carboxylic acid intermediates
# These can disrupt cysteine catabolism (Cdo1 suggested in serum; Csad in transcriptome; consistent Fmo1 expr across groups – this is final conv. to taurine)
# Incr. hypotaurine leading to potential spillover
# increasing adipocyte uptake (Slc6a6)/Incr. Slc6a6 associated with obesity
# Lit. says incr. hypotaurine:taurine ratio depletes B-cat for Wnt signaling (altered Wnt5a methylation bc hypotaurine?)
# depleted B-cat reduces Wnt complex/signal forcing stem cells toward adipocyte fate via paths like Ppar (suggestion of both Wnt and Ppar in serum data)


```








## Network 1: keyword subset network
contains terms of interest relevant to both prev. proposed models w/o excl any terms diablo may have ID'd as highly covariable across blocks


```{r}

# # define hypothesis feature lists
keywords <- c(
  "Lachno", "thetaiotaomicron", "scindens", "symbiosum","hiranonis",
  "Csad","Cdo1","Fmo1","Stra","Aldh","Rdh","Scd",
  "Rbp","Fasn","Acaca","Elovl","Retsat","Cyp","Slc10a1","Hsd","Ctnnb1","Wnt","Ppar","Sreb","Abcb11","Sfrp1","Fzd",
  "carb","sulf","chol","taur","oxy","stero","glucos","mann",
  "acetyl","phos","sugar","glyco","sacch","retin","malonyl", "palmit", "steara","acyl", "olea",
  "sialic","fucose","hexose","pentose","heptose","octulose","uronic","lactone"
)

top_features <- lapply(names(final.diablo.model_data$X), function(block) {
  load_block <- final.diablo.model_data$loadings[[block]][, 1]   #loadings comp1
  load_block <- load_block[!is.na(load_block)] #chuck NA
  keep_idx <- grepl(paste(keywords, collapse = "|"), names(load_block), ignore.case = TRUE) #keyword pull
  load_block <- load_block[keep_idx]
  feats <- names(sort(abs(load_block), decreasing = TRUE))[1:min(30, length(load_block))] #get abs vals for loading matches + pull top 30
  feats <- gsub(" ", "_", feats) #clean names
  paste0(block, "_", feats)
})

##combine blocks with block-prefix feat names
all_feats <- do.call(cbind, lapply(names(final.diablo.model_data$X), function(b) {
  mat <- final.diablo.model_data$X[[b]]
  colnames(mat) <- paste0(b, "_", gsub(" ", "_", colnames(mat)))  #fill spaces w/underscore
  mat
}))

#correlation matrix across entire feature sets
cors_mat <- cor(all_feats, use="pairwise.complete.obs", method = "spearman")
diag(cors_mat) <- 0  #no self-corr
selected_feats <- intersect(unlist(top_features), colnames(cors_mat))
cors_mat <- cors_mat[selected_feats, selected_feats, drop = FALSE]
#feature filtering by keyword gets all matches from entire dataset ends up being 200 some features
# mod_feats <- grep(paste(keywords, collapse="|"), colnames(cors_mat),ignore.case=TRUE, value=TRUE)
# cors_mat <- cors_mat[unique(mod_feats),
#                      unique(mod_feats)]

### thresholding for correlated features of interset (taking abs value of 0.75 so can be highly pos/neg corr.)
edges <- which(abs(cors_mat) >= 0.75 & lower.tri(cors_mat), arr.ind=TRUE)
g <- graph_from_edgelist(cbind(rownames(cors_mat)[edges[,1]],
                               rownames(cors_mat)[edges[,2]]), directed=FALSE)
E(g)$weight <- cors_mat[edges]

###select colors/can update  this (colors of points for each blank)
V(g)$block <- sapply(strsplit(V(g)$name, "_"), `[`, 1)

####plot net
ggraph(g, layout="stress") +
  geom_edge_link(aes(color = weight), width=1) +
  geom_node_point(aes(color = block), size=3) +
  geom_node_text(aes(label = name), repel=TRUE, size=2) +
  scale_edge_color_gradient2(low="blue", mid="grey", high="red", midpoint=0,
                             name="Correlation") +
  theme_void() +
  theme(legend.position="right")

##extract features from most connected plot
features_largest_subgraph.key <- V(g)$name[components(g)$membership == which.max(components(g)$csize)]
View(data.frame(features_largest_subgraph.key))
length(features_largest_subgraph.key)



```



## Network 2: DIABLO id'd top features-only

```{r}

#### leaving only the top 30 DIABLO id'd features co-varying the most across each block
top_features <- lapply(names(final.diablo.model_data$X), function(b) {
  loadings_block <- final.diablo.model_data$loadings[[b]][, 1]       #comp1
  loadings_block <- loadings_block[!is.na(loadings_block)]
  feats <- names(sort(loadings_block, decreasing = TRUE))[1:30]      #top 30
  feats <- gsub(" ", "_", feats)                                     #replace spaces
  paste0(b,"_", feats)                                                   #add block prefix
})
names(top_features) <- names(final.diablo.model_data$X)

##combine blocks with block-prefix feat names
all_feats <- do.call(cbind, lapply(names(final.diablo.model_data$X), function(b) {
  mat <- final.diablo.model_data$X[[b]]
  colnames(mat) <- paste0(b, "_", gsub(" ", "_", colnames(mat)))  #fill spaces w/underscore
  mat
}))

#correlation matrix across entire feature sets
cors_mat3 <- cor(all_feats, use="pairwise.complete.obs", method = "spearman")
diag(cors_mat3) <- 0  #no self-corr
selected_feats <- intersect(unlist(top_features), colnames(cors_mat3))
cors_mat3 <- cors_mat3[selected_feats, selected_feats, drop = FALSE]
# cors_mat <- cor(all_feats, use="pairwise.complete.obs")
# dimnames(cors_mat) <- lapply(dimnames(cors_mat), function(x) gsub(" ", "_", x))
# cors_mat3 <- cors_mat[unlist(top_features) %>% as.vector(),
#                       unlist(top_features) %>% as.vector()]
# diag(cors_mat3) <- 0

### thresholding for correlated features of interset (taking abs value of 0.75 so can be highly pos/neg corr.)
edges <- which(abs(cors_mat3) >= 0.75 & lower.tri(cors_mat3), arr.ind=TRUE)
g <- graph_from_edgelist(cbind(rownames(cors_mat3)[edges[,1]],
                               rownames(cors_mat3)[edges[,2]]), directed=FALSE)
E(g)$weight <- cors_mat3[edges]

###select colors/can update  this (colors of points for each blank)
V(g)$block <- sapply(strsplit(V(g)$name, "_"), `[`, 1)

theme_bold <- theme(
  axis.title.x = element_text(size = 22, face = "bold"),
  axis.title.y = element_text(size = 22,face = "bold"),
  axis.text.x = element_text(face = "bold", size = 22),
  axis.text.y = element_text(size = 22,face = "bold"),
  legend.title = element_text(size = 22,face = "bold"),
  legend.text = element_text(size = 22,face = "bold"),
  plot.title = element_text(size = 22,face = "bold"),
  strip.text = element_text(size = 22,face = "bold"),
  axis.line = element_line(linewidth = 3, color = "black")
)

####plot net
ggraph(g, layout="stress") +
  geom_edge_link(aes(color = weight), width=1) +
  geom_node_point(aes(color = block), size=3) +
  geom_node_text(aes(label = name), repel=TRUE, size=2) +
  scale_edge_color_gradient2(low="blue", mid="grey", high="red", midpoint=0,
                             name="Correlation") +
  theme_void() +
  theme_bold +
  theme(legend.position="right")

##extract features from most connected plot
features_largest_subgraph.diablo <- V(g)$name[components(g)$membership == which.max(components(g)$csize)]
print((features_largest_subgraph.diablo))
length(features_largest_subgraph.diablo)
#features_largest_subgraph

```


## orgaanizing collective of features

I think it's kind of impressive here...
You can use the DIABLO graph and find about a third of features in the largest, connected subgraph

If you use a logical subset based on independent analysis of features 80% of those of features connect back to the suggested adipocyte hypertrophy/hyperplasia model mediated by incomplete carbohydrate fermentation and disrupted sulfur/bile metabolism

```{r}

library(dplyr)

# Example input: a character vector of features
##features_largest_subgraph.key
##features_largest_subgraph.diablo

adiposity_related_features <- c(
  "liver_Slc10a1","liver_Cbs","liver_Csad","liver_Fasn","liver_Acaca","liver_Acly","liver_Fads1","liver_Fads2","liver_Gpd1",
  "liver_Gpd2","liver_Insig2","liver_Hnf4a","liver_Igf1","liver_Cyb5r3",
  "liver_Acaa1b","liver_Retsat",
  "microbe_Campylobacter_sp._RM12651","microbe_Campylobacter_jejuni","microbe_Helicobacter_cinaedi",
  "microbe_Loigolactobacillus_coryniformis",
  "cecum_s,s-dimethyl-beta-propiothetin",
  "cecum_diisopropyl_sulfate",
  keywords
)
#keywords
#combine, collapse name, assign mechanism, and set 'from' label
combined_table <- bind_rows(
    tibble(feature = features_largest_subgraph.diablo, from = "DIABLO"),
    tibble(feature = features_largest_subgraph.key, from = "SingleOmicSelection")
  ) %>%
  mutate(
    feature = gsub(" ", "_", feature),
    mechanism = ifelse(
      grepl(paste(adiposity_related_features, collapse = "|"), feature),
      #"incomplete carb fermentation -> sulfur/bile disruption -> hypertrophy/hyperplasia",
      "model coherent",
      "Other"
    ),
    correlation_with_neighbors = sapply(feature, function(f) {
      cm <- if(f %in% features_largest_subgraph.diablo) cors_mat3 else cors_mat #select corr matrix to assess dirn
      if(!(f %in% colnames(cm))) return(NA)
      nbrs <- cm[f, setdiff(colnames(cm), f)]  #get negighbs above 0.75,
      nbrs <- nbrs[abs(nbrs) >= 0.75]
      if(length(nbrs) == 0) return(NA)
      s <- sign(nbrs)
      if(all(s > 0)) "positive"
      else if(all(s < 0)) "negative"
      else "both"
    })
  ) %>%
  group_by(feature, mechanism) %>%
  summarise(
    from = ifelse(all(c("DIABLO","SingleOmicSelection") %in% from), "both", unique(from)),
    correlation_with_neighbors = first(correlation_with_neighbors),
    .groups = "drop"
  ) %>%
  arrange(feature)
write.csv(combined_table, "combined_table.csv")



nrow(combined_table %>%
       filter((from == "SingleOmicSelection" | from == "both" & mechanism != "Other")))/length(c(features_largest_subgraph.key,features_largest_subgraph.diablo)) *100

nrow(combined_table %>%
       filter((from == "DIABLO" | from == "both" & mechanism != "Other")))/length(c(features_largest_subgraph.key,features_largest_subgraph.diablo)) *100

nrow(combined_table %>%
       filter((from == "DIABLO" | from == "both" & mechanism == "Other")))/length(c(features_largest_subgraph.key,features_largest_subgraph.diablo)) *100

print("some DIABLO features complement the features set we used to filter and project the first network, either because they were found jointly across investigations, or because they were not previously identified but consitute interactors in mechanisms of interest\n
      This is about 44% of the features")
nrow(combined_table %>%
       filter((from == "SingleOmicSelection" | from == "both" & mechanism != "Other") | (from == "DIABLO" & mechanism != "Other")))/length(c(features_largest_subgraph.key,features_largest_subgraph.diablo)) *100

print("The other 50% of the features seem to be more detox oriented...")
combined_table %>%
       filter(mechanism == "Other") %>% pull(feature)
combined_table %>%
       filter(mechanism != "Other") %>% pull(feature)


#Microbial metabolism: sulfur cycling, carbohydrate fermentation
#Host liver metabolism: lipid synthesis, bile acid transport, detoxification, retinol metabolism
### features both expand and link back to previous notions of host-microbe mechanism hypothesized to 
### coordinate sulfur and bile handling to mediated increased adiposity

print("features unique between largest connected subgraphs")
combined_table %>% filter(from != "both" & mechanism != "Other") %>% pull(feature) %>% length()
print("features shared across largest connected subgraphs of diablo/keyword-searched feature sets")
combined_table %>% filter(from == "both" & mechanism != "Other") %>% pull(feature) %>% length()
#### provide lots of additional context .




```




## suppl.


```{r, not convinced this is needed at all actually!!!!!! }
## keyword search for features supporting central (adipose hypertrophy) and supporting (adipose hyperplasia) mechanisms prev. suggested 


# collect selected features from your model
all_feats <- unlist(lapply(final.diablo.model_data$loadings, rownames))

#### barplot draft
# barplot(c(`Hypertrophy-feature sum` = sum(grepl(paste(hypert, collapse="|"), all_feats, ignore.case=TRUE)),
#           `Hyperplasia-features sum` = sum(grepl(paste(hyperp, collapse="|"), all_feats, ignore.case=TRUE))),
#         ylab="Count of features", col=c("steelblue","firebrick"),
#         main = "Count of features supporting hypertrophy/hyperplasia models")


```



```{r}

#substring for multiple features/ or write in specific one
#### bit of draft code for searching nearest neighbs of features of interest
feature_substring <- "microbe_Bacteroides_thetaiotaomicron"
features <- grep(feature_substring, V(g)$name, value = TRUE)  # returns all matching node names
#first neighb for all features pulled by substring
primary <- unique(unlist(lapply(features, function(f) neighbors(g, v = f, mode = "all"))))
primary_names <- V(g)$name[primary]
#secondary neighbs
secondary <- unique(unlist(lapply(primary, function(f) neighbors(g, v = f, mode = "all"))))
secondary_names <- setdiff(V(g)$name[secondary], c(features, primary_names))
#results
 features        #nodes match string
 primary_names   #directly connected nodes
 secondary_names #secondarily connected neighb


```




```{r}

#### removing top 30 from diablo and using the NEXT top 30 loadings instead

# #### leaving only the top 30 DIABLO id'd features co-varying the most across each block
# top_features <- lapply(names(final.diablo.model_data$X), function(b) {
#   loadings_block <- final.diablo.model_data$loadings[[b]][, 1]       #comp1
#   loadings_block <- loadings_block[!is.na(loadings_block)]
#   feats <- names(sort(loadings_block, decreasing = TRUE))[1:30]      #top 30
#   feats <- gsub(" ", "_", feats)                                     #replace spaces
#   paste0(b,"_", feats)                                                   #add block prefix
# })
# names(top_features) <- names(final.diablo.model_data$X)
# #top_features
# 
# #### removing top30 DIABLO id'd features co-varying the most across each block
# exclude_feats <- unlist(lapply(top_features, `[`, 1:30)) %>%
#                  gsub("-", "_", .)  # replace hyphens
# #match up for exclusion of top 20 diablo features
# matched_names <- colnames(cors_mat)[unique(unlist(lapply(exclude_feats, function(s) grep(s, colnames(cors_mat), fixed = TRUE))))]
# 
# cors_mat <- cor(all_feats, use="pairwise.complete.obs")
# #subset corr matrix
# cors_mat2 <- cors_mat[!rownames(cors_mat) %in% matched_names,
#                       !colnames(cors_mat) %in% matched_names]
# 
# diag(cors_mat2) <- 0  #no self-corr-ing
# #get top 30 per block
# new_top_features <- setNames(
#   lapply(c("microbe_", "serum_", "cecum_", "liver_"), function(pref) {
#     features_block <- grep(paste0("^", pref), colnames(cors_mat2), value = TRUE)
#     max_abs_cor <- sapply(features_block, function(f) max(abs(cors_mat2[f, ]), na.rm = TRUE))
#     names(sort(max_abs_cor, decreasing = TRUE))[1:30]
#   }),
#   c("microbe_", "serum_", "cecum_", "liver_")
# )
# cors_mat2 <- cors_mat2[rownames(cors_mat2) %in% unlist(new_top_features) %>% as.vector(),
#                       colnames(cors_mat2) %in% unlist(new_top_features) %>% as.vector()]
# 
# ### thresholding for correlated features of interset (taking abs value of 0.75 so can be highly pos/neg corr.)
# edges <- which(abs(cors_mat2) >= 0.75 & lower.tri(cors_mat2), arr.ind=TRUE)
# g <- graph_from_edgelist(cbind(rownames(cors_mat2)[edges[,1]],
#                                rownames(cors_mat2)[edges[,2]]), directed=FALSE)
# E(g)$weight <- cors_mat2[edges]
# 
# ###select colors/can update  this (colors of points for each blank)
# V(g)$block <- sapply(strsplit(V(g)$name, "_"), `[`, 1)
# 
# ## make sure we're subsetting logically
# 
# ####plot net
# ggraph(g, layout="stress") +
#   geom_edge_link(aes(color = weight), width=1) +
#   geom_node_point(aes(color = block), size=3) +
#   geom_node_text(aes(label = name), repel=TRUE, size=2) +
#   scale_edge_color_gradient2(low="blue", mid="grey", high="red", midpoint=0,
#                              name="Correlation") +
#   theme_void() +
#   theme(legend.position="right")
# 
# ##extract features from most connected plot
# features_largest_subgraph <- V(g)$name[components(g)$membership == which.max(components(g)$csize)]
# #features_largest_subgraph
# length(features_largest_subgraph)
# 
# 
# #wow very replete and uninformative feature set as it relates to our proposed model after removing top 30 across blocks and then grabbing the next top 30 per block... to me this signals the importance and performance of diablos work in feature selection

```


```{r}

block_filters <- list(
 microbiome = c("microbe_Bacteroides_thetaiotaomicron","scindens", "symbiosum", "sulf"),
 cecal_metabolome  = c("carb", "mannitol", "buty", "buta", "propa", "propi", "sterol", "chol", "sacch", "ose"),
 serum_metabolome  = c("carb", "mannitol", "buty", "buta", "propa", "propi", "sulf", "cyst", "sterol", "chol", "sacch", "ose"),
 liverA = c("Csad", "Cdo1", "Fmo1", "Gclc", "Cth", "Mfsd"),
 liverB = c("Cyp", "Hnf4a", "Slc10a1"),
 liverC = c("Rxr", "Nr1h3","Nr1h4", "Nr1i3", "Nr1i2", "Retsat", "Rbp1"),
 liverD = c("Fasn", "Acaca", "Acacb", "Sreb", "Scd1", "Acly", "Slc27a5"),
 liverE = c("Slc6a6","Slc6a13","Slc", "Ctnnb", "Wnt", "Mfrp", "Fzd"),
 liverF = c("Ppar", "Cebp", "Sreb", "Klf", "Fabp", "Adipoq", "Cidec", "Gpd1", "Stat5", "Wnt", "Ctnnb")
)


all_paths <- list(
 c("microbiome","cecal metabolome","serum metabolome","liverA","liverB","liverE", "liverF"),
 c("microbiome","cecal metabolome","serum metabolome","liverA","liverB","liverC","liverD"),
 c("microbiome","cecal metabolome","serum metabolome","liverA","liverB")
)







write_xlsx(x =  , path = "4nclip.xlsx", col_names = TRUE, format_headers = TRUE)

```




```{r} 

#Updated code 10/17/2025
set.seed(123)

# # define hypothesis feature lists
keywords <- c(
  "Lachno", "thetaiotaomicron", "scindens", "symbiosum","hiranonis",
  "Csad","Cdo1","Fmo1","Stra","Aldh","Rdh","Scd",
  "Rbp","Fasn","Acaca","Elovl","Retsat","Cyp","Slc10a1","Hsd","Ctnnb1","Wnt","Ppar","Sreb","Abcb11","Sfrp1","Fzd",
  "carb","sulf","chol","taur","oxy","stero","glucos","mann",
  "acetyl","phos","sugar","glyco","sacch","retin","malonyl", "palmit", "steara","acyl", "olea",
  "sialic","fucose","hexose","pentose","heptose","octulose","uronic","lactone"
)

top_features_filt <- lapply(names(final.diablo.model_data$X), function(block) {
  load_block <- final.diablo.model_data$loadings[[block]][, 1]   #loadings comp1
  load_block <- load_block[!is.na(load_block)] #chuck NA
  keep_idx <- grepl(paste(keywords, collapse = "|"), names(load_block), ignore.case = TRUE) #keyword pull
  load_block <- load_block[keep_idx]
  feats <- names(sort(abs(load_block), decreasing = TRUE))[1:min(30, length(load_block))] #get abs vals for loading matches + pull top 30
  feats <- gsub(" ", "_", feats) #clean names
  paste0(block, "_", feats)
})
#top_features

##combine blocks with block-prefix feat names
all_feats <- do.call(cbind, lapply(names(final.diablo.model_data$X), function(b) {
  mat <- final.diablo.model_data$X[[b]]
  colnames(mat) <- paste0(b, "_", gsub(" ", "_", colnames(mat)))  #fill spaces w/underscore
  mat
}))

#correlation matrix across entire feature sets
cors_mat <- cor(all_feats, use="pairwise.complete.obs", method = "spearman")
diag(cors_mat) <- 0  #no self-corr
selected_feats <- intersect(unlist(top_features_filt), colnames(cors_mat))
cors_mat2 <- cors_mat[selected_feats, selected_feats, drop = FALSE]
#feature filtering by keyword gets all matches from entire dataset ends up being 200 some features
# mod_feats <- grep(paste(keywords, collapse="|"), colnames(cors_mat),ignore.case=TRUE, value=TRUE)
# cors_mat <- cors_mat[unique(mod_feats),
#                      unique(mod_feats)]

### thresholding for correlated features of interset (taking abs value of 0.75 so can be highly pos/neg corr.)

edges <- which(abs(cors_mat2) >= 0.7 & lower.tri(cors_mat2), arr.ind=TRUE)
g <- graph_from_edgelist(
  cbind(rownames(cors_mat2)[edges[,1]], rownames(cors_mat2)[edges[,2]]),
  directed = FALSE
)

### assign block info
V(g)$block <- sapply(strsplit(V(g)$name, "_"), `[`, 1)
#magnitude for layout
E(g)$weight <- abs(cors_mat2[edges])
#signed correlation
E(g)$weight_signed <- cors_mat2[edges]   

#get largest subgraph 
comps <- components(g)
largest_comp_id <- which.max(comps$csize)
V(g)$in_largest <- comps$membership == largest_comp_id

###pagerank/consistent w trying to assess  directionality + consistent w/ metabolomics
E(g)$pr_weight <- abs(E(g)$weight_signed) / max(abs(E(g)$weight_signed))
pr <- page_rank(g, directed = TRUE, weights = E(g)$pr_weight)$vector

#centraliry thresh for pagerank score
pr_thresh <- quantile(pr[V(g)$in_largest], 0.7)
V(g)$is_hub <- V(g)$in_largest & pr >= pr_thresh

## color the nodes And add labels
V(g)$node_color <- ifelse(V(g)$is_hub, V(g)$block, "gray70")
V(g)$label_text <- ifelse(V(g)$in_largest, V(g)$name, "")
V(g)$label_color <- ifelse(V(g)$node_color == "gray70", "gray70", "black")
###color pallete
dark2_cols <- brewer.pal(4, "Dark2")
names(dark2_cols) <- c("microbe", "cecum", "serum", "liver")

hub_labels <- V(g)$label_text
hub_labels[!V(g)$is_hub] <- ""

nonhub_labels <- V(g)$label_text
nonhub_labels[V(g)$is_hub] <- ""

#jpeg("filtered_network_1.jpeg", units="in", height = 16, width= 20, res=600)
jpeg("filtered_network_1.jpeg", units="cm", height = 60, width=90, res=600)
ggraph(g, layout = layout_with_kk(g, weights = E(g)$weight)) +
  geom_edge_link(aes(
    color = weight_signed), 
    width = 0.8) +
  geom_node_point(aes(
    color = node_color), 
    size = 8, 
    show.legend = TRUE) +
  # geom_node_text(aes(
  #   label = label_text),
  #   color = V(g)$label_color,
  #   repel = TRUE, 
  #   size = 8,
  #   fontface = "bold",
  #   show.legend = FALSE
  # ) +
  geom_node_text(
    aes(label = nonhub_labels),
    color = "gray50",
    repel = TRUE, 
    size = 8, 
    show.legend = FALSE
    ) +
  geom_node_text(
    aes(label = hub_labels),
    color = "black",
    repel = TRUE, 
    size = 8, 
    fontface = "bold", 
    show.legend = FALSE
    ) +
  scale_edge_color_gradient2(
    low = "skyblue1", mid = "grey90", high = "red",
    midpoint = 0, limits = c(-1, 1),
    name = "Relative Correlation"
  ) +
  scale_color_manual(
    values = dark2_cols,
    name = "Omic Block",
    breaks = c("microbe","cecum","serum","liver"),
    labels = c("Microbiome","Cecum Metabolome","Serum Metabolome","Liver Transcriptome")
  ) +
  theme_void() +
  theme(legend.position = c(0.95, 0.6),
        legend.text = element_text(size = 20, face = "bold"),
        legend.title = element_text(size = 20, face = "bold"),
        legend.key.size = unit(1, "cm"),
        panel.grid = element_blank()
  )
dev.off()

# ##Pull out counts of features per block in this keyword filtered network
# Featuresperblock <- table(V(g)$block[V(g)$label_color == "black"])
# print(Featuresperblock)

#print(keyword)

#ggsave("filtered_keyword_subset.png", plot = keyword, width=16, height=9, dpi=600, bg = "white")

# ##extract features from most connected plot
comps <- components(g)
largest_comp <- which.max(comps$csize)
features_largest_subgraph.key <- V(g)$name[comps$membership == largest_comp]
filtered_datakey_suppl1 <- data.frame(
  feature = V(g)$name[which(comps$membership == largest_comp)],
  color   = V(g)$label_color[which(comps$membership == largest_comp)]
  )

# ##extract features from most connected plot
# comps <- components(g)
# largest_comp <- which.max(comps$csize)
# features_largest_subgraph.key <- V(g)$name[comps$membership == largest_comp]
# #View(data.frame(features_largest_subgraph.key))
# #length(features_largest_subgraph.key)

```


```{r}

##updated code 10/17/2025

set.seed(123)
#### leaving only the top 30 DIABLO id'd features co-varying the most across each block
top_features <- lapply(names(final.diablo.model_data$X), function(b) {
  loadings_block <- final.diablo.model_data$loadings[[b]][, 1]       #comp1
  loadings_block <- loadings_block[!is.na(loadings_block)]
  feats <- names(sort(loadings_block, decreasing = TRUE))[1:30]      #top 30
  feats <- gsub(" ", "_", feats)                                     #replace spaces
  paste0(b,"_", feats)                                                   #add block prefix
})
names(top_features) <- names(final.diablo.model_data$X)

top_features
##combine blocks with block-prefix feat names
all_feats <- do.call(cbind, lapply(names(final.diablo.model_data$X), function(b) {
  mat <- final.diablo.model_data$X[[b]]
  colnames(mat) <- paste0(b, "_", gsub(" ", "_", colnames(mat)))  #fill spaces w/underscore
  mat
}))

#correlation matrix across entire feature sets
cors_mat3 <- cor(all_feats, use="pairwise.complete.obs", method = "spearman")
diag(cors_mat3) <- 0  #no self-corr
selected_feats <- intersect(unlist(top_features), colnames(cors_mat3))
cors_mat3 <- cors_mat3[selected_feats, selected_feats, drop = FALSE]
# cors_mat <- cor(all_feats, use="pairwise.complete.obs")
# dimnames(cors_mat) <- lapply(dimnames(cors_mat), function(x) gsub(" ", "_", x))
# cors_mat3 <- cors_mat[unlist(top_features) %>% as.vector(),
#                       unlist(top_features) %>% as.vector()]
# diag(cors_mat3) <- 0

### thresholding for correlated features of interset (taking abs value of 0.75 so can be highly pos/neg corr.)
edges <- which(abs(cors_mat3) >= 0.7 & lower.tri(cors_mat3), arr.ind=TRUE)
g <- graph_from_edgelist(
  cbind(rownames(cors_mat3)[edges[,1]], rownames(cors_mat3)[edges[,2]]),
  directed = FALSE
)

### assign block info
V(g)$block <- sapply(strsplit(V(g)$name, "_"), `[`, 1)
E(g)$weight <- abs(cors_mat3[edges])       # magnitude for layout
E(g)$weight_signed <- cors_mat3[edges]     # signed correlation

#get largest subgraph 
comps <- components(g)
largest_comp_id <- which.max(comps$csize)
V(g)$in_largest <- comps$membership == largest_comp_id

###pagerank/consistent w trying to assess  directionality + consistent w/ metabolomics
E(g)$pr_weight <- abs(E(g)$weight_signed) / max(abs(E(g)$weight_signed))
pr <- page_rank(g, directed = TRUE, weights = E(g)$pr_weight)$vector

#centraliry thresh for pagerank score
pr_thresh <- quantile(pr[V(g)$in_largest], 0.7)
V(g)$is_hub <- V(g)$in_largest & pr >= pr_thresh

## color the nodes And add labels
V(g)$node_color <- ifelse(V(g)$is_hub, V(g)$block, "grey70")
V(g)$label_text <- ifelse(V(g)$in_largest, V(g)$name, "")
V(g)$label_color <- ifelse(V(g)$node_color == "grey70", "grey90", "black")

## Set label size based on label color
#V(g)$label_size <- ifelse(V(g)$label_color == "black", 8, 6)

###color pallete
dark2_cols <- brewer.pal(4, "Dark2")
names(dark2_cols) <- c("microbe", "cecum", "serum", "liver")

hub_labels <- V(g)$label_text
hub_labels[!V(g)$is_hub] <- ""

nonhub_labels <- V(g)$label_text
nonhub_labels[V(g)$is_hub] <- ""

#jpeg("unfiltered_network_2.jpeg", units="in", height = 16, width= 20, res=600)
jpeg("unfiltered_network_2.jpeg", units="cm", height = 60, width= 90, res=600)
ggraph(g, layout = layout_with_kk(g, weights = E(g)$weight)) +
  geom_edge_link(aes(
    color = weight_signed), 
    width = 0.8) +
  geom_node_point(aes(
    color = node_color), 
    size = 8, 
    show.legend = TRUE) +
  geom_node_text(
    aes(label = nonhub_labels),
    color = "gray50",
    repel = TRUE, size =8, show.legend = FALSE
    ) +
  geom_node_text(
    aes(label = hub_labels),
    color = "black",
    repel = TRUE, size = 8, fontface = "bold", show.legend = FALSE
    ) +
  scale_size_identity() +
  scale_edge_color_gradient2(
    low = "skyblue1", mid = "grey90", high = "red",
    midpoint = 0, limits = c(-1, 1),
    name = "Relative Correlation"
  ) +
  scale_color_manual(
    values = dark2_cols,
    name = "Omic Block",
    breaks = c("microbe","cecum","serum","liver"),
    labels = c("Microbiome","Cecum Metabolome","Serum Metabolome","Liver Transcriptome")
  ) +
  theme_void()+
  theme(legend.position = c(0.95, 0.6),
        legend.text = element_text(size = 20, face = "bold"),
        legend.title = element_text(size = 20, face = "bold"),
        legend.key.size = unit(1, "cm"),
        panel.grid = element_blank()
  )
dev.off()

# ##features per block in this diablo top-feature only network
# diablofeatures<-table(V(g)$block[V(g)$label_color == "black"])
# print(diablofeatures)

##extract features from most connected plot
comps <- components(g)
largest_comp <- which.max(comps$csize)
features_largest_subgraph.key <- V(g)$name[comps$membership == largest_comp]
filtered_datakey_suppl1 <- data.frame(
  feature = V(g)$name[which(comps$membership == largest_comp)],
  color   = V(g)$label_color[which(comps$membership == largest_comp)]
  )

# ##extract features from most connected plot
# features_largest_subgraph.diablo <- V(g)$name[comps$membership == largest_comp]
# print(features_largest_subgraph.diablo)
# length(features_largest_subgraph.diablo)
# #features_largest_subgraph

```
