---
title: "2_cecal_shotgun"
output: html_document
date: "2024-04-05"
---

Kraken2/Bracken output first

```{r, load packs}

set.seed(1) # for reproducible stochastic processes
library(devtools)
#BiocManager::install("phyloseq")
library(phyloseq)
library(ggplot2)
library(readr)
library(stringr)
library(tibble)
library(dplyr)
library(cowplot) #for combining plots
library(tidyr)
library(patchwork) # for combining multiple plots
#BiocManager::install("microViz")
#if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
#BiocManager::install(c("phyloseq", "microbiome", "ComplexHeatmap"), update = FALSE)
# install.packages("microViz", 
#                  repos = c(davidbarnett = "https://david-barnett.r-universe.dev", getOption("repos")
#                            )
#                  )
# install.packages("ggtext") # for rotated labels on ord_plot() 
# install.packages("ggraph") # for taxatree_plots()
# install.packages("DT") # for tax_fix_interactive()
# install.packages("remotes")
# remotes::install_github("statdivlab/corncob") # for example datasets and beta binomial models
library(ggtext)
library(ggraph)
library(DT)
library(data.table)
library(corncob)
library(reshape2)
#BiocManager::install("microbiome")
library(microbiome)
#BiocManager::install("ComplexHeatmap")
library(ComplexHeatmap)
#install.packages("ggheatmap")
library(ggheatmap)
#install.packages("microViz",repos = c(davidbarnett = "https://david-barnett.r-universe.dev", getOption("repos")))
library(microViz)
library(ape)
library(tibble)
#install.packages("RColorBrewer")
library(RColorBrewer)
#install.packages("ggpubr")
library(ggpubr)
library(vegan)
#install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
library(pairwiseAdonis)
library(gridExtra)
#install.packages("ggh4x")
library(ggh4x)
library(igraph)
library(writexl)

```




```{r, formatting }

#### making draft otu table
getwd()
otu_mat <- data.frame(read_tsv("bracken_abundance_species_mpa.txt"))
otu_mat <- otu_mat %>% column_to_rownames("X.Classification") %>% as.matrix()
#otu_mat2 <- otu_mat[,-c(4,5,7,8,10,11,13,15)]


### making draft tax table
taxmat <- data.table(rownames(otu_mat))
taxmat[, paste0(c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species")) := tstrsplit(V1, ",") ]
taxmat$Kingdom <- str_extract(taxmat$Kingdom, "k__[A-Z]+[a-z]+")
taxmat$Phylum <- str_extract(taxmat$Phylum, "p__[A-Z]+[a-z]+")
taxmat$Class <- str_extract(taxmat$Class, "c__[A-Z]+[a-z]+")
taxmat$Order <- str_extract(taxmat$Order, "o__[A-Z]+[a-z]+")
taxmat$Family <- str_extract(taxmat$Family, "f__[A-Z]+[a-z]+")
taxmat$Genus <- str_extract(taxmat$Genus, "g__[A-Z]+[a-z]+")
taxmat$Species <- str_extract(taxmat$Species, "s__.*")
taxmat <- taxmat %>% column_to_rownames(var = "V1")
#taxmat %>% filter(taxmat, taxmat$Species != "NA")
no<-na.omit(taxmat)

#newnames <- paste(as.vector(taxmatv0[,1]),"\\|",as.vector(taxmatv0[,2]),"\\|",as.vector(taxmatv0[,3]), sep = "")
#rownames(otu_matv01) <- newnames

#rownames(taxmatv0) <- rownames(otu_matv01)
#rownames(taxmatv0) <- paste(as.vector(taxmatv0[,1]),"|",as.vector(taxmatv0[,2]),"|",as.vector(taxmatv0[,3]), sep = "")
#rownames(taxmatv0) <- paste(as.vector(taxmatv0[,1]))
#rownames(taxmatv0) <- newnames


### making draft metadata
metadata <- read.csv("metadata.txt")
metadata2 <- metadata[,-1]
metadata2$file <- metadata2$sample_name
rownames(metadata2) <- metadata[,1]
metadata2$sample_name <- paste(metadata2$group, substr(rownames(metadata2), start = 2, stop = 10), sep = "")
metadata2$file <- metadata2$sample
metadata2$sample <- metadata2$sample_name
rownames(metadata2) <- metadata2$sample_name

### fixing otu colnames to match metadata for phyloseq obj init
colnames(otu_mat) <- metadata2$sample_name
colnames(otu_mat) <- rownames(metadata2)

### setting up phyloseq obj components
OTU <- otu_table(as.matrix(otu_mat), taxa_are_rows = TRUE)
TAX <- tax_table(as.matrix(no))
META <- sample_data(metadata2)

### init phyloseq obj and print obj overview
physeq <- phyloseq(OTU, TAX, META)
#View(physeq@tax_table)
# used unfiltered data and get same result/can ignore warning for this dataset; results looked the identical
#View(OTU)

```


## Taxonomic barplots and qc
```{r, preprocessing and top taxonomic compositional abundances}


theme_bold <- theme(
  axis.title.x = element_text(size = 22, face = "bold"),
  axis.title.y = element_text(size = 22,face = "bold"),
  axis.text.x = element_text(face = "bold", size = 22),
  axis.text.y = element_text(size = 22,face = "bold"),
  legend.title = element_text(size = 22,face = "bold"),
  legend.text = element_text(size = 22,face = "bold"),
  plot.title = element_text(size = 22,face = "bold"),
  strip.text = element_text(size = 22,face = "bold"),
  axis.line = element_line(size = 3, color = "black")
)


#setwd("/Users/apulvino/AmatoLab/2_cecal_shotgun/")
# https://joey711.github.io/phyloseq-demo/Restroom-Biogeography
any(taxa_sums(physeq) == 0)
sum(taxa_sums(physeq) == 0)

physeq = prune_taxa(taxa_sums(physeq) > 0, physeq)

any(sample_sums(physeq) == 0)
ps = prune_samples(sample_sums(physeq) > 0, physeq)

readsumsdf = data.frame(nreads = sort(taxa_sums(ps), TRUE), sorted = 1:ntaxa(ps), 
    type = "OTUs")
readsumsdf = rbind(readsumsdf, data.frame(nreads = sort(sample_sums(ps), 
    TRUE), sorted = 1:nsamples(ps), type = "Samples"))

p = ggplot(readsumsdf, aes(x = sorted, y = nreads)) + geom_bar(stat = "identity")
p + ggtitle("Total Number of Reads per Sample") + scale_y_log10() + facet_wrap(~type, 1, scales = "free") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  xlab("Sample") + ylab("Number of Reads")

#ps = prune_samples(sample_sums(ps) > 500, ps)

## Transform to equal sample sum
#library(compositions)
#psP = transform_sample_counts(ps, function(OTU) clr(OTU))
psP = transform_sample_counts(ps, function(OTU) OTU/sum(OTU) * 100)

# par(mfrow = c(1, 2))
# title = "Sum of reads for each sample"
# plot(sort(sample_sums(psP), TRUE), type = "h", main = title, ylab = "reads", 
#     ylim = c(0, 1000))
# title = "Sum of reads for each sample"
# plot(sort(sample_sums(psP), TRUE), type = "h", main = title, ylab = "reads", 
#     ylim = c(0, 1000))
top19otus = names(sort(taxa_sums(psP), TRUE)[1:55])
taxtab19 =data.frame( cbind(tax_table(psP), top20 = "Subsequent Taxa"))
taxtab19[top19otus, "top20"] <- as(tax_table(psP)[top19otus, "Genus"], 
    "character")
taxtab19$top20<- sub("g__", "",taxtab19$top20)

A1leg.ttl = 2.5
A1leg.txt =2.5
A1axis.ttl.y = 2.5
A1axis.txt.y =2.5
A1hum = 1
A1mac = 1
A1sqm = 1
A1strip.txt = 28

# PAPER 1
# 20 top genera
tax_table(psP) <- tax_table(as.matrix(taxtab19))
jpeg("qual2_taxplot1.jpeg", units="in", width=20, height=12, res=300)
phyloseq::plot_bar(psP, fill = "top20") + 
  theme(legend.title = element_text(size = rel(A1leg.ttl), face = "bold"),
        legend.text = element_text(size = rel(A1leg.txt), face = "bold", color = "black"),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = rel(A1axis.ttl.y), face = "bold", color = "black"),
        axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size = rel(A1axis.txt.y), color = "black", face = "bold"),
        #axis.title.x = element_text(size = rel(1.5)),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        #strip.background = element_rect(fill="black"),
        strip.text=element_text(color="black", face = "bold", size = A1strip.txt)
        )  + 
  scale_y_continuous(expand = c(0,0)) +
  expand_limits(y=100.25) +
  xlab("Donor Group") + 
  ylab("Percent Abundance") +
  scale_fill_manual("Top 20 Abundant Genera",values = colorRampPalette(brewer.pal(8, "Dark2"))(21)) +
  ggh4x::facet_wrap2(~group,
                     scales = "free_x",
                     strip = strip_themed(background_x = list(hum = element_rect(fill = "#1b9e77", size = A1hum),
                                                              mac = element_rect(fill = "#a6761d", size = A1mac),
                                                              sqm = element_rect(fill = "#666666", size = A1sqm)
                                                              )
                                          )
                     )

dev.off()

psP = transform_sample_counts(ps, function(OTU) OTU/sum(OTU) * 100)
top19otus = names(sort(taxa_sums(psP), TRUE)[1:20])
taxtab19 =data.frame( cbind(tax_table(psP), top20 = "Subsequent Taxa"))
taxtab19[top19otus, "top20"] <- as(tax_table(psP)[top19otus, "Species"], 
    "character")
taxtab19$top20 <- taxtab19$top20 %>%
  stringr::str_replace("^s__", "") %>%
  stringr::str_replace_all("_", " ")


#### write out phyloseq obj for HsMmSb_MouseRNAseq.Rmd/for initial spearman dim red analysis PAPER 1
write_rds(psP, "C:/Users/willh/OneDrive - Loyola University Chicago/Documents/Mouse_data_AnthonyP2025/LiverRNAseqPackage_19082025/LiverRNAseqPackage_19082025/psP.Rds")


##work on making font bigger and plots wider

A2leg.ttl = 2.25
A2leg.txt =2.25
A2axis.ttl.y = 2.5
A2axis.txt.y = 3
A2hum = 1
A2mac = 1
A2sqm = 1
A2strip.txt = 28

#set labels
label_vec <- c("hum", "mac", "sqm")
styled_labels <- setNames(
  paste0("<span style='color:white; text-shadow:-1px -1px 0 black, 1px -1px 0 black, -1px 1px 0 black, 1px 1px 0 black;'>", label_vec, "</span>"),
  label_vec
)

facet_wrap2(~group, labeller = as_labeller(styled_labels))

## abundance plot for top 10
# PAPER 1
# 20 top species
tax_table(psP) <- tax_table(as.matrix(taxtab19))
# sample_data(psP) <- data.frame(sample_data(psP)) %>%
  # mutate(group = case_when(group %in% "hum" ~ "human",
   #                                group %in% "mac" ~ "macaque",
    #                               group %in% "sqm" ~ "squirrel monkey"
     #                              ))
tiff("qual2_taxplot2.black.tiff", units="in", width=20, height=10, res=300)
 phyloseq::plot_bar(psP, fill = "top20") + 
  theme(legend.title = element_text(size = rel(A2leg.ttl), face = "bold"),
        legend.text = element_text(size = rel(A2leg.txt), face = "bold", color = "black"),
        #axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = rel(1.5), color = "black"),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = rel(A2axis.ttl.y), face = "bold", colour = "black"),
        axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size = rel(A2axis.txt.y), color = "black", face = "bold"),
        #axis.title.x = element_text(size = rel(1.5)),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        #strip.background = element_rect(fill="black"),
        strip.text = ggtext::element_markdown(face = "bold", size = A2strip.txt)
        ) +
  scale_y_continuous(expand = c(0,0)) +
  expand_limits(y=100.25) +
  xlab("Donor Group") + 
  ylab("Percent Abundance") +
  scale_fill_manual("Top 20 Abundant Species",values = colorRampPalette(brewer.pal(8, "Dark2"))(21))+
  ggh4x::facet_wrap2(~group,
                     scales = "free_x",
                     strip = strip_themed(background_x = list(hum = element_rect(fill = "#1b9e77"),
                                                              mac = element_rect(fill = "#a6761d"),
                                                              sqm = element_rect(fill = "#666666") +theme_bold 
                                                              )
                                          )
                     )



dev.off()



# write csv for tax table
#write.csv(taxtab19, "taxtab19.csv")


```


## Dim. reduction/plot

```{r, sample distances, dim red, PCoA/nmds}

library(ggrepel)
set.seed(123)
#options(device = "quartz")
## sample distance
# PAPER 1

p1axis.txt = 27
p1axis.ttl = 22
p1plot.ttl = 22
p1elipse.size = 2
p1mac.label = 8
p1hum.label = 8
p1sqm.label = 8

psp_ord = ordinate(psP, "NMDS", "bray")
tiff("qual2_distplot1.good.tiff", units="in", width=8, height=8, res=300)
bray <- plot_ordination(psP, psp_ord, color = "group", type = "group") + 
  geom_point(size = 6, alpha = 0.7, stat = "identity") + 
  stat_ellipse(size = p1elipse.size) + 
  theme(panel.grid.major = element_blank(), aspect.ratio = 1/2.5, panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black", size = 4), axis.text = element_text(size = p1axis.txt, face = "bold", colour = "black"), axis.title = element_text(size = p1axis.ttl, face = "bold"), plot.title = element_text(size = p1plot.ttl, face = "bold", hjust = 0.5),
        legend.position="none") +
  labs(title = "NMDS of Bray-Curtis distance", "Donor Group\n" )+ 
  annotate("text",  x = 0.1, y = -0.35, label = "sqm", color  = "#666666", fontface = "bold", size = p1sqm.label) +
  annotate("text", x = -0.38, y = 0.25, label = "hum", color = "#1b9e77", fontface = "bold", size = p1hum.label) +
  annotate("text", x = 1.2, y = 0.43, label = "mac", color = "#a6761d", fontface = "bold", size = p1mac.label) +
  scale_color_manual(values = c(hum="#1b9e77",mac="#a6761d",sqm="#666666"))

bray

dev.off()

######### anybody and everybody
#psp_ord = ordinate(psP, "PCoA", "bray")
#tiff("qual2_distplot2.tiff", units="in", width=11, height=11, res=300)
#plot_ordination(psP, psp_ord, color = "group", type = "sites")+
#  geom_point(size = 6, alpha = 0.7, stat = "identity") +
 # stat_ellipse(type = "t", linetype = "solid", size = 0.8) +
#  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
 #       panel.background = element_blank(), axis.line = element_line(colour = "black"),
  #      legend.position = "none") +
  #labs(title = "PCoA of Bray-Curtis distance", color = "Donor Group\n")+ 
  #annotate("text", x = 0.10, y = -0.3, label = "sqm", color = "#666666") +
  #annotate("text", x = 0.35, y = 1, label = "mac", color = "#a6761d") +
  #annotate("text", x = -0.22, y = 0.27, label = "hum", color = "#1b9e77")  +
#  scale_color_manual(values = c(hum="#1b9e77",mac="#a6761d",sqm="#666666")) 
#dev.off()



```


## Shannon plots 
```{r}
## PAPER 1
# shannon alpha div
shannonp <- plot_richness(physeq, x="group", measures=c("Shannon")) 


theme_bold <- theme(
  axis.title.x = element_text(size = 22, face = "bold"),
  axis.title.y = element_text(size = 22,face = "bold"),
  axis.text.x = element_text(face = "bold", size = 22),
  axis.text.y = element_text(size = 22,face = "bold"),
  legend.title = element_text(size = 22,face = "bold"),
  legend.text = element_text(size = 22,face = "bold"),
  plot.title = element_text(size = 22,face = "bold"),
  strip.text = element_text(size = 22,face = "bold"),
  axis.line = element_line(size = 3, color = "black")
)


ap1 <- ggboxplot(shannonp$data, x = "group", y = "value", size = 2,
          color = "group", order = c("hum", "mac","sqm")) +
  stat_compare_means(aes(group = group) , method = 'wilcox.test', label = "p.signif", bracket.size = 2,
                     comparisons = list( c("hum", "mac"), c("sqm", "mac")), size = 7,
                      vjust = .5 , label.y = 4.6, label.x = 8, step.increase = .1
             ) +  
  xlab("") + ylab("") + ggtitle("Shannon Diversity") +
  geom_point(stat = "identity") +
  scale_color_manual(values = c(hum="#1b9e77",mac="#a6761d",sqm="#666666"),guide = "none") +
  theme(aspect.ratio = 1/2,legend.position="none") + theme_bold + theme(
    aspect.ratio = 1/2,
    legend.position = "none",
    axis.text.x = element_text(face = "bold", size = 22),
    axis.title.x = element_text(size = 22, face = "bold"),
    axis.ticks.x = element_line(size = 3, color = "black")
  ) + theme_bold

#to save un comment these lines!!
#ap1
#ggsave("shannondiversity.11.tiff", device = "jpeg")

ap2 <- ggboxplot(plot_richness(physeq, x="group", measures=c("Simpson"))$data,
          x = "group", y = "value", size = 2, color = "group", order = c("hum", "mac","sqm")) +
  stat_compare_means(aes(group = group), method = 'wilcox.test', label = "p.signif", bracket.size = 2,
                     comparisons = list( c("hum", "mac"), c("sqm", "mac")), size = 5.5,
                      vjust = .5 ,  step.increase = .1
  ) + 
  xlab("") + ylab("") + ggtitle("Simpson Diversity") +
  geom_point(stat = "identity") +
  scale_color_manual(values = c(hum="#1b9e77",mac="#a6761d",sqm="#666666")) +
  theme(legend.position="none",
        axis.text.x = element_blank(),  
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank()) + theme_bold+ theme(
    aspect.ratio = 1/2,
    legend.position = "none",
    axis.text.x = element_text(face = "bold", size = 22),
    axis.title.x = element_text(size = 22, face = "bold"),
    axis.ticks.x = element_line(size = 3, color = "black")
  ) + theme_bold

ap3 <- ggboxplot(plot_richness(physeq, x="group", measures=c("Chao1"))$data,
          x = "group", y = "value", size = 2,color = "group", order = c("hum", "mac","sqm")) +
  stat_compare_means(aes(group = group), method = 'wilcox.test', label = "p.signif", bracket.size = 2,
                     comparisons = list( c("hum", "mac"), c("sqm", "mac")), size = 4.4,
                      vjust = ,  step.increase = .1
  ) + 
  xlab("") + ylab("") + ggtitle("Chao Diversity") +
  geom_point(stat = "identity") +
  scale_color_manual(values = c(hum="#1b9e77",mac="#a6761d",sqm="#666666")) +
  theme(legend.position="none")  + theme_bold+ theme(
    aspect.ratio = 1/2,
    legend.position = "none",
    axis.text.x = element_text(face = "bold", size = 22),
    axis.title.x = element_text(size = 22, face = "bold"),
    axis.ticks.x = element_line(size = 3, color = "black")
  ) + theme_bold

ap4 <- ggboxplot(plot_richness(physeq, x="group", measures=c("Observed"))$data,
          x = "group", y = "value", size = 2, color = "group", order = c("hum", "mac","sqm")) +
  stat_compare_means(aes(group = group),
                     method = 'wilcox.test',
                     label = "p.signif",
                     bracket.size = 2,
                     comparisons = list( c("hum", "mac"), c("sqm", "mac")), size = 5.5,
                      vjust = .5 , step.increase = .1) + 
  xlab("") + ylab("") + ggtitle("Observed Diversity") +
  geom_point(stat = "identity") +
  scale_color_manual(values = c(hum="#1b9e77",mac="#a6761d",sqm="#666666")) +
  theme(legend.position="none") + theme_bold + theme(
    aspect.ratio = 1/2,
    legend.position = "none",
    axis.text.x = element_text(face = "bold", size = 22),
    axis.title.x = element_text(size = 22, face = "bold"),
    axis.ticks.x = element_line(size = 3, color = "black")
  ) + theme_bold

tiff("diversity.comp.tiff", units="in", width=12, height=8, res=300)

grid.arrange(ap1, ap2, ap3, ap4,nrow=2 )

dev.off


```






```{r, permanova}



META$mac <- if_else(data.frame(META)$group == "mac", 1, 0)
META$hum <- if_else(data.frame(META)$group == "hum", 1, 0)
META$sqm <- if_else(data.frame(META)$group == "sqm", 1, 0)
## permanova stats
## broader effect of group, significant
print("global (all three groups) microbiome dataset permanova for effect of group")
global_permS1 <- adonis2(phyloseq::distance(psP, method = "bray") ~ group,
        data=data.frame(META),
        by = "margin",
        permutations = 5000)

print(global_permS1)

write_xlsx(global_permS1, path = "Global_permS1.xlsx")

# ## pairwise b/w groups
# pairwise.adonis(phyloseq::distance(psP, method = "bray"), data.frame(META)$group,
#                 sim.method = "bray",
#                 p.adjust.m = "bonferroni")
### humac pairwise permanova
print("human- vs. macaque-donor microbiome permanova")
phylo_humac <- subset_samples(psP, group == "mac" | group == "hum")
pairw_humacS3 <- adonis2(phyloseq::distance(phylo_humac, method = "bray") ~ group,
        data=data.frame(META) %>% filter(group != "sqm"),
        by = "margin",
        permutations = 5000)

print(pairw_humacS3)

write_xlsx(pairw_humacS3, path = "Pairw_humacS3.xlsx")

### sqmac pairwise permanova
print("squirrel monkey- vs. macaque-donor microbiome permanova")
phylo_sqmac <- subset_samples(psP, group == "mac" | group == "sqm")
pairw_sqmacs2 <- adonis2(phyloseq::distance(phylo_sqmac, method = "bray") ~ group,
        data=data.frame(META) %>% filter(group != "hum"),
        by = "margin",
        permutations = 5000)

print(pairw_sqmacs2)

write_xlsx(pairw_sqmacs2, path = "Pairw_sqmacs2.xlsx")

```







```{r, corncob in pairwise for paper 1}


set.seed(100)
########### comparing mac to hum, pairwise; the corncob model is set up to test differential abundance (formula) and variability (phi.formula) while controlling for the effect of group on dispersion (phi.formula null) 
## in the manner it's currently written
##https://cran.r-project.org/web/packages/corncob/vignettes/corncob-intro.html
sample_data(ps)$group <- relevel(factor(sample_data(ps)$group), ref = "mac")
ps_humac <- subset_samples(ps, group %in% c("mac", "hum"))
otu_table(ps_humac) <- (otu_table(ps_humac))

# beta diverisity dispersion is  significant/"overdispersed", so we change phi.formula_null = ~ group rather than ~ 1
permutest(vegan::betadisper(phyloseq::distance(ps_humac, method = "bray"),
                  phyloseq::sample_data(ps_humac)$group))

corn_da_humac <- differentialTest(formula = ~ group,
                            phi.formula = ~ group,
                            formula_null = ~ 1,
                            phi.formula_null = ~ group,
                            data = ps_humac,
                            test = "Wald", boot = FALSE,
                            fdr_cutoff = 0.001)

#write out to rnaseq uncomment next line

#write_rds(corn_da_humac,"C:/Users/willh/OneDrive - Loyola University Chicago/Documents/Mouse_data_AnthonyP2025/LiverRNAseqPackage_19082025/LiverRNAseqPackage_19082025/corn_da_humac" )

fdr_corncob_humac <- corn_da_humac$significant_taxa
dim(data.frame(fdr_corncob_humac))
## pick top 33 of macaque corncob sig hits
hits_humac <- sort(corn_da_humac$p_fdr < 0.001)
hits_humac <- data.frame(hits_humac) %>% rownames_to_column() %>% filter(hits_humac == TRUE)
humactaxspec <- str_extract(hits_humac$rowname, "s__.*")

humactax <- data.frame(sort(corn_da_humac$p_fdr)) %>% filter(., `sort.corn_da_humac.p_fdr.` < 0.001)
humactax$tax <- rownames(humactax)
humactaxtop <- data.frame(humactax[1:870,])


########### comparing mac to sqm, pairwise
sample_data(ps)$group <- relevel(factor(sample_data(ps)$group), ref = "mac")
ps_sqmac <- subset_samples(ps, group %in% c("mac", "sqm"))
otu_table(ps_sqmac) <- (otu_table(ps_sqmac))

# beta diverisity dispersion is significant/"overdispersed", so we keep phi.formula_null = ~ group, rather than modifying to ~ 1
permutest(vegan::betadisper(phyloseq::distance(ps_sqmac, method = "bray"),
                  phyloseq::sample_data(ps_sqmac)$group))

corn_da_sqmac <- differentialTest(formula = ~ group,
                            phi.formula = ~ group,
                            formula_null = ~ 1,
                            phi.formula_null = ~ group,
                            data = ps_sqmac,
                            test = "Wald", boot = FALSE,
                            fdr_cutoff = 0.001)
fdr_corncob_sqmac <- corn_da_sqmac$significant_taxa
dim(data.frame(fdr_corncob_sqmac))
hits_sqmac <- sort(corn_da_sqmac$p_fdr < 0.001)
hits_sqmac <- data.frame(hits_sqmac) %>% rownames_to_column() %>% filter(hits_sqmac == TRUE)
sqmactaxspec <- str_extract(hits_sqmac$rowname, "s__.*")
#write out to rna seq
write_rds(corn_da_sqmac,"C:/Users/willh/OneDrive - Loyola University Chicago/Documents/Mouse_data_AnthonyP2025/LiverRNAseqPackage_19082025/LiverRNAseqPackage_19082025/corn_da_sqmac" )


sqmactax <- data.frame(sort(corn_da_sqmac$p_fdr)) %>% filter(., `sort.corn_da_sqmac.p_fdr.` < 0.001)
sqmactax$tax <- rownames(sqmactax)
sqmactaxtop <- data.frame(sqmactax[1:870,])

humactab <- data.frame(table(sub(".*s__(.*)\\|.*","\\1",humactax$tax)))
humactab$group <- "humac"
sqmactab <- data.frame(table(sub(".*s__(.*)\\|.*","\\1",sqmactax$tax)))
sqmactab$group <- "sqmac"

totpack <- data.frame(rbind(humactab, sqmactab))

definers <- totpack[!(duplicated(totpack$Var1)|duplicated(totpack$Var1, fromLast=TRUE)),, drop=FALSE]
## gives you only mutually, shared taxa from pwise comparisons from each corncob model 
humacfin <- humactax %>% filter(tax %in% 
                                  (definers %>% filter(group == "humac")%>% pull(.,"Var1")))
sqmacfin <- sqmactax %>% filter(tax %in% 
                                  (definers %>% filter(group == "sqmac")%>% pull(.,"Var1"))) 


```

## corncob model plotting

```{r, plotting on corncob for paper 1}

####### now we need to focus on publication quality figures
####### this means filtering each corncob res object by mutual taxa and plotting top 10 or so side by side...
glance_significant_models <- function(corncob_results){
  df <- data.frame()
  for(i in 1:length(corncob_results$significant_taxa)){
    taxa_name <- corncob_results$significant_taxa[i]
    taxa_df <- corncob_results$significant_models[[i]]$coefficients %>%
      as.data.frame() %>%
      rownames_to_column("covariate") %>%
      mutate(taxa_name = taxa_name)
    df <- bind_rows(df, taxa_df)
  }
  df <- janitor::clean_names(df) %>%
    select(covariate, estimate, std_error, t_value, fdr = pr_t, taxa_name)
  return(df)
}

mutual_taxa_total <- 
  intersect(
    glance_significant_models(corn_da_sqmac) %>%
      #filter(t_value < -3 | t_value > 3) %>%
      filter(fdr < 0.05 & std_error < 0.3) %>%
      #filter(!grepl("Intercept", covariate)) %>%
      pull(taxa_name),
    
    glance_significant_models(corn_da_humac) %>%
      #filter(t_value < -3 | t_value > 3) %>%
      filter(fdr < 0.05 & std_error < 0.3) %>%
      #filter(!grepl("Intercept", covariate))%>%
      pull(taxa_name)
) 

subset_corncob_result_full <- function(result, taxa_to_keep, subset_data = TRUE) {
  taxa_to_keep <- intersect(names(result$p), taxa_to_keep)

  result$p <- result$p[taxa_to_keep]
  result$p_fdr <- result$p_fdr[taxa_to_keep]

  #taking care of the subset on significant_models slot
  #and this is actually enough for plotting going fwd
  #because i think corncob plot fxn natively uses sig taxa
  names(result$significant_models) <- result[["significant_taxa"]]
   result[["significant_taxa"]] <- names(result$p_fdr)
   if (!is.null(result$significant_models)) {
     result$significant_models <- result$significant_models[
       intersect(names(result$significant_models), taxa_to_keep)
     ]
   }
  return(result)
}

# genct <- glance_significant_models(corn_da_sqmac) %>%
#       #filter(t_value < -5 | t_value > 5) %>% 
#       #filter(fdr < 0.05 & std_error < 0.3) %>%
#       filter(!grepl("Intercept", covariate)) 

# these are the top 5 genera
# sub(".*g__(.*)\\|.*","\\1",mutual_taxa) %>% 
#   table() %>%
#   sort() %>% tail(5)

#Of the mutual taxa Clostridia and Bacteroidia by far out-number the rest, Bacilli, Gammaproteobacteria dbl digits
#table(sub(".*c__(.*?)\\|.*", "\\1", mutual_taxa_total))
topclass <- c("Alphaproteobacteria", "Bacilli", "Bacteroidia", "Clostridia", "Gammaproteobacteria")
mutual_class <- sub(".*c__(.*?)\\|.*", "\\1", mutual_taxa_total)
### at order level we have raw count dominance of Bacteroidales, Lachnospirales, and Eubacteriales; the rest don't make double digits
#table(sub(".*o__(.*?)\\|.*", "\\1", mutual_taxa_total))
toporder <- c("Bacteroidales", "Eubacteriales", "Lachnospirales")
mutual_order <- sub(".*f__(.*?)\\|.*", "\\1", mutual_taxa_total)
### family level: Lachnospiraceae(20 ct.);Clostridiaceae(18 ct.);Prevotellaceae(15 ct.);Bacteroidaceae(14 ct.);Oscillospiraceae (10 ct.)
data.frame(table(sub(".*f__(.*?)\\|.*", "\\1", mutual_taxa_total))) %>% filter(Freq > 10)
topfamily <- c("Bacteroidaceae","Clostridiaceae","Lachnospiraceae","Oscillospiraceae","Flavobacteriaceae","Prevotellaceae","Sphingobacteriaceae")
mutual_family <- sub(".*f__(.*?)\\|.*", "\\1", mutual_taxa_total)
### genera: Bacteroides Clostridium Prevotella
#table(sub(".*g__(.*)\\|.*","\\1",mutual_taxa_total))
#topgenera <- c("Bacteroides", "Clostridium", "Prevotella")
topgenera <- c("Bacteroides", "Clostridium", "Prevotella")
mutual_genera <- sub(".*g__(.*)\\|.*","\\1",mutual_taxa_total)

    
 # tiff("corncob.family.1v1.tiff", units="in", width=12, height=10, res=300)
  
for (taxlvl in unique(topfamily)) 
  
  taxa_to_plot <- grep(taxlvl, mutual_taxa_total, value = TRUE)
  
  sqmac_subset <- subset_corncob_result_full(corn_da_sqmac, taxa_to_plot)
  humac_subset <- subset_corncob_result_full(corn_da_humac, taxa_to_plot)

  species_labels <- stringr::str_extract(taxa_to_plot, "s__[^|]+$")
  #names(species_labels) <- taxa_to_plot

  
  
#to export corncob plots as panels follow below  
  
library(ggplot2)
library(patchwork)
library(gridExtra)
library(dplyr)
library(tidyr)
library(stringr)
library(tibble)

# ... your glance_significant_models(), subset_corncob_result_full(), and data prep above ...

plot_list <- list()

# drop "Flavobacteriaceae" and other families from the loop set
for (taxlvl in setdiff(unique(topfamily), c("Flavobacteriaceae", "Oscillospiraceae", "Clostridiaceae", "Prevotellaceae","Sphingobacteriaceae"))) {

  taxa_to_plot <- grep(taxlvl, mutual_taxa_total, value = TRUE)

  sqmac_subset <- subset_corncob_result_full(corn_da_sqmac, taxa_to_plot)
  humac_subset <- subset_corncob_result_full(corn_da_humac, taxa_to_plot)

  species_labels <- stringr::str_extract(taxa_to_plot, "s__[^|]+$") %>%
  stringr::str_replace("^s__", "")  %>%
  stringr::str_replace_all("_", " ")


  fg1 <- plot(sqmac_subset) +
    scale_y_discrete(labels = species_labels) +
    scale_x_continuous(limits = c(-6, 6), breaks = seq(-6, 6, by = 2)) +
      theme(
      strip.background = element_blank(),
      strip.text = element_blank(),
      axis.text.y = element_text(size = 16),
      axis.text.x = element_text(size = 16),
      plot.title = element_text(size = 26)
    ) +
    guides(color = "none") +
    ggtitle(paste0("Family: ", taxlvl), subtitle = "Mac v Sqm") +
    theme( 
      plot.margin = margin(t = 0, r = 0, b = 0, l = 30),
      plot.title = element_text(face = "bold", size = 28, hjust = .5),
      axis.title.y = element_blank(),
      axis.text.y = element_text(colour = "black", face = "bold"),
      axis.line = element_line(size = 3, colour = "black"),
      axis.title.x = element_text(face = "bold"),
      plot.subtitle = element_text(face = "bold", size = 24, color = "black"),
      axis.ticks = element_line(color = "black"),
      axis.text.x.bottom = element_text(face = "bold", color = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 5)
    )

  fg2 <- plot(humac_subset) +
    scale_y_discrete(labels = species_labels) +
    scale_x_continuous(limits = c(-6, 6), breaks = seq(-6, 6, by = 2)) +
      theme(
      strip.background = element_blank(),
      strip.text = element_blank(),
      axis.text.y = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(size = 14),
      plot.title = element_text(size = 22),
      axis.ticks.y = element_blank()) +
    guides(color = "none") +
    ggtitle("", subtitle = "Mac v Hum") +
    theme(
      plot.title = element_text(face = "bold", size = 28, hjust = .5),
      axis.title.y = element_blank(),
      axis.line = element_line(size = 3, colour = "black"),
      axis.title.x = element_text(face = "bold"),
      plot.subtitle = element_text(face = "bold", size = 24, color = "black"),
      axis.ticks = element_line(color = "black"),
      axis.text.x.bottom = element_text(color = "black", face = "bold"),
      panel.border = element_rect(colour = "black", fill = NA, size = 5)
    ) +
    ylab(NULL)

  combined_plot <- fg1 + fg2 + plot_layout(ncol = 2, guides = "collect")

  # store the finished patchwork
  plot_list[[length(plot_list) + 1]] <- combined_plot
}

# build the grid
ncol_grid <- 2
nrow_grid <- ceiling(length(plot_list) / ncol_grid)

final_grid <- wrap_plots(plot_list, ncol = ncol_grid) 

# save all at once
ggsave(
  filename = "corncob_families_excluding.seperated.tiff",
  plot = final_grid,
  width = 10 * ncol_grid,
  height = 7 * nrow_grid,
  dpi = 300
)

#effort to put all plots into a grid




```




## corncob Suppl model plotting

```{r, plotting on corncob for paper 1}

####### now we need to focus on publication quality figures for supplemental

plot_list2 <- list()

target_families <- c("Flavobacteriaceae", "Clostridiaceae", "Prevotellaceae", "Oscillospiraceae")

for (taxlvl2 in target_families) {
  
    taxa_to_plot <- grep(taxlvl2, mutual_taxa_total, value = TRUE)
  sqmac_subset <- subset_corncob_result_full(corn_da_sqmac, taxa_to_plot)
  humac_subset <- subset_corncob_result_full(corn_da_humac, taxa_to_plot)

  species_labels <- stringr::str_extract(taxa_to_plot, "s__[^|]+$") %>%
    stringr::str_replace("^s__", "") %>%
    stringr::str_replace_all("_", " ")

     fg3 <- plot(sqmac_subset) +
    scale_y_discrete(labels = species_labels) +
    scale_x_continuous(limits = c(-6, 6), breaks = seq(-6, 6, by = 2)) +
      theme(
      strip.background = element_blank(),
      strip.text = element_blank(),
      axis.text.y = element_text(size = 14.5),
      axis.text.x = element_text(size = 16),
      plot.title = element_text(size = 26)
    ) +
    guides(color = "none") +
    ggtitle(paste0("Family: ", taxlvl2), subtitle = "Mac v Sqm") +
    theme( 
      plot.margin = margin(t = 0, r = 0, b = 0, l = 30),
      plot.title = element_text(face = "bold", size = 28, hjust = .5),
      axis.title.y = element_blank(),
      axis.text.y = element_text(colour = "black", face = "bold"),
      axis.line = element_line(size = 3, colour = "black"),
      axis.title.x = element_text(face = "bold"),
      plot.subtitle = element_text(face = "bold", size = 24, color = "black"),
      axis.ticks = element_line(color = "black"),
      axis.text.x.bottom = element_text(face = "bold", color = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 5)
    )

  fg4 <- plot(humac_subset) +
    scale_y_discrete(labels = species_labels) +
    scale_x_continuous(limits = c(-6, 6), breaks = seq(-6, 6, by = 2)) +
      theme(
      strip.background = element_blank(),
      strip.text = element_blank(),
      axis.text.y = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(size = 14),
      plot.title = element_text(size = 22),
      axis.ticks.y = element_blank()) +
    guides(color = "none") +
    ggtitle("", subtitle = "Mac v Hum") +
    theme(
      plot.title = element_text(face = "bold", size = 28, hjust = .5),
      axis.title.y = element_blank(),
      axis.line = element_line(size = 3, colour = "black"),
      axis.title.x = element_text(face = "bold"),
      plot.subtitle = element_text(face = "bold", size = 24, color = "black"),
      axis.ticks = element_line(color = "black"),
      axis.text.x.bottom = element_text(color = "black", face = "bold"),
      panel.border = element_rect(colour = "black", fill = NA, size = 5)
    ) +
    ylab(NULL)

  combined_plot2 <- fg3 + fg4 + plot_layout(ncol = 2, guides = "collect")
  plot_list2[[length(plot_list2) + 1]] <- combined_plot2
}


# build the grid
ncol_grid <- 2
nrow_grid <- ceiling(length(plot_list2) / ncol_grid)


final_grid2 <- wrap_plots(plot_list2, ncol = 2)
ggsave("corncob_selected_supp_families.tiff", plot = final_grid2,
       width = 10 * 2, height = 8 * ceiling(length(plot_list2) / 2), dpi = 300)


```



## microbiome co-abundance network -- corncob microbes
human v macaque

```{r }

set.seed(123)
mutual_taxa_total <-
  intersect(
    glance_significant_models(corn_da_sqmac) %>%
      filter(t_value < -3 | t_value > 3) %>%
      filter(fdr < 0.1 & std_error < 0.3) %>%
      filter(!grepl("Intercept", covariate)) %>%
      pull(taxa_name),

    glance_significant_models(corn_da_humac) %>%
      filter(t_value < -3 | t_value > 3) %>%
      filter(fdr < 0.1 & std_error < 0.3) %>%
      filter(!grepl("Intercept", covariate))%>%
      pull(taxa_name)
)
glance_significant_models(corn_da_sqmac) %>%
      filter(t_value < -3 | t_value > 3) %>%
      filter(fdr < 0.1 & std_error < 0.3) %>%
      filter(!grepl("Intercept", covariate)) %>%
      filter(taxa_name %in% mutual_taxa_total) 
coabund_phylo <- subset_samples(psP, group %in% c("hum", "mac"))
coabund_phylo <- prune_taxa(mutual_taxa_total, coabund_phylo)

abundance_matrix <- otu_table(coabund_phylo)
abundance_matrix <- as.matrix(abundance_matrix)
abundance_matrix <- t(abundance_matrix)  #taxa as columns

cor_mat <- cor(abundance_matrix, method = "spearman")
colnames(cor_mat) <- str_extract(colnames(cor_mat), "s__.*")
rownames(cor_mat) <- str_extract(rownames(cor_mat), "s__.*")

cor_mat[lower.tri(cor_mat)] <- NA  #upper triangle only
cor_mat[lower.tri(cor_mat)] <- NA  #upper triangle only
cor_df <- reshape2::melt(cor_mat, na.rm = TRUE)
names(cor_df) <- c("Gene1", "Gene2", "correlation")
##strong thresh for network
edges <- cor_df %>% filter(abs(correlation) > 0.70)
g <- graph_from_data_frame(edges, directed = FALSE)
V(g)$degree <- igraph::degree(g)

### detect louvain clustered modules; switching to infomap --
## more robust for a smaller network; less coarse network results giving us more specific focused modularity
### more robust than WGCNA which focuses on defining groups by co-expression (and also requires bit larger sample size/#input genes)
comm <- cluster_louvain(g)

V(g)$module <- factor(membership(comm))
modules <- levels(V(g)$module)

module_colors <- brewer.pal(length(modules), "Set3")
names(module_colors) <- modules

# Prepare node info
node_info <- tibble(
  name = V(g)$name,
  module = V(g)$module,
  degree = V(g)$degree
)

# Identify hubs per module
top_hubs <- node_info %>%
  group_by(module) %>%
  filter(degree == max(degree)) %>%
  pull(name)

# Assign shapes: triangles for hubs, circles for others
node_info <- node_info %>%
  mutate(
    shape_group = ifelse(name %in% top_hubs, 24, 21),  # 24=triangle, 21=circle
    outline_color = ifelse(degree >= 60, "red", "black")  # red outline for degree ≥50
  )

# Match order
node_info <- node_info[match(V(g)$name, node_info$name), ]

# Assign to graph vertices
V(g)$module <- factor(node_info$module, levels = modules)
V(g)$shape_group <- node_info$shape_group
V(g)$degree <- node_info$degree
V(g)$name <- node_info$name
V(g)$outline_color <- node_info$outline_color

# Plot
library(grid)  # for custom legend key grobs
ggraph(g, layout = "stress") +
  geom_edge_link(aes(edge_alpha = abs(correlation)), color = "grey70", show.legend = FALSE) +
geom_node_point(aes(size = degree, fill = module, shape = factor(shape_group), color = outline_color),
                stroke = 1.5,
                show.legend = c(fill = TRUE, shape = FALSE, size = FALSE, color = FALSE)) +
  geom_node_text(aes(label = name), repel = TRUE, size = 3.5) +
  scale_fill_manual(
    values = module_colors,
    name = "Module",
    guide = guide_legend(
      override.aes = list(shape = 21)  # legend keys are circles, colored by module
    )
  ) +
  scale_shape_manual(
    values = c(`21` = 21, `24` = 24),
    guide = "none"  # hide shape legend
  ) +
  scale_color_identity() +  # use outline colors literally, no legend
  theme_void() +
  labs(title = "Human-Macaque Donor Gene Correlation Network:",
       subtitle = "Louvain module, Hub Triangles and High-Degree Centrality (>=30) Highlighted nodes (red)")







```


squirrel monkey v macaque
```{r }

set.seed(123)

mutual_taxa_total <-
  intersect(
    glance_significant_models(corn_da_sqmac) %>%
      filter(t_value < -3 | t_value > 3) %>%
      filter(fdr < 0.1 & std_error < 0.3) %>%
      filter(!grepl("Intercept", covariate)) %>%
      pull(taxa_name),

    glance_significant_models(corn_da_humac) %>%
      filter(t_value < -3 | t_value > 3) %>%
      filter(fdr < 0.1 & std_error < 0.3) %>%
      filter(!grepl("Intercept", covariate))%>%
      pull(taxa_name)
)
coabund_phylo <- subset_samples(psP, group %in% c("sqm", "mac"))
coabund_phylo <- prune_taxa(mutual_taxa_total, coabund_phylo)

abundance_matrix <- otu_table(coabund_phylo)
abundance_matrix <- as.matrix(abundance_matrix)
abundance_matrix <- t(abundance_matrix)  # taxa as columns

cor_mat <- cor(abundance_matrix, method = "spearman")
colnames(cor_mat) <- str_extract(colnames(cor_mat), "s__.*")
rownames(cor_mat) <- str_extract(rownames(cor_mat), "s__.*")

cor_mat[lower.tri(cor_mat)] <- NA  # upper triangle only
cor_mat[lower.tri(cor_mat)] <- NA  # upper triangle only
cor_df <- reshape2::melt(cor_mat, na.rm = TRUE)
names(cor_df) <- c("Gene1", "Gene2", "correlation")
##strong thresh for network
edges <- cor_df %>% filter(abs(correlation) > 0.70)
g <- graph_from_data_frame(edges, directed = FALSE)
V(g)$degree <- igraph::degree(g)

### detect louvain clustered modules; switching to infomap --
## more robust for a smaller network; less coarse network results giving us more specific focused modularity
### more robust than WGCNA which focuses on defining groups by co-expression (and also requires bit larger sample size/#input genes)
comm <- cluster_louvain(g)

V(g)$module <- factor(membership(comm))
modules <- levels(V(g)$module)

module_colors <- brewer.pal(length(modules), "Set3")
names(module_colors) <- modules

# Prepare node info
node_info <- tibble(
  name = V(g)$name,
  module = V(g)$module,
  degree = V(g)$degree
)

# Identify hubs per module
top_hubs <- node_info %>%
  group_by(module) %>%
  filter(degree == max(degree)) %>%
  pull(name)

# Assign shapes: triangles for hubs, circles for others
node_info <- node_info %>%
  mutate(
    shape_group = ifelse(name %in% top_hubs, 24, 21),  # 24=triangle, 21=circle
    outline_color = ifelse(degree >= 60, "red", "black")  # red outline for degree ≥50
  )

# Match order
node_info <- node_info[match(V(g)$name, node_info$name), ]

# Assign to graph vertices
V(g)$module <- factor(node_info$module, levels = modules)
V(g)$shape_group <- node_info$shape_group
V(g)$degree <- node_info$degree
V(g)$name <- node_info$name
V(g)$outline_color <- node_info$outline_color

# Plot
library(grid)  # for custom legend key grobs
ggraph(g, layout = "stress") +
  geom_edge_link(aes(edge_alpha = abs(correlation)), color = "grey70", show.legend = FALSE) +
geom_node_point(aes(size = degree, fill = module, shape = factor(shape_group), color = outline_color),
                stroke = 1.5,
                show.legend = c(fill = TRUE, shape = FALSE, size = FALSE, color = FALSE)) +
  geom_node_text(aes(label = name), repel = TRUE, size = 3.5) +
  scale_fill_manual(
    values = module_colors,
    name = "Module",
    guide = guide_legend(
      override.aes = list(shape = 21)  # legend keys are circles, colored by module
    )
  ) +
  scale_shape_manual(
    values = c(`21` = 21, `24` = 24),
    guide = "none"  # hide shape legend
  ) +
  scale_color_identity() +  # use outline colors literally, no legend
  theme_void() +
  labs(title = "Squirrel Monkey-Macaque Donor Gene Correlation Network:",
       subtitle = "Louvain module, Hub Triangles and High-Degree Centrality (>=30) Highlighted nodes (red)")


abund_df <- ps_melt(coabund_phylo) %>%
  group_by(OTU, group) %>%
  summarize(mean_abund = mean(Abundance), .groups = "drop") %>%
  mutate(log2_abund = log2(mean_abund + 1e-6))  # avoid log(0)
abund_df$OTU <- str_extract(abund_df$OTU, "s__.*")

abund_wide <- abund_df %>%
  pivot_wider(names_from = group, values_from = log2_abund, values_fill = 0)

V(g)$log2abund_sqm <- abund_wide$sqm[match(V(g)$name, abund_wide$OTU)]
V(g)$log2abund_mac <- abund_wide$mac[match(V(g)$name, abund_wide$OTU)]

abund_wide <- abund_wide %>%
  mutate(mac_minus_sqm = mac - sqm)

V(g)$log2fc_mac_sqm <- abund_wide$mac_minus_sqm[match(V(g)$name, abund_wide$OTU)]

ggraph(g, layout = "stress") +
  geom_edge_link(alpha = 0.2) +
  geom_node_point(aes(color = log2fc_mac_sqm), size = 4) +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  scale_color_gradient2(
    low = "blue", mid = "white", high = "red", midpoint = 0,
    name = "log2(Abundance)\nMacaque - Sqm"
  ) + 
  theme_graph()



```




## taxonomic BOXplots
Not sure it's worth using these given the taxonomic barplots, but can slap error bars and p-val thresholds on if it makes more sense to use them in the end...

```{r, contains boxplots for a few taxa of interest (slides 25-37) }

############### MIGHT END UP BEING SUPPL. BOX PLOTS FOR CHAPTER/PAPER 1;
############### THESE ARE ALREADY ON SLIDE DECK
library(ggplot2)
library(ggpubr)

#add theme 

theme_bold <- theme(
  axis.title.x = element_text(size = 22, face = "bold"),
  axis.title.y = element_text(size = 22,face = "bold"),
  axis.text.x = element_text(face = "bold", size = 22),
  axis.text.y = element_text(size = 22,face = "bold"),
  legend.title = element_text(size = 18,face = "bold"),
  legend.text = element_text(size = 22,face = "bold"),
  plot.title = element_text(size = 18,face = "bold"),
  strip.text = element_text(size = 22,face = "bold"),
  axis.line = element_line(size = 3, color = "black"),
)

# given data integration results (network projection)
pdata <- phyloseq::psmelt(psP)
my_comparisons <- list( c("mac", "sqm"), c("hum", "mac"), c("sqm", "hum") )

#####Clostridia (Class-lvl)######
 ggboxplot(filter(pdata, Class == "c__Clostridia"),
          x = "group", y = "Abundance",
          color = "group", order = c("hum", "mac", "sqm"))+ 
  scale_y_continuous(trans="log10")+
  stat_compare_means(aes(group,Freq),comparisons = my_comparisons, method = "wilcox.test") + 
  ggtitle("Clostridia") +
  scale_color_manual(values = colorRampPalette(brewer.pal(8, "Dark2"))(3)) +
  theme(legend.position="none") + ylab("Relative Abundance (log10)") + xlab("Donor Group")
#####Lachnospirales (Order-lvl)######
 ggboxplot(filter(pdata, Order == "o__Lachnospirales"),
          x = "group", y = "Abundance",
          color = "group", order = c("hum", "mac", "sqm"))+ 
  scale_y_continuous(trans="log10")+
  stat_compare_means(aes(group,Freq),comparisons = my_comparisons, method = "wilcox.test") + 
  ggtitle("Lachnospirales") +
  scale_color_manual(values = colorRampPalette(brewer.pal(8, "Dark2"))(3)) +
  theme(legend.position="none") + ylab("Relative Abundance (log10)") + xlab("Donor Group")
#####Lachnospiraceae (Family-lvl)######
 ggboxplot(filter(pdata, Family == "f__Lachnospiraceae"),
          x = "group", y = "Abundance",
          color = "group", order = c("hum", "mac", "sqm"))+ 
  scale_y_continuous(trans="log10")+
  stat_compare_means(aes(group,Freq),comparisons = my_comparisons, method = "wilcox.test") + 
  ggtitle("Lachnospiraceae") +
  scale_color_manual(values = colorRampPalette(brewer.pal(8, "Dark2"))(3)) +
  theme(legend.position="none") + ylab("log10(Rel. Abundance)") + xlab("Donor Group")






#####Genus Level######
genvec <- c("g__Lachnoclostridium","g__Enterocloster")
for( x in genvec){
  ggboxplot(filter(pdata, Genus == x),
          x = "group", y = "Abundance",
          color = "group", order = c("hum", "mac", "sqm"))+ 
  scale_y_continuous(trans="log10")+
  stat_compare_means(aes(group,Freq),comparisons = my_comparisons, method = "wilcox.test") + 
  ggtitle(paste0("", x)) +
  scale_color_manual(values = colorRampPalette(brewer.pal(8, "Dark2"))(3)) +
  theme(legend.position="none") + ylab("log10(Rel. Abundance)") + xlab("Donor Group")
}


#####Species Level######
specvec <- c("s__[Clostridium]_hylemonae","s__[Clostridium]_scindens","s__[Clostridium]_symbiosum",
              "s__[Clostridium]_innocuum","s__Clostridioides_difficile","s__Butyrivibrio_crossotus",
             "s__Lachnospira_eligens","s__Simiaoa_sunii","s__Agathobacter_rectalis","s__Enterocloster_bolteae","s__Clostridium_perfringens", "s__Peptacetobacter_hiranonis" )

for (x in specvec) {
  # Clean species name for display and file naming
  clean_label <- x %>%
    stringr::str_replace("^s__", "") %>%
    stringr::str_replace_all("_", " ")

  # Create plot
  p <- ggboxplot(filter(pdata, Species == x),
          x = "group", y = "Abundance", size = 2,
          color = "group", order = c("hum", "mac", "sqm")) +
    scale_y_continuous(trans = "log10") +
    stat_compare_means(aes(group, Freq), comparisons = my_comparisons, method = "wilcox.test", size = 4.3, bracket.size = 2, font=list(face = "bold", step.increase = .1)) +
    ggtitle(paste0(" ", clean_label)) +
    scale_color_manual(values = colorRampPalette(brewer.pal(8, "Dark2"))(3)) +
    theme(legend.position = "none") + 
    theme_bold + theme(
    aspect.ratio = 1/2,
    legend.position = "none",
    axis.ticks.x = element_line(size = 3, color = "black")) + 
    theme_bold  +
  theme(
    aspect.ratio = 1/2,
    legend.position = "none",
    axis.text.x = element_text(face = "bold", size = 22),
    axis.title.x = element_text(size = 22, face = "bold"),
    axis.ticks.x = element_line(size = 3, color = "black")) +
    ylab("log10(Rel. Abundance)") +
    xlab("Donor Group")

  print(p)

  # Save with cleaned filename
  ggsave(filename = paste0("boxplot_S1C_", clean_label, ".tiff"), plot = p,
         width = 8, height = 4.5, dpi = 300)
}
#################### Bacteroides heparinolyticus
ggboxplot(filter(pdata, Species == "s__Clostridioides_difficile"),
          x = "group", y = "Abundance",
          color = "group", order = c("hum", "mac", "sqm"))+ 
  scale_y_continuous(trans="log10")+
  stat_compare_means(aes(group,Freq),comparisons = my_comparisons, method = "wilcox.test") + 
  ggtitle("Bacteroides_heparinolyticus,
          a **famously** bile-sensitive taxon") +
  scale_color_manual(values = colorRampPalette(brewer.pal(8, "Dark2"))(3)) +
  theme(legend.position="none") + ylab("log10(Rel. Abundance)") + xlab("Donor Group")

############### Some extras... don't necessarily need them, but could be useful in context. we'll see the writeup

# filter(pdata, Species == "s__Phocaeicola_vulgatus") %>%
# ggplot(data = ., aes(x = group, y = Abundance)) +
#   geom_boxplot(outlier.shape  = NA) +
#   geom_jitter(aes(color = group), height = 0, width = .2) +
#   ggtitle("Bacteroides boxplots")

# filter(pdata, Genus == "g__Bacteroides") %>% filter(group == "mac"| group == "hum") %>%
# ggplot(data = ., aes(x = group, y = Abundance)) +
#   geom_boxplot(outlier.shape  = NA) +
#   geom_jitter(aes(color = group), height = 0, width = .2) +
#   ggtitle("Lactobacillus plots")
# 
# filter(pdata, Genus == "g__Bifidobacterium") %>%
# ggplot(data = ., aes(x = group, y = log(Abundance))) +
#   geom_boxplot(outlier.shape  = NA) +
#   geom_jitter(aes(color = group), height = 0, width = .2) +
#   ggtitle("Bifido boxplots")
# 
# filter(pdata, Species == "s__Parabacteroides_goldsteinii") %>%
# ggplot(data = ., aes(x = group, y = log(Abundance))) +
#   geom_boxplot(outlier.shape  = NA) +
#   geom_jitter(aes(color = group), height = 0, width = .2) +
#   ggtitle("Bifido boxplots")


```




