---
title: "liver RNAseq look"
output: html_document
date: "2022-11-21"
---



```{r}

library(readr)
library(stringr)
#BiocManager::install("DESeq2")
#BiocManager::install("SummarizedExperiment")
library(SummarizedExperiment)
library(DESeq2)
library(tibble)
library(tidyr)
#BiocManager::install("clusterProfiler")
library(clusterProfiler)
library(ggplot2)
library(RColorBrewer)
library(pheatmap)
library(glmpca)
library(reshape2)
library(tibble)
 #if(!require(devtools)) install.packages("devtools")
 #devtools::install_github("kassambara/ggpubr")
library(ggpubr)
library(enrichplot)
#library(AnnotationDbi)
#BiocManager::install("org.Mm.eg.db")
library(org.Mm.eg.db)
library(dplyr)
# this next i think
#BiocManager::install("DOSE")
#BiocManager::install("pathview")
#if (!require("BiocManager", quietly = TRUE))
 #   install.packages("BiocManager")
#
#BiocManager::install("clusterProfiler")
#BiocManager::install("org.Mm.eg.db")
library(DOSE)
library(pathview)
#BiocManager::install("biomaRt")
library(biomaRt)
library(rbioapi)
library(purrr)
library(AnnotationDbi)
library(ggrepel)
library(phyloseq)
library(gridExtra)

```



## corncob microbe-liver gene correlation

```{r}

############## generate counts
filenames <- list.files(full.names=TRUE)
filenames <- grep("readcounts.txt$", filenames, value = TRUE)
All_files <- lapply(filenames, function(i)
  read_tsv(i)
  )
counts <- All_files %>% 
  purrr::reduce(left_join, by = "geneID") %>% tibble::column_to_rownames(var = "geneID")
counts <- counts[!grepl("^_", rownames(counts)), ]

metadata_rna <- read.csv("Liver_Gene_expression_metadata.csv")
#metadata <- metadata[-14,]
metadata_rna[26,3] <- "Squirrel_monkey"
metadata_humac <- metadata_rna %>% filter(Species !="Squirrel_monkey") #metadata_rna[-14,] %>% filter(Species !="Squirrel_monkey")
metadata_sqmac <- metadata_rna %>% filter(Species !="Human")#metadata_rna[-14,] %>% filter(Species !="Human")
metadata_sqmac <- metadata_sqmac[-11,]
metadata_rna$fat <- as.factor(ifelse(metadata_rna$Species == "Macaque", 1, 0))
#metadata_rna$fat <- as.factor(metadata_rna$fat)
#metadata_rna$sampleid <- as.factor(metadata_rna$sampleid)
metadata_rna$renamed_for_phys <- NULL

metadata_rna$sampleid
metadata_rna$sample_num <- c("sqm12", "sqm14","sqm16","sqm18", "mac8","mac7", "sqm13", "sqm9", "sqm11","sqm10", "sqm15", "mac4", "mac2","mac6","mac3","mac1","mac5","hum22","hum20","hum21","hum26","hum23","hum25","hum24","hum19", "sqm17")

#counts <- counts[,-14]
counts_humac <- counts[,paste(metadata_humac$sampleid, sep = "")]
counts_humac <- counts_humac[!grepl("^_", rownames(counts_humac)), ]
counts_sqmac <- counts[,paste(metadata_sqmac$sampleid, sep = "")]
counts_sqmac <- counts_sqmac[!grepl("^_", rownames(counts_sqmac)), ]


dds <- DESeqDataSetFromMatrix(countData = counts,
                                 colData = metadata_rna,
                                 design = ~Species)
smallestGroupSize <- 8
keep <- rowSums(counts(dds) >= 10) >= smallestGroupSize
dds <- dds[keep,]
#dds <- estimateSizeFactors(dds)
# rlog will estimate size factors as necessary internally  
rld <- rlog(dds, blind = TRUE)
sampleDists <- dist(t(assay(rld)))



theme_bold <- theme(
  axis.title.x = element_text(size = 44, face = "bold"),
  axis.title.y = element_text(size = 44,face = "bold"),
  axis.text.x = element_text(face = "bold", size = 36, color = "black"),
  axis.text.y = element_text(size = 36,face = "bold", color = "black"),
  legend.title = element_text(size = 28,face = "bold"),
  legend.text = element_text(size = 28,face = "bold"),
  plot.title = element_text(size = 28,face = "bold", hjust = -.5),
  strip.text = element_text(size = 32,face = "bold"),
  axis.line = element_line(size = 4.2, color = "black"),
  (plot.margin = margin(t = 20, r = 20, b = 20, l = 20))
)

#set label size
p1sqm.label=12.5
p1mac.label=12.5
p1hum.label=12.5

### baby brained dim red plot
set.seed(123)
plotPCA(rld, intgroup = c("Species")) + 
    geom_point(size = 3) +
  theme_minimal() +
  geom_point(size = 6, alpha = 0.7, stat = "identity") +
  stat_ellipse(type = "t", linetype = "solid", size = 1.8) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.position = "none") + theme_bold +
  labs( color = "Donor Group\n")+ 
  annotate("text", x = -10, y = 11, label = "sqm", color = "#666666", fontface="bold", size=p1sqm.label) +
  annotate("text", x = -2.5, y = -10, label = "mac", color = "#a6761d", fontface= "bold",size=p1mac.label) +
  annotate("text", x = 10, y = 12.2, label = "hum", color = "#1b9e77",fontface="bold", size=p1hum.label)  +
  scale_color_manual(values = c(Human="#1b9e77",Macaque="#a6761d",Squirrel_monkey="#666666"))

ggsave("PCA_Bulk_Liver.tiff", width=12.5, height=8, bg="white")

gene_mat <- t(assay(rld))
###### just so much noise... no padj < 0.3
#dds <- DESeq(dds)
#res <- data.frame(results(dds))
###### for microbe-gene correlation informed PCA; big-brained dim red plot
meta_df <- as.data.frame(colData(rld))
id_map <- setNames(meta_df$sample_num, rownames(meta_df))
rownames(gene_mat) <- id_map[rownames(gene_mat)]
###### Need to run first 3 blocks of cecal_shotgun.Rmd for this to work
library(phyloseq)
psP <- read_rds("psP.Rds")
microbe_mat <- as(otu_table(psP), "matrix") 
 microbe_mat <- t(microbe_mat)
 
#if (taxa_are_rows(physeq))  {
 # microbe_mat <- t(microbe_mat)
#}
microbe_mat <- sweep(microbe_mat, 1, rowSums(microbe_mat), FUN = "/")

common_samples <- intersect(rownames(gene_mat), rownames(microbe_mat))
####final clean-up/double check for evenness (26/all samps from all groups in each mat)
gene_mat <- gene_mat[common_samples, , drop=FALSE]
microbe_mat <- microbe_mat[common_samples, , drop=FALSE]

## quickly grabbing gene_mat/microbe_mat for 4n data integration (diablo, mofa2 run)
gene_matOut <- data.frame(gene_mat)
gene_matOut$samples <- rownames(gene_matOut)
write.csv(data.frame(gene_matOut), "4nLiver.csv", row.names = TRUE)
microbe_matOut <- data.frame(microbe_mat)
microbe_matOut$samples <- rownames(microbe_matOut)
getwd()
write.csv(data.frame(microbe_matOut),"4nMicrobeMat.csv", row.names = TRUE)

##### make correlation. matrix
cor_mat <- cor(gene_mat, microbe_mat, method = "spearman") 
##### get group-wise/sample-wise info accessible for plotting
groups <- colData(rld)$Species
samples <- rownames(gene_mat)
cor_df <- as.data.frame(cor_mat) %>%
  mutate(gene = rownames(cor_mat)) %>%
  pivot_longer(-gene, names_to = "microbe", values_to = "correlation")
#### get top n variable genes for plotting
top_genes <- rownames(cor_mat)[order(apply(abs(cor_mat), 1, max), decreasing = TRUE)][1:500]
#### make the subset of the top n genes for plotting
subset_expr <- t(assay(rld)[top_genes, ])
### prepare for PCA
pca_res <- prcomp(subset_expr, scale. = TRUE)
pca_df <- as.data.frame(pca_res$x)
pca_df$group <- colData(rld)[rownames(pca_df), "Species"]
pca_var <- pca_res$sdev^2
pca_var_explained <- round(100 * pca_var / sum(pca_var), 1)
### plot
p2sqm.label=12.5
p2mac.label=12.5
p2hum.label=12.5

set.seed(123)
ggplot(pca_df, aes(PC1, PC2, color = group)) +
  geom_point(size = 3) +
  geom_point(size = 6, alpha = 0.7, stat = "identity") +
  stat_ellipse(type = "t", linetype = "solid", size = 2.4) +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    panel.background = element_blank(), 
    axis.line = element_line(colour = "black"),
    legend.position = "none"
  ) + theme_bold +
  labs(
    title = "", 
    color = "Donor Group\n",
    x = paste0("PC1 (", pca_var_explained[1], "% variance)"),
    y = paste0("PC2 (", pca_var_explained[2], "% variance)")
  ) +
  annotate("text", x = -20, y = -7, label = "sqm", color = "#666666", fontface = "bold", size = p2sqm.label) +
  annotate("text", x = 15, y = -9.3, label = "mac", color = "#a6761d", fontface = "bold", size = p2mac.label) +
  annotate("text", x = 20, y = 7, label = "hum", color = "#1b9e77", fontface = "bold", size = p2hum.label) +
  scale_color_manual(values = c(Human="#1b9e77", Macaque="#a6761d", Squirrel_monkey="#666666"))

ggsave("PCA_500.tiff", width = 12.5, height = 8, bg = "white")


```



```{r}

#################### run the above to get what is needed for 4nMicrobeSerumLiverRNA.Rmd
comps <- c("humac", "sqmac")
sample_dist_viz <- list()
sample_pca_viz <- list()
sample_glmpca_viz <- list()
##### PREP DESeq2 DATASET / SAMPLE DIST METRICS / DIFF EXP ANALYSIS RESULTS AND FILT.
for (i in comps){
  
  setClassUnion("ExpData", c("matrix", "RangedSummarizedExperiment"))
  
  meta <- eval(as.name(paste("metadata_", i, sep = "")))

  #ddetermine the control group
  if ("Squirrel_monkey" %in% meta$Species) {
    ref_level <- "Squirrel_monkey"
  } else if ("Human" %in% meta$Species) {
    ref_level <- "Human"
  } else {
    stop()
  }

  #relevel Species so  control = reference
  meta$Species <- relevel(factor(meta$Species), ref = ref_level)

  #assign   global metadata object
  assign(paste("metadata_", i, sep = ""), meta)
  
  assign(
    paste("dds_", i, sep = ""),
    DESeqDataSetFromMatrix(countData = eval(as.name(paste("counts_",i,sep=""))),
                             colData = eval(as.name(paste("metadata_",i,sep=""))),
                             design = ~Species))
  smallestGroupSize <- 8
  assign(
    paste("keep",i,sep="_"),
         rowSums(counts(eval(as.name(paste("dds_", i, sep = "")))) >= 10) >= smallestGroupSize)
  
  assign(
    paste("dds_", i, sep = ""),
    eval(as.name(paste("dds_", i, sep = "")))[eval(as.name(paste("keep_", i, sep = ""))),]
  )
  
  assign(
    paste("dds_", i, sep = ""),
    estimateSizeFactors(eval(as.name(paste("dds_", i, sep = ""))))
  )
  
  assign(
    paste("rld_", i, sep = ""),
    rlog(eval(as.name(paste("dds_", i, sep = ""))), blind = FALSE)
    )
  

  
  sample_dist_viz[[i]] <- pheatmap(as.matrix(dist(t(assay(eval(as.name(paste("rld_", i, sep = ""))))))),
                                 clustering_distance_rows = dist(t(assay(eval(as.name(paste("rld_", i, sep = "")))))),
                                 clustering_distance_cols = dist(t(assay(eval(as.name(paste("rld_", i, sep = "")))))),
                                 col = colorRampPalette( rev(brewer.pal(9, "Blues")) )(255))
  
  assign(
    paste("pcaData_", i, sep = ""),
    plotPCA(eval(as.name(paste("rld_", i, sep = ""))), intgroup = c("Species"), returnData = TRUE)
    )
  
  sample_pca_viz[[i]] <- 
    ggplot(eval(as.name(paste("pcaData_", i, sep = ""))),
           aes(x = PC1, y = PC2, color = Species, shape = Species, label = name)) +
    geom_point(size =3) +
    xlab(paste0("PC1: ", round(100 * attr(eval(as.name(paste("pcaData_", i, sep = ""))), "percentVar"))[1], "% variance")) +
    ylab(paste0("PC2: ", round(100 * attr(eval(as.name(paste("pcaData_", i, sep = ""))), "percentVar"))[2], "% variance")) +
    coord_fixed() +
    ggtitle(paste("PCA with rld", i, "data")) + 
    geom_text()
  
  sample_glmpca_viz[[i]] <- data.frame(glmpca(counts(eval(as.name(paste("dds_", i, sep = "")))), L=2)$factors)
  sample_glmpca_viz[[i]]$Species <- eval(as.name(paste("dds_", i, sep = "")))$Species
  sample_glmpca_viz[[i]] <- ggplot(sample_glmpca_viz[[i]],
                                   aes(x = dim1, y = dim2, color = Species)) +
    geom_point(size =3) +
    coord_fixed() +
    ggtitle(paste("glmpca - Generalized PCA w/ Rld", i, "data"))
    
  assign(
    paste("dds_", i, sep = ""),
    DESeq(eval(as.name(paste("dds_", i, sep = ""))))
    )
  
    assign(
    paste("res_", i, sep = ""),
    data.frame(
      results(eval(as.name(paste("dds_", i, sep = ""))),
              alpha = 0.05, lfcThreshold=0)) #%>%
      #subset(., padj <=0.05)
    )

}

# write_rds(res_humac, "res_humac.Rds")
# write_rds(res_sqmac, "res_sqmac.Rds")

```






## microbially-correlated, clustered gene co-expression network 

```{r, humac corncob X humac padj sig genes}

set.seed(123)
#### must run cecal_shotgun.Rmd first before this can be run
library(corncob)
library(igraph)
library(ggraph)
library(tidyverse)
library(RColorBrewer)
library(igraph)
library(ggraph)
library(tidyverse)

corn_da_humac <- read_rds("corn_da_humac")
corn_da_sqmac <- read_rds("corn_da_sqmac")
# for correlating with pairwise humac rna-seq; get humac corncob taxa subset from OG phyloseq
humac_corntax <- t(otu_table(prune_taxa(corn_da_humac$significant_taxa, psP))) %>% 
  .[!grepl("sqm", rownames(.)), ]
### the preceding block has essentially the full deseq2 workflow with draft viz and stats, but for this we want to backtrack a bit
### we want to use the rld_humac transformed counts so that
### significant, corncob ID'd microbes from the humac comparison can direct a selection of genes
### this gene set will be used to subset the res_humac dataframe from deseq() workflow, and then we can grab observations with padj < 0.1 
###### make sure rownames match for matrix correlation bw abundance matrix and rna-seq
colnames(rld_humac)
colnames(rld_humac) <- c("mac8","mac7","mac4","mac2","mac6","mac3","mac1","mac5",
                         "hum22","hum20","hum21","hum26","hum23","hum25","hum24","hum19")
#### make sure we have same samples across each matrix for 16 tot , 8per group
shared_samples <- intersect(rownames(humac_corntax), colnames(assay(rld_humac)))
## make the subsets and matrix-ify/reshape w/ t() as needed so samps are rows
microbe_mat <- humac_corntax[shared_samples, ]
expr_mat <- t(assay(rld_humac)[, shared_samples])
#### actually correlate; we're correlating all the transformed genes with all *significant* microbes from corncob
cor_HumacMat <- cor(expr_mat, microbe_mat, method = "spearman")
###### get top 500 pos/neg correlated genes per microbe
top_pos <- apply(cor_HumacMat, 2, function(x) names(sort(x, decreasing = TRUE))[1:100])
pos_genes <- unique(as.vector(as.matrix(top_pos)))
top_neg <- apply(cor_HumacMat, 2, function(x) names(sort(x, decreasing = FALSE))[1:100])
neg_genes <- unique(as.vector(as.matrix(top_neg)))
top_genes <- unique(c(pos_genes, neg_genes))

humacRes_microbeCord <- res_humac[rownames(dds_humac) %in% top_genes, ]
humac_microbeCordsig <- humacRes_microbeCord %>% filter(padj < 0.1 & pvalue < 0.05)

sqmacRes_microbeCord <- res_sqmac[rownames(dds_sqmac) %in% top_genes, ]
sqmac_microbeCordsig <- sqmacRes_microbeCord %>% filter(padj < 0.1 & pvalue < 0.05)

##expr_mat: genes x samples (only ~70 DE genes which are significant and correlated to diff abun of corncob microbes)
filtered_expr <- expr_mat[, grep(paste(intersect(rownames(humac_microbeCordsig),rownames(sqmac_microbeCordsig)), collapse = "|"), colnames(expr_mat))]
gene_symbols <- mapIds(org.Mm.eg.db, keys = colnames(filtered_expr), 
                       keytype = "ENSEMBL", 
                       column = "SYMBOL", 
                       multiVals = "first")
colnames(filtered_expr) <- ifelse(is.na(gene_symbols), colnames(filtered_expr), gene_symbols)
###correlation and matrix elongate for plotting
cor_mat <- cor((filtered_expr), method = "spearman") 
cor_mat[lower.tri(cor_mat)] <- NA  # upper triangle only
cor_df <- reshape2::melt(cor_mat, na.rm = TRUE)
names(cor_df) <- c("Gene1", "Gene2", "correlation")
##strong thresh for network
edges <- cor_df %>% filter(abs(correlation) > 0.70)
g <- graph_from_data_frame(edges, directed = FALSE)
V(g)$degree <- igraph::degree(g)

### detect louvain clustered modules; switching to infomap --
## more robust for a smaller network; less coarse network results giving us more specific focused modularity
### more robust than WGCNA which focuses on defining groups by co-expression (and also requires bit larger sample size/#input genes)
comm <- cluster_louvain(g)

V(g)$module <- factor(membership(comm))
modules <- levels(V(g)$module)

module_colors <- brewer.pal(length(modules), "Set3")
names(module_colors) <- modules

# Prepare node info
node_info_humac <- tibble(
  name = V(g)$name,
  module = V(g)$module,
  degree = V(g)$degree
)


# Identify hubs per module
top_hubs <- node_info_humac %>%
  group_by(module) %>%
  filter(degree == max(degree)) %>%
  pull(name)

# Assign shapes: triangles for hubs, circles for others
node_info_humac <- node_info_humac %>%
  mutate(
    shape_group = ifelse(name %in% top_hubs, 24, 21),  # 24=triangle, 21=circle
    outline_color = ifelse(degree >= 30, "red", "black")  # red outline for degree ≥50
  )


# Match order
node_info_humac <- node_info_humac[match(V(g)$name, node_info_humac$name), ]

# Assign to graph vertices
V(g)$module <- factor(node_info_humac$module, levels = modules)
V(g)$shape_group <- node_info_humac$shape_group
V(g)$degree <- node_info_humac$degree
V(g)$name <- node_info_humac$name
V(g)$outline_color <- node_info_humac$outline_color

# Plot
library(grid)  # for custom legend key grobs


# remove prefix 
#node_info_humac <- node_info_humac %>%
 # mutate(name = str_replace_all(name,"ENSMUSG0000", ""))


#node_info_sqmac <- node_info_sqmac %>%
 # mutate(name = str_replace_all(name,"ENSMUSG0000", ""))

```
```{r sqmac corncob X sqmac padj sig genes}}
set.seed(123)
library(corncob)
# for correlating with pairwise humac rna-seq; get humac corncob taxa subset from OG phyloseq
sqmac_corntax <- t(otu_table(prune_taxa(corn_da_sqmac$significant_taxa, psP))) %>% 
  .[!grepl("hum", rownames(.)), ]
### the preceding block has essentially the full deseq2 workflow with draft viz and stats, but for this we want to backtrack a bit
### we want to use the rld_humac transformed counts so that
### significant, corncob ID'd microbes from the humac comparison can direct a selection of genes
### this gene set will be used to subset the res_humac dataframe from deseq() workflow, and then we can grab observations with padj < 0.1 
###### make sure rownames match for matrix correlation bw abundance matrix and rna-seq
colnames(rld_sqmac)
colnames(rld_sqmac) <- c("sqm12","sqm14","sqm16","sqm18","mac8","mac7","sqm13","sqm9",
                         "sqm11","sqm10","mac4","mac2","mac6","mac3","mac1","mac5","sqm17")
#### make sure we have same samples across each matrix for 16 tot , 8per group
shared_samples <- intersect(rownames(sqmac_corntax), colnames(assay(rld_sqmac)))
## make the subsets and matrix-ify/reshape w/ t() as needed so samps are rows
microbe_mat <- sqmac_corntax[shared_samples, ]
expr_mat <- t(assay(rld_sqmac)[, shared_samples])
#### actually correlate; we're correlating all the transformed genes with all *significant* microbes from corncob
cor_SqmacMat <- cor(expr_mat, microbe_mat, method = "spearman")
###### get top 500 pos/neg correlated genes per microbe
top_pos <- apply(cor_SqmacMat, 2, function(x) names(sort(x, decreasing = TRUE))[1:100])
pos_genes <- unique(as.vector(as.matrix(top_pos)))
top_neg <- apply(cor_SqmacMat, 2, function(x) names(sort(x, decreasing = FALSE))[1:100])
neg_genes <- unique(as.vector(as.matrix(top_neg)))
top_genes <- unique(c(pos_genes, neg_genes))

length(intersect(rownames(humac_microbeCordsig),rownames(sqmac_microbeCordsig)))
##expr_mat: genes x samples (only ~70 DE genes which are significant and correlated to diff abun of corncob microbes)
filtered_expr <- expr_mat[, grep(paste(intersect(rownames(humac_microbeCordsig),rownames(sqmac_microbeCordsig)), collapse = "|"), colnames(expr_mat))]
gene_symbols <- mapIds(org.Mm.eg.db, keys = colnames(filtered_expr), 
                       keytype = "ENSEMBL", 
                       column = "SYMBOL", 
                       multiVals = "first")
colnames(filtered_expr) <- ifelse(is.na(gene_symbols), colnames(filtered_expr), gene_symbols)
###correlation and matrix elongate for plotting
cor_mat <- cor((filtered_expr), method = "spearman") 
cor_mat[lower.tri(cor_mat)] <- NA  # upper triangle only
cor_df <- reshape2::melt(cor_mat, na.rm = TRUE)
names(cor_df) <- c("Gene1", "Gene2", "correlation")
##strong thresh for network
edges <- cor_df %>% filter(abs(correlation) > 0.70)
g <- graph_from_data_frame(edges, directed = FALSE)
V(g)$degree <- igraph::degree(g)

### detect louvain clustered modules; switching to infomap --
## more robust for a smaller network; less coarse network results giving us more specific focused modularity
### more robust than WGCNA which focuses on defining groups by co-expression (and also requires bit larger sample size/#input genes)
comm <- cluster_louvain(g)

V(g)$module <- factor(membership(comm))
modules <- levels(V(g)$module)

module_colors <- brewer.pal(length(modules), "Set3")
names(module_colors) <- modules

# Prepare node info
node_info_sqmac <- tibble(
  name = V(g)$name,
  module = V(g)$module,
  degree = V(g)$degree
)

# Identify hubs per module
top_hubs <- node_info_sqmac %>%
  group_by(module) %>%
  filter(degree == max(degree)) %>%
  pull(name)

# Assign shapes: triangles for hubs, circles for others
node_info_sqmac <- node_info_sqmac %>%
  mutate(
    shape_group = ifelse(name %in% top_hubs, 24, 21),  # 24=triangle, 21=circle
    outline_color = ifelse(degree >= 30, "red", "black")  # red outline for degree ≥50
  )

# Match order
node_info_sqmac <- node_info_sqmac[match(V(g)$name, node_info_sqmac$name), ]

# Assign to graph vertices
V(g)$module <- factor(node_info_sqmac$module, levels = modules)
V(g)$shape_group <- node_info_sqmac$shape_group
V(g)$degree <- node_info_sqmac$degree
V(g)$name <- node_info_sqmac$name
V(g)$outline_color <- node_info_sqmac$outline_color




theme_bold <- theme(
  legend.title = element_text(size = 30,face = "bold"),
  legend.text = element_text(size = 28,face = "bold"),
  plot.title = element_text(size = 24,face = "bold"),
  plot.subtitle = element_text(size = 18, face = "bold"),
  strip.text = element_text(size = 16,face = "bold"),
  legend.position = c(0.1,0.88),
  panel.spacing = unit(0.8, "lines"))

layout_df <- create_layout(g, layout = "stress")

layout_df$x <- jitter(layout_df$x, amount = 0.25)

layout_df$y <- jitter(layout_df$y, amount = 0.25) 


# Plot
library(grid)  # for custom legend key grobs

Node_plot1 <- ggraph(g, layout = layout_df) +
  geom_edge_link(aes(edge_alpha = abs(correlation)), color = "grey70", show.legend = FALSE) +
geom_node_point(aes(size = degree, fill = module, shape = factor(shape_group), color = outline_color),
                stroke = 1.5,
                show.legend = c(fill = TRUE, shape = FALSE, size = FALSE, color = FALSE)) +
  geom_node_text(aes(label = name), repel = TRUE, size = 8, fontface = "bold") +
  scale_fill_manual(
    values = module_colors,
    name = "Module",
    guide = guide_legend(
      override.aes = list(shape = 21)  # legend keys are circles, colored by module
    )
  ) +
  scale_shape_manual(
    values = c(`21` = 21, `24` = 24),
    guide = "none"  # hide shape legend
  ) +
  scale_color_identity() + #use outline colors literally, no legend
  guides( # enlarge guide nodes
  fill = guide_legend(
    override.aes = list(
      size = 4.5,
      stroke = 1,
      shape = 21,
      color = "black" ))) +
  theme_void() +  
 theme(panel.border = element_rect(color = "black", fill = NA, size = 1.2)) + 
  theme_bold 

#  labs(title = "Squirrel Monkey-Macaque Donor Gene Correlation Network:",
 #      subtitle = "Louvain module, Hub Triangles and High-Degree Centrality (>=30) Highlighted nodes (red)")

ggsave("sqmac.1.node.plot.tiff", width = 12, height = 10, bg= "white")
```

```{r}
set.seed(123)
library(gridExtra)
# Use ENSEMBL IDs as vertex names for consistency:
# When creating filtered_expr, keep colnames as ENSEMBL IDs, but get symbols separately
filtered_expr_ens <- expr_mat[, grep(paste(intersect(rownames(humac_microbeCordsig), rownames(sqmac_microbeCordsig)), collapse = "|"), colnames(expr_mat))]
# Keep colnames as ENSEMBL IDs for graph construction

gene_symbols <- mapIds(org.Mm.eg.db, keys = colnames(filtered_expr_ens), 
                       keytype = "ENSEMBL", 
                       column = "SYMBOL", 
                       multiVals = "first")

cor_mat <- cor(filtered_expr_ens, method = "spearman") 
cor_mat[lower.tri(cor_mat)] <- NA
cor_df <- reshape2::melt(cor_mat, na.rm = TRUE)
names(cor_df) <- c("Gene1", "Gene2", "correlation")
edges <- cor_df %>% filter(abs(correlation) > 0.7)

g <- graph_from_data_frame(edges, directed = FALSE)
V(g)$degree <- igraph::degree(g)

comm <- cluster_louvain(g)
V(g)$module <- factor(membership(comm))
modules <- levels(V(g)$module)
module_colors <- brewer.pal(length(modules), "Set3")

node_info_sqmac <- tibble(
  name = V(g)$name,   # ENSEMBL IDs here
  module = V(g)$module,
  degree = V(g)$degree
)

top_hubs <- node_info_sqmac %>%
  group_by(module) %>%
  filter(degree == max(degree)) %>%
  pull(name)
 
# Add DE info by matching ENSEMBL IDs (graph vertex names) to res_humac rownames (also ENSEMBL IDs)
node_info_sqmac <- node_info_sqmac %>%
  mutate(
    log2FC = res_sqmac$log2FoldChange[match(name, rownames(res_sqmac))],
    padj = res_sqmac$padj[match(name, rownames(res_sqmac))],
    sig = ifelse(padj < 0.1, TRUE, FALSE),
    shape_group = ifelse(name %in% top_hubs, 24, 21),
    outline_color = ifelse(degree >= 30, "red", "black"),
    symbol = gene_symbols[match(name, names(gene_symbols))]  # this is key: gene symbols mapped by ENSEMBL IDs
  )


# Then match to graph vertices:
node_info_sqmac <- node_info_sqmac[match(V(g)$name, node_info_sqmac$name), ]

V(g)$symbol <- node_info_sqmac$symbol  # assign symbols to vertices
V(g)$log2FC <- node_info_sqmac$log2FC
V(g)$padj <- node_info_sqmac$padj
V(g)$sig <- node_info_sqmac$sig
V(g)$shape_group <- node_info_sqmac$shape_group
V(g)$outline_color <- node_info_sqmac$outline_color
V(g)$degree <- node_info_sqmac$degree


theme_bold <- theme(
  legend.title = element_text(size = 30,face = "bold"),
  legend.text = element_text(size = 22,face = "bold"),
  plot.title = element_text(size = 24,face = "bold"),
  plot.subtitle = element_text(size = 18, face = "bold"),
  strip.text = element_text(size = 16,face = "bold"),
  legend.position = c(0.2,0.88),
  panel.spacing = unit(0.8, "lines"))

layout_df <- create_layout(g, layout = "stress")

layout_df$x <- jitter(layout_df$x, amount = 0.2)

layout_df$y <- jitter(layout_df$y, amount = 0.3)

 Node_plot2 <- ggraph(g, layout = layout_df) +
  geom_edge_link(aes(edge_alpha = abs(correlation)), color = "grey70", show.legend = FALSE) +
  geom_node_point( aes(size = degree,fill = log2FC, shape = factor(shape_group), color = outline_color),  stroke = 1.5, show.legend = c(fill = TRUE, shape = FALSE, size = FALSE, color = FALSE)) +
geom_node_text(aes(label = symbol), repel = TRUE, size = 8, fontface = "bold") + 
  scale_fill_gradient2(
    low = "blue", mid = "white", high = "red", midpoint = 0,
    name = "Log2 Fold Change") +
  scale_shape_manual(
    values = c(`21` = 21, `24` = 24),
    guide = "none") +
  scale_color_identity() + # use outline_color literally (black or red)
  theme_void() + 
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1.2)
  )+
  theme_bold 
 
 # labs(
  #  title = "Macaque v Squirrel Monkey Donor Gene Correlation Network",
   # subtitle = "Microbially correlated DE genes, Triangular Hubs, High-Degree centrality (≥30) in Red"
  #)
 
ggsave("sqmac.2.node.plot.tiff", width = 12, height = 10, bg = "white")


```


```{r, humac corncob plot 1}
corn_da_humac <- read_rds("corn_da_humac")
corn_da_sqmac <- read_rds("corn_da_sqmac")
# for correlating with pairwise humac rna-seq; get humac corncob taxa subset from OG phyloseq
humac_corntax <- t(otu_table(prune_taxa(corn_da_humac$significant_taxa, psP))) %>% 
  .[!grepl("sqm", rownames(.)), ]
### the preceding block has essentially the full deseq2 workflow with draft viz and stats, but for this we want to backtrack a bit
### we want to use the rld_humac transformed counts so that
### significant, corncob ID'd microbes from the humac comparison can direct a selection of genes
### this gene set will be used to subset the res_humac dataframe from deseq() workflow, and then we can grab observations with padj < 0.1 
###### make sure rownames match for matrix correlation bw abundance matrix and rna-seq
colnames(rld_humac)
colnames(rld_humac) <- c("mac8","mac7","mac4","mac2","mac6","mac3","mac1","mac5",
                         "hum22","hum20","hum21","hum26","hum23","hum25","hum24","hum19")
#### make sure we have same samples across each matrix for 16 tot , 8per group
shared_samples <- intersect(rownames(humac_corntax), colnames(assay(rld_humac)))
## make the subsets and matrix-ify/reshape w/ t() as needed so samps are rows
microbe_mat <- humac_corntax[shared_samples, ]
expr_mat <- t(assay(rld_humac)[, shared_samples])
#### actually correlate; we're correlating all the transformed genes with all *significant* microbes from corncob
cor_HumacMat <- cor(expr_mat, microbe_mat, method = "spearman")
###### get top 500 pos/neg correlated genes per microbe
top_pos <- apply(cor_HumacMat, 2, function(x) names(sort(x, decreasing = TRUE))[1:100])
pos_genes <- unique(as.vector(as.matrix(top_pos)))
top_neg <- apply(cor_HumacMat, 2, function(x) names(sort(x, decreasing = FALSE))[1:100])
neg_genes <- unique(as.vector(as.matrix(top_neg)))
top_genes <- unique(c(pos_genes, neg_genes))

humacRes_microbeCord <- res_humac[rownames(dds_humac) %in% top_genes, ]
humac_microbeCordsig <- humacRes_microbeCord %>% filter(padj < 0.1 & pvalue < 0.05)

sqmacRes_microbeCord <- res_sqmac[rownames(dds_sqmac) %in% top_genes, ]
sqmac_microbeCordsig <- sqmacRes_microbeCord %>% filter(padj < 0.1 & pvalue < 0.05)

##expr_mat: genes x samples (only ~70 DE genes which are significant and correlated to diff abun of corncob microbes)
filtered_expr <- expr_mat[, grep(paste(intersect(rownames(humac_microbeCordsig),rownames(sqmac_microbeCordsig)), collapse = "|"), colnames(expr_mat))]
gene_symbols <- mapIds(org.Mm.eg.db, keys = colnames(filtered_expr), 
                       keytype = "ENSEMBL", 
                       column = "SYMBOL", 
                       multiVals = "first")
colnames(filtered_expr) <- ifelse(is.na(gene_symbols), colnames(filtered_expr), gene_symbols)
###correlation and matrix elongate for plotting
cor_mat <- cor((filtered_expr), method = "spearman") 
cor_mat[lower.tri(cor_mat)] <- NA  # upper triangle only
cor_df <- reshape2::melt(cor_mat, na.rm = TRUE)
names(cor_df) <- c("Gene1", "Gene2", "correlation")
##strong thresh for network
edges <- cor_df %>% filter(abs(correlation) > 0.70)
g <- graph_from_data_frame(edges, directed = FALSE)
V(g)$degree <- igraph::degree(g)

### detect louvain clustered modules; switching to infomap --
## more robust for a smaller network; less coarse network results giving us more specific focused modularity
### more robust than WGCNA which focuses on defining groups by co-expression (and also requires bit larger sample size/#input genes)
comm <- cluster_louvain(g)

V(g)$module <- factor(membership(comm))
modules <- levels(V(g)$module)

module_colors <- brewer.pal(length(modules), "Set3")
names(module_colors) <- modules

# Prepare node info
node_info_humac <- tibble(
  name = V(g)$name,
  module = V(g)$module,
  degree = V(g)$degree
)


# Identify hubs per module
top_hubs <- node_info_humac %>%
  group_by(module) %>%
  filter(degree == max(degree)) %>%
  pull(name)

# Assign shapes: triangles for hubs, circles for others
node_info_humac <- node_info_humac %>%
  mutate(
    shape_group = ifelse(name %in% top_hubs, 24, 21),  # 24=triangle, 21=circle
    outline_color = ifelse(degree >= 30, "red", "black")  # red outline for degree ≥50
  )


# Match order
node_info_humac <- node_info_humac[match(V(g)$name, node_info_humac$name), ]

# Assign to graph vertices
V(g)$module <- factor(node_info_humac$module, levels = modules)
V(g)$shape_group <- node_info_humac$shape_group
V(g)$degree <- node_info_humac$degree
V(g)$name <- node_info_humac$name
V(g)$outline_color <- node_info_humac$outline_color

# Plot
library(grid)  # for custom legend key grobs


# remove prefix 
#node_info_humac <- node_info_humac %>%
 # mutate(name = str_replace_all(name,"ENSMUSG0000", ""))


#node_info_sqmac <- node_info_sqmac %>%
 # mutate(name = str_replace_all(name,"ENSMUSG0000", ""))


theme_bold <- theme(
  legend.title = element_text(size = 30,face = "bold"),
  legend.text = element_text(size = 30,face = "bold"),
  plot.title = element_text(size =,face = "bold", hjust = ),
  plot.subtitle = element_text(size = , face = "bold"),
  strip.text = element_text(size = 16,face = "bold"),
  legend.position = c(0.1,0.88),
  panel.spacing = unit(0.8, "lines"))

layout_df <- create_layout(g, layout = "stress")

layout_df$x <- jitter(layout_df$x, amount = 0.25)

layout_df$y <- jitter(layout_df$y, amount = 0.25) 

set.seed(123)
Node_plot3 <- ggraph(g, layout = layout_df) +
  geom_edge_link(aes(edge_alpha = abs(correlation)), color = "grey70", show.legend = FALSE) +
geom_node_point(aes(size = degree, fill = module, shape = factor(shape_group), color = outline_color),
                stroke = 1.5,
                show.legend = c(fill = TRUE, shape = FALSE, size = FALSE, color = FALSE)) +
  geom_node_text(aes(label = name), repel = TRUE, size = 7.5, fontface = "bold") +
    scale_fill_manual(
    values = module_colors,
    name = "Module",
    guide = guide_legend(
      override.aes = list(shape = 21)  # legend keys are circles, colored by module
    )
  ) + theme_bold + 
  scale_shape_manual(
    values = c(`21` = 21, `24` = 24), 
    guide = "none"  # hide shape legend
  ) +
  scale_color_identity() +  # use outline colors literally, no legend
   guides( # enlarge guide nodes
  fill = guide_legend(
    override.aes = list(
      size = 4.5,
      stroke = 1,
      shape = 21,
      color = "black" ))) +
  theme_void() + 
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1.2)) +
  theme_bold 

  #labs(title = "Human-Macaque Donor Gene Correlation Network:",
     #  subtitle = "Louvain module, Hub Triangles and High-Degree Centrality (>=30) Highlighted nodes (red)")

ggsave("humac.1.node.plot.tiff", bg = "white", height = 10, width = 12 )

```

```{r}

set.seed(123)
# Use ENSEMBL IDs as vertex names for consistency:
# When creating filtered_expr, keep colnames as ENSEMBL IDs, but get symbols separately

filtered_expr_ens <- expr_mat[, grep(paste(intersect(rownames(humac_microbeCordsig), rownames(sqmac_microbeCordsig)), collapse = "|"), colnames(expr_mat))]
# Keep colnames as ENSEMBL IDs for graph construction

gene_symbols <- mapIds(org.Mm.eg.db, keys = colnames(filtered_expr_ens), 
                       keytype = "ENSEMBL", 
                       column = "SYMBOL", 
                       multiVals = "first")

cor_mat <- cor(filtered_expr_ens, method = "spearman") 
cor_mat[lower.tri(cor_mat)] <- NA
cor_df <- reshape2::melt(cor_mat, na.rm = TRUE)
names(cor_df) <- c("Gene1", "Gene2", "correlation")
edges <- cor_df %>% filter(abs(correlation) > 0.7)
######## directed  = TRUE now because we're interested in seeing how differentially expressed genes across network influence others 
######## roundabout way to try to map some kind of modularity onto this log2fc network
g <- graph_from_data_frame(edges, directed = FALSE)
V(g)$degree <- igraph::degree(g)

comm <- cluster_louvain(g)
V(g)$module <- factor(membership(comm))
modules <- levels(V(g)$module)
module_colors <- brewer.pal(length(modules), "Set3")

node_info_humac <- tibble(
  name = V(g)$name,   # ENSEMBL IDs here
  module = V(g)$module,
  degree = V(g)$degree
)

top_hubs <- node_info_humac %>%
  group_by(module) %>%
  filter(degree == max(degree)) %>%
  pull(name)

# Add DE info by matching ENSEMBL IDs (graph vertex names) to res_humac rownames (also ENSEMBL IDs)
node_info_humac <- node_info_humac %>%
  mutate(
    log2FC = res_humac$log2FoldChange[match(name, rownames(res_humac))],
    padj = res_humac$padj[match(name, rownames(res_humac))],
    sig = ifelse(padj < 0.1, TRUE, FALSE),
    shape_group = ifelse(name %in% top_hubs, 24, 21),
    outline_color = ifelse(degree >= 30, "red", "black"),
    symbol = gene_symbols[match(name, names(gene_symbols))]  # this is key: gene symbols mapped by ENSEMBL IDs
  )

# Then match to graph vertices:
node_info_humac <- node_info_humac[match(V(g)$name, node_info_humac$name), ]

V(g)$symbol <- node_info_humac$symbol  # assign symbols to vertices
V(g)$log2FC <- node_info_humac$log2FC
V(g)$padj <- node_info_humac$padj
V(g)$sig <- node_info_humac$sig
V(g)$shape_group <- node_info_humac$shape_group
V(g)$outline_color <- node_info_humac$outline_color
V(g)$degree <- node_info_humac$degree


theme_bold <- theme(
  legend.title = element_text(size = 30,face = "bold"),
  legend.text = element_text(size = 20.5,face = "bold"),
  plot.title = element_text(size = 24,face = "bold"),
  plot.subtitle = element_text(size = 18, face = "bold"),
  strip.text = element_text(size = 16,face = "bold"),
  legend.position = c(0.2,0.88),
  panel.spacing = unit(0.8, "lines"))

layout_df <- create_layout(g, layout = "stress")

layout_df$x <- jitter(layout_df$x, amount = 0.25)

layout_df$y <- jitter(layout_df$y, amount = 0.25) 



 Node_plot4 <- ggraph(g, layout = layout_df) +
  geom_edge_link(aes(edge_alpha = abs(correlation)), color = "grey70", show.legend = FALSE) +
  geom_node_point( aes(size = degree,fill = log2FC, 
                       shape = factor(shape_group), color = outline_color),  stroke = 1.5,
                   show.legend = c(fill = TRUE, shape = FALSE, size = FALSE, color = FALSE)) +
geom_node_text(aes(label = symbol), repel = TRUE, size = 8, fontface = "bold") +
  scale_fill_gradient2(
    low = "blue", mid = "white", high = "red", midpoint = 0,
    name = "Log2 Fold Change"
  ) +
  scale_shape_manual(
    values = c(`21` = 21, `24` = 24),
    guide = "none"
  ) +
  scale_color_identity() +  # use outline_color literally (black or red)
  theme_void() + 
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1.2)) +
  theme_bold 
 
 # labs(
  #  title = "Macaque v Human Donor Gene Correlation Network",
   # subtitle = "Microbially correlated DE genes, Triangular Hubs, High-Degree centrality (≥30) in Red"
  #)

ggsave("humac.2.Node.graph.tiff", width = 12, height = 10, bg = "white")
 

#Warning plot 1 and 2 come out on bottom Warning
tiff(".node.comb.tiff", units="in", width=22, height=10, res=300)

grid.arrange(Node_plot1, Node_plot2, Node_plot3, Node_plot4, nrow=2)


dev.off

 

```


## subset microbially correlated, sig observed liver genes (padj<0.1) 

```{r}

humac_microbeCordsig$Gene <- mapIds(org.Mm.eg.db, 
                         keys = as.character(rownames(humac_microbeCordsig)),
                         keytype = "ENSEMBL", column = "SYMBOL")
sqmac_microbeCordsig$Gene <- mapIds(org.Mm.eg.db, 
                         keys = as.character(rownames(sqmac_microbeCordsig)),
                         keytype = "ENSEMBL", column = "SYMBOL")

Mac_DEs_loose <- intersect(humac_microbeCordsig$Gene,sqmac_microbeCordsig$Gene)
#### 74 Genes total we report this in PAPER 1
Mac_DEs_loose

merged_df <- merge(node_info_humac, node_info_sqmac, by = "symbol", suffixes = c("_humac", "_sqmac"))
#### we should now determine which set of these genes from any given module are shared between the modular network plots -- we find 3:
#### module 1, 2, and 4
##### we report these genes in brief lists outlined in PAPER 1
merged_df %>%
  filter(module_humac == module_sqmac, module_humac == 4) %>%
#### adjust number below to select cluster to examine which genes are mutual between pw comparisons (clusters = 1,2,3,4,5; see above projections)
  pull(symbol) %>% noquote() %>% print()

### we should also consider which genes from the projected networks are shared, and mutually UPREGULATED ; PAPER 1
merged_df %>%
  filter(module_humac == module_sqmac, module_humac == 4) %>%
  filter(log2FC_humac > 0, log2FC_sqmac > 0) %>%
  pull(symbol) %>% noquote() %>% print()
### and probably also DOWNREGULATED too; PAPER 1
merged_df %>%
  filter(module_humac == module_sqmac, module_humac == 4) %>%
  filter(log2FC_humac < 0, log2FC_sqmac < 0) %>%
  pull(symbol) %>% noquote() %>% print()

```


## Normalized count for mutually microbially correlated gene set; plotting significance

```{r}

library(rstatix)

#set theme
theme_bold <- theme(
  axis.title.x = element_text(size = 38, face = "bold", color = "black"),
  axis.title.y = element_text(size = 32,face = "bold", color = "black", vjust = -5),
  axis.text.x = element_text(face = "bold", size = 19, color = "black"),
  axis.text.y = element_text(size = 24,face = "bold", color = "black"),
  legend.title = element_text(size = 36,face = "bold"),
  legend.text = element_text(size = 30 ,face = "bold"),
  plot.title = element_text(size = 0,face = "bold", hjust = 0),
  axis.line = element_line(size = 3.5, color = "black"))


meta_sqmac <- counts(dds_sqmac, normalized = TRUE) %>% reshape2::melt() %>% 
  mutate(sampleid = Var2) %>% left_join(metadata_sqmac, by = "sampleid") %>%
  mutate(Gene = mapIds(org.Mm.eg.db, keys = as.character(Var1), keytype = "ENSEMBL", column = "SYMBOL")) %>% as_tibble()
meta_humac <- counts(dds_humac, normalized = TRUE) %>% reshape2::melt() %>%
  mutate(sampleid = Var2) %>% left_join(metadata_humac, by = "sampleid") %>%
  mutate(Gene = mapIds(org.Mm.eg.db, keys = as.character(Var1), keytype = "ENSEMBL", column = "SYMBOL")) %>% as_tibble()

sqmac_macde <- filter(meta_sqmac, grepl(paste(c(Mac_DEs_loose), collapse='|'), Gene))
humac_macde <- filter(meta_humac, grepl(paste(c(Mac_DEs_loose), collapse='|'), Gene))


humac_macde <- humac_macde %>%
  mutate(Species = str_replace_all(Species, c(
    "Human" = "Hum",
    "Macaque" = "Mac",
    "Squirrel_monkey" = "Sqm"
  )))

sqmac_macde <- sqmac_macde %>%
  mutate(Species = str_replace_all(Species, c(
    "Human" = "Hum",
    "Macaque" = "Mac",
    "Squirrel_monkey" = "Sqm"
  )))

stat.test.sm <- sqmac_macde %>% group_by(Gene) %>% wilcox_test(value ~ Species) %>%
  adjust_pvalue(method = "BH") %>% filter(p.adj < 0.1 & p < 0.05) 
stat.test.hm <- humac_macde %>% group_by(Gene) %>% wilcox_test(value ~ Species) %>%
  adjust_pvalue(method = "BH") %>% filter(p.adj < 0.1 & p < 0.05)

p1 <- ggplot(sqmac_macde %>% filter(Gene %in% stat.test.sm$Gene),
       aes(x = Gene, y = value)) +
  geom_boxplot(aes(x = Gene, y = value, fill = Species), position = position_dodge(0.9)) +
  theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust=1, size = 11), axis.title.y = element_text(hjust = 3)) +
  theme_bold +
  theme(axis.text.x = ggtext::element_markdown())+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))+
  scale_fill_manual(values = c("Mac" ="#a6761d", "Sqm" = "#666666")) +
  xlab("\nGene") +
  ylab("Normalized Count\n") +
  stat_compare_means(aes(group = Species),method = 'wilcox.test', label = "p.signif", size = 3)
p2 <- ggplot(humac_macde %>% filter(Gene %in% stat.test.hm$Gene),
       aes(x = Gene, y = value)) +
  geom_boxplot(aes(fill = Species), position = position_dodge(0.9)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 11), axis.title.y = element_text(hjust = 3)) +
   theme_bold +
  theme(axis.text.x = ggtext::element_markdown()) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  scale_fill_manual(values = c("Mac" ="#a6761d", "Hum" = "#1b9e77")) +
  xlab("\nGene") + 
  ylab("Normalized Count\n") +
  stat_compare_means(aes(group = Species), method = 'wilcox.test', label = "p.signif", size = 3)

#save 
tiff("normalizedcount.diff.microb.tiff", units="in", width=22, height=14, res=300)

grid.arrange(p1,p2)

dev.off

#humac_macde %>% filter("Ncor1" %in% stat.test.hm$Gene)

#to plot specific genes

# ggplot(meta_sqmac %>% filter(Gene %in% ""),
#        aes(x = Gene, y = value)) +
#   geom_boxplot(aes(fill = Species), position = position_dodge(0.9)) +
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 11)) +
#   theme(axis.text.x = ggtext::element_markdown()) +
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
#         panel.background = element_blank(), axis.line = element_line(colour = "black")) +
#   scale_fill_manual(values = c(Macaque ="#a6761d", Squirrel_monkey = "#1b9e77")) +
#   xlab("\ngene") +
#   ylab("normalized count\n") +
#   stat_compare_means(aes(group = Species), method = 'wilcox.test', label = "p.signif", size = 3)







```


## GO term enrichment at biological process and molecular function ont. levels

```{r}

#set theme
theme_bold <- theme(
  axis.title.x = element_text(size = 32, face = "bold", color = "black"),
  axis.title.y = element_text(size = 32,face = "bold", color = "black"),
  axis.text.x = element_text(face = "bold", size = 26, color = "black"),
  axis.text.y = element_text(size = 26,face = "bold", color = "black"),
  legend.title = element_text(size = 21,face = "bold"),
  legend.text = element_text(size = 20 ,face = "bold"),
  plot.title = element_text(size = 28,face = "bold", hjust = -.5), 
  panel.border = element_rect(colour = "black", fill = NA, size = 5),
  strip.text = element_text(size = 20,face = "bold"),
  axis.line = element_line(size = 4, color = "black"),
  (plot.margin = margin(t = 1, r = 1, b = 1, l = 2000, unit = "cm"))
)

###################### grouping function does not take a universe option so this is sufficient
ggo_sharegeneBP <- groupGO(gene = na.omit(intersect(rownames(humac_microbeCordsig),rownames(sqmac_microbeCordsig))), 
                           OrgDb = org.Mm.eg.db,
                           ont      = "BP",
                           level    = 3,
                           readable = TRUE,
                           keyType = "ENSEMBL")
#View(ggo_sharegeneBP[order(-ggo_sharegeneBP$Count),])
barplot(ggo_sharegeneBP %>% arrange(-Count),
                         drop = TRUE, title = "G.O. Biol. Path", font.size = 8) + 
  geom_bar(stat = "identity", fill = "#800000") +
  theme_bold +
  theme(legend.position = "none" ,
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5) )

ggsave("Gobiol.path.tiff",height=8, width=14)
###################### grouping function does not take a universe option so this is sufficient
ggo_sharegeneMF <- groupGO(gene = na.omit(intersect(rownames(humac_microbeCordsig),rownames(sqmac_microbeCordsig))), 
                        
                           OrgDb = org.Mm.eg.db,
                           ont      = "MF",
                           level    = 3,
                           readable = TRUE,
                           keyType = "ENSEMBL")
#View(ggo_sharegeneMF[order(-ggo_sharegeneMF$Count),])
barplot(ggo_sharegeneMF %>% arrange(-Count), 
        drop = TRUE, title = "G.O. Molec. Function", font.size = 8) + 
  theme_bold +
  geom_bar(stat = "identity", fill = "#800000") +
  theme(legend.position = "none", 
        panel.grid = element_blank(),
         plot.title = element_text(hjust = 0.5) )

ggsave("Liverpathenrich.path.tiff",height=8, width=14)
##### Classic run of clusterprofiler/enrichGO function for hypergeometric path ID
ego_sharegene  <- enrichGO(
                            gene = na.omit(intersect(rownames(humac_microbeCordsig),rownames(sqmac_microbeCordsig))), 
                universe = c(rownames(res_humac),rownames(res_sqmac)),
                keyType = "ENSEMBL",
                OrgDb = org.Mm.eg.db,
                ont = "ALL",
                pAdjustMethod = "BH",
                pvalueCutoff = 0.05,
                qvalueCutoff = 0.1,
                readable = TRUE)


theme_bold <- theme(
  axis.title.x = element_text(size = 32, face = "bold", color = "black"),
  axis.title.y = element_text(size = 32,face = "bold", color = "black"),
  axis.text.x = element_text(face = "bold", size = 26, color = "black"),
  axis.text.y = element_text(size = 26,face = "bold", color = "black"),
  legend.title = element_text(size = 21,face = "bold"),
  legend.text = element_text(size = 20 ,face = "bold"),
  plot.title = element_text(size = 28,face = "bold", hjust = -.5),
  strip.text = element_text(size = 20,face = "bold"),
  axis.line = element_line(size = 4, color = "black"),
  (plot.margin = margin(t = 1, r = 1, b = 1, l = 2000, unit = "cm"))
)

#View(ego_sharegene[order(-ego_sharegene$Count),])
dotplot(ego_sharegene) +
  scale_y_discrete(labels = function(x) str_replace_all(x, "process", "")) +  
  ylab("Biological Process") +
  theme_classic() +
  theme( 
  axis.line.x.bottom = element_line(size = 3, color = "black"),
  axis.line.x.top    = element_line(size = 3, color = "black"),
  axis.line.y.left   = element_line(size = 3, color = "black"),
  axis.line.y.right  = element_line(size = 3, color = "black"),
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  panel.grid = element_blank())+ 
  theme_bold 


ggsave("Go-all-onts.tiff", height = 6, width =11)



humac_microbeCordsig$Ent <-  mapIds(org.Mm.eg.db, keys = rownames(humac_microbeCordsig), keytype = "ENSEMBL", column = "ENTREZID")
sqmac_microbeCordsig$Ent <-  mapIds(org.Mm.eg.db, keys = rownames(sqmac_microbeCordsig), keytype = "ENSEMBL", column = "ENTREZID")

res_humac$Gene <-  mapIds(org.Mm.eg.db, keys = rownames(res_humac), keytype = "ENSEMBL", column = "SYMBOL")
res_sqmac$Gene <-  mapIds(org.Mm.eg.db, keys = rownames(res_sqmac), keytype = "ENSEMBL", column = "SYMBOL")

##### Classic run of clusterprofiler/enrichkegg function for hypergeometric path ID
mapping <- AnnotationDbi::select(
  org.Mm.eg.db,
  keys = c(rownames(res_humac),rownames(res_sqmac)),
  keytype = "ENSEMBL",
  columns = "ENTREZID"
) 
kegg_res <- enrichKEGG(gene = intersect((humac_microbeCordsig$Ent),(sqmac_microbeCordsig$Ent)),
                       universe = mapping$ENTREZID,organism = "mmu", keyType = "kegg",
                       pvalueCutoff = 0.05, qvalueCutoff = 0.1, pAdjustMethod = "BH",
                       minGSSize = 10, maxGSSize = 500, use_internal_data = FALSE)

# add columns

ego_sharegene@result$label <- "GO_Biological_Process"
kegg_res@result$label <- "KEGG_Pathway"


#filter ontology out
ego_sharegene@result$ONTOLOGY <- NULL

#remove more
kegg_res@result$category <- NULL
kegg_res@result$subcategory <- NULL


#make dataframe
GO_Bio_gene <- ego_sharegene@result

#remove process 
GO_Bio_gene <- GO_Bio_gene %>%
  mutate( Description = str_replace_all(Description, "process", ""))

#rbind
combined_enrich <- rbind(kegg_res@result, GO_Bio_gene)

#load packages
library(ggplot2)
library(dplyr)
library(forcats)

#take each plots terms
#terms_kegg <- kegg_res@result %>%
 # slice_min(order_by = pvalue, Description n = 9) %>%
#  pull(Description)

#terms_go <- ego_sharegene@result %>%
 # slice_min(order_by = p.adjust, ), n = 10) %>%
  #pull(Description)


target_kegg <- c("Complement and coagulation cascades", "Glycine, serine and threonine metabolism",
"Fatty acid metabolism", "Retinol metabolism", "Inflammatory mediator regulation of TRP channels",
"Linoleic acid metabolism", "2-Oxocarboxylic acid metabolism", "Cysteine and methionine metabolism",
"Ovarian steroidogenesis")

target_go <-  c("small molecule biosynthetic ","monocarboxylic acid metabolic ","organic acid biosynthetic ","carboxylic acid biosynthetic ","lipid biosynthetic ","fatty acid metabolic ","sulfur compound metabolic ","monocarboxylic acid biosynthetic ","acyl-CoA metabolic ","thioester metabolic ")


Go_biocheck <- GO_Bio_gene %>%
  filter(Description %in% (target_go))


#filter to have those terms 
combined_filtered <- combined_enrich %>%
  filter(Description %in% c(target_go, target_kegg))

# Convert GeneRatio from "x/y" string to numeric
combined_filtered$GeneRatio <- sapply(
  strsplit(as.character(combined_filtered$GeneRatio), "/"),
  function(x) as.numeric(x[1]) / as.numeric(x[2])
)


#build top terms
top_terms <- combined_filtered %>%
  arrange(label, GeneRatio) %>%
  mutate(term = paste(label, Description, sep = "___")) %>%
  mutate(term = factor(term, levels = rev(unique(term)))) %>%
  group_by(label) %>%
  # Order each group so lowest GeneRatio comes first
  arrange(GeneRatio, .by_group = TRUE) %>%
  # Create a term factor with group-wise ordering preserved
  mutate(term = factor(Description, levels = unique(Description))) %>%
  ungroup()

# Arrange terms so lower GeneRatio appears at the bottom of the plot
#top_terms <- top_terms %>%
 # arrange(label, GeneRatio) %>%
  #mutate(term = fct_inorder(Description)) %>%  # preserve order for plotting
  #mutate(term = fct_rev(term))                 # reverse so low ratios are at bottom
 
#order
top_terms <- top_terms %>%
  mutate(label = factor(label, levels = c("KEGG_Pathway", "GO_Biological_Process")))



#for fixing
top_terms <- top_terms %>%
  mutate(label = str_replace_all(label, "_", " "))

top_terms <- top_terms %>%
  mutate(label = str_replace_all(label, "GO", ""))

# Assign group colors
label_colors <- c(" Biological Process" = "darkviolet", "KEGG Pathway" = "darkgreen")


#order labels on right side
label_positions <- top_terms %>%
  group_by(label) %>%
  summarise(y = mean(as.numeric(term))- 3, .groups = "drop") %>%
  mutate(color = label_colors[label])




#change position
#top_terms <- top_terms %>%
  # Make sure KEGG comes first in the label order
 # mutate(label = factor(label, levels = c("KEGG Pathway", "Biological Process"))) %>%
  # Arrange data by label, then by GeneRatio within each group
  #arrange(label, GeneRatio) %>%
  # Build a unique term label that preserves group structure
  #mutate(term = paste(label, Description, sep = "___")) %>%
  # Reverse the term order so KEGG ends up visually at the bottom
  #mutate(term = factor(term, levels = rev(unique(term))))

top_terms <- top_terms %>%
  mutate(term_color = label_colors[label])


top_terms <- top_terms %>%
  arrange(GeneRatio) %>%                                   # low → high across all rows
  mutate(term = factor(Description, levels = unique(Description)))

#plot

ggplot(top_terms, aes(x = GeneRatio, y = term)) +
  # Draw connector lines from baseline to each dot
  geom_segment(aes(x = 0, xend = GeneRatio, yend = term),
               color = "gray70", size = 0.5) +
  geom_point(aes(size = Count, color = p.adjust), alpha = 0.9) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray40") +
  geom_text(data = label_positions,
            aes(x = max(top_terms$GeneRatio) + 0.2, y = y, label = label),
            inherit.aes = FALSE,
            angle = 90, hjust = 0.1,
            fontface = "bold", size = 10.5,
            colour = label_colors[label_positions$label]) +
  scale_color_gradient(low = "red", high = "blue", name = "Adjusted p-value") +
  scale_size(range = c(3, 7.5)) +
  labs(
    x = "Gene Ratio",
    y = NULL,
    size = "Gene Count",
    color = "Adjusted p-value"
  ) +
  # Apply classic theme 
theme_classic(base_size = 14) +
  theme(
    axis.text.y = element_text(face = "bold", color =  "#00000000", size = 21),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(face = "bold", color = "black", size =22.8),
    axis.title.x = element_text(face = "bold", size = 24),
    legend.title = element_text(face = "bold", size = 24),
    legend.text = element_text(size = 21), 
    plot.margin = margin(t = 1, r = 1.5, b = 1, l = 0, unit = "cm"),
    panel.border = element_rect(color = "black", fill = NA, size = 1)
  ) + 
  #scale_y_discrete(labels = NULL) +
  geom_text(data = top_terms,
          aes(x = -Inf, y = term, label = Description),
          inherit.aes = FALSE,
          hjust = 1.02,
          size = 7.2, fontface = "bold", colour = top_terms$term_color) +
  coord_cartesian(clip = "off") 




#save to directory as tiff
ggsave("combinedKegg_Go.tiff", height = 10, width = 18)

dotplot(kegg_res)  +
  scale_y_discrete(labels = function(x) str_replace_all(x, "process", "")) +  
  ylab("KEGG Pathway") +
  theme_classic() + 
      geom_point() +
      theme(legend.background = element_rect(fill = NA),
  axis.line.x.bottom = element_line(size = 3, color = "black"),
  axis.line.x.top    = element_line(size = 3, color = "black"),
  axis.line.y.left   = element_line(size = 3, color = "black"),
  axis.line.y.right  = element_line(size = 3, color = "black"),
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank())+ 
  theme_bold 


ggsave("KEGG.3.tiff", height = 6, width =13.45)



```


## Plot on enrichr universe of results


```{r}

######## not working with background set... which... incidentally is something they also don't make available via ma'ayan lab gui
######## https://maayanlab.cloud/Enrichr/
# enrichres_backid <- rba_enrichr_add_background(background_genes = c(res_humac$Gene,res_sqmac$Gene))
# 
# enrichres <- rba_enrichr(gene_list = intersect(humac_microbeCordsig$Gene,sqmac_microbeCordsig$Gene),
#                          #gene_set_library = "all",
#                          background_genes = enrichres_backid)
########## we don't have a background so this is inherently riskier to interpret from than results like those above, BUT
########## we have so many more databases to access in using enrichr (218)-- maybe results from a run of enrichr can at least help contextualize what we ########## observed from our run of clusterprofiler
enrichres <- rba_enrichr(gene_list = intersect(humac_microbeCordsig$Gene,sqmac_microbeCordsig$Gene),
                         gene_set_library = "all")
enrichres_filt_hits <- lapply(enrichres, function(x)
       filter(x, P.value < 0.05 &
                Adjusted.P.value < 0.1 &
                as.numeric(sub("/.*", "", Overlap)) >= 20 &
                str_detect(Term, c("cho|fatty|fat|lip|amino|amine|prop|but|acet|bil|scfa|short-chain|short chain"))
              )
       )
enrichres_filt_score <- lapply(enrichres, function(x)
       filter(x, P.value < 0.05 &
                Adjusted.P.value < 0.1 &
                Combined.Score > 1000 &
                #as.numeric(sub("/.*", "", Overlap)) >= 38 &
                #Combined.Score > 1500 &
                str_detect(Term, c("cho|fatty|fat|lip|amin|amino|amine|prop|but|acet|bil|scfa|short-chain|short chain"))
              )
       )



theme_bold <- theme(
  axis.title.x = element_text(size = 16, face = "bold"),
  axis.title.y = element_text(size = 16,face = "bold"),
  axis.text.x = element_text(face = "bold", size = 14, color = "black"),
  axis.text.y = element_text(size = 14,face = "bold", color = "black"),
  legend.title = element_text(size = 14.8,face = "bold"),
  legend.text = element_text(size = 14,face = "bold"),
  plot.title = element_text(size = 22,face = "bold", hjust = .5),
  strip.text = element_text(size = 16,face = "bold"),
  axis.line = element_line(size = 2.2, color = "black"),
  panel.spacing = unit(0.8, "lines"))
  

enrichres_filt_hits <- purrr::discard(enrichres_filt_hits, ~nrow(.) == 0)
enrichres_filt_score <- purrr::discard(enrichres_filt_score, ~nrow(.) == 0)
enrichr_list <- c(enrichres_filt_hits,enrichres_filt_score)
#combine dfs across enrichr results and filter for columns to plot on (e.g. padj, db info, res/term to plot on y);sort on decr padj val
combined_df <- bind_rows(enrichr_list, .id = "database") %>%
  filter(!is.na(Adjusted.P.value)) %>%
  select(Term, Adjusted.P.value, database) %>%
  group_by(database) %>%
  slice_min(order_by = Adjusted.P.value, n = 10, with_ties = FALSE) %>%
  ungroup() %>% unique()

combined_df <- combined_df %>%
  mutate(database = str_replace_all(database, "_", " "))
#install and load palletteer
#install.packages("paletteer")
library(paletteer)
paletteer_d("nbapalettes::jazz_city")

# Step 1: Combine and clean all enrichment results
combined_df <- bind_rows(enrichr_list, .id = "database") %>%
  filter(!is.na(Adjusted.P.value)) %>%
  mutate(database = str_replace_all(database, "_", " "))

# Step 2: Select top 20 terms by significance (across all databases)
top_terms <- combined_df %>%
  arrange(Adjusted.P.value) %>%
  distinct(Term, .keep_all = TRUE) %>%
  slice_head(n = 20) %>%
  pull(Term)

# Step 3: Filter full dataset to keep all database entries for those 20 terms
plot_df <- combined_df %>%
  filter(Term %in% top_terms)


# Step 4: Plot only those terms
ggplot(plot_df, aes(x = reorder(Term, -Adjusted.P.value), y = -log10(Adjusted.P.value), fill = database)) +
  geom_col(position = position_dodge()) +
  coord_flip() +
  labs(title = "",
       x = "",
       y = "-log10 Adjusted P-value",
       fill = "Database") +
  theme_minimal() + 
  theme_bold +
  theme(panel.grid = element_blank())+
  scale_fill_manual(
  values = c(
    "#010101FF", "#5D2A2CFF", "#C8102EFF","#DC582AFF", "#FFC72CFF"
  )
)

ggsave("enrichd.new.color.tiff", bg = "white", height = 7, width = 14.4)

# Plot all bars, colored by database
ggplot(combined_df, aes(x = reorder(Term, -Adjusted.P.value), y = -log10(Adjusted.P.value), fill = database)) +
  geom_col(position = position_dodge()) +
  coord_flip() +
  labs(title = "Top Enrichr Terms Across Databases",
       x = "",
       y = "-log10 Adjusted P-value",
       fill = "Database") +
  theme_minimal() + theme_bold 


combined_df %>% arrange(Adjusted.P.value) %>% head(n = 21) %>% pull(Term)

ggsave("enrichd.tiff", bg = "white", height = 7, width = 12)

# terms <- combined_df %>% arrange(Adjusted.P.value) %>% head(n = 21) %>% pull(Term)
# library(stringdist)
# dist_mat <- stringdistmatrix(terms, terms, method = "cosine", q = 2)
# sim_mat <- 1 - dist_mat  # similarity = 1 - distance
# 
# rownames(sim_mat) <- colnames(sim_mat) <- terms
# mds <- cmdscale(as.dist(1 - sim_mat), k = 2)
# mds_df <- as.data.frame(mds)
# mds_df$term <- terms
# 
# ggplot(mds_df, aes(V1, V2, label = term)) +
#   geom_point() +
#   geom_text_repel() +
#   theme_minimal() +
#   labs(title = "MDS of Pathway Term Similarity", x = "Dim 1", y = "Dim 2")

```






```{r}
# Basically validated findings from above 

# enrichres_filt_hitsHmac <- lapply(enrichres_humac, function(x)
#        filter(x, P.value < 0.05 &
#                 Adjusted.P.value < 0.05 &
#                 #Combined.Score > 1500 &
#                 str_detect(Overlap, stringr::str_flatten(paste("^",c(50:100), sep = ""), collapse = "|"))
#               )
#        )
# 
# enrichres_filt_scoreHmac <- lapply(enrichres_humac, function(x)
#        filter(x, P.value < 0.05 &
#                 Adjusted.P.value < 0.05 &
#                 Combined.Score > 1500 &
#                 str_detect(Term, c("bile|cho|fatty|fat|amino|amine"))
#               )
#        )
# enrichres_filt_hitsHmac <- purrr::discard(enrichres_filt_hitsHmac, ~nrow(.) == 0)
# enrichres_filt_scoreHmac <- purrr::discard(enrichres_filt_scoreHmac, ~nrow(.) == 0)
# 
# View(enrichres_filt_hitsHmac)
# View(enrichres_filt_scoreHmac)
# Basically validated findings from above 
# enrichres_filt_hitsSmac <- lapply(enrichres_sqmac, function(x)
#        filter(x, P.value < 0.05 &
#                 Adjusted.P.value < 0.05 &
#                 #Combined.Score > 1500 &
#                 str_detect(Overlap, stringr::str_flatten(paste("^",c(50:100), sep = ""), collapse = "|"))
#               )
#        )
# enrichres_filt_scoreSmac <- lapply(enrichres_sqmac, function(x)
#        filter(x, P.value < 0.05 &
#                 Adjusted.P.value < 0.05 &
#                 Combined.Score > 1500 &
#                 str_detect(Term, c("bile|cho|fatty|fat|amino"))
#               )
# 
#        )
# enrichres_filt_hitsSmac <- purrr::discard(enrichres_filt_hitsSmac, ~nrow(.) == 0)
# enrichres_filt_scoreSmac <- purrr::discard(enrichres_filt_scoreSmac, ~nrow(.) == 0)
# 
# View(enrichres_filt_hitsSmac)
# View(enrichres_filt_scoreSmac)

```


The ERV genes I was previously interested showed no distinct differential expression patterning between groups. Nup153, an ERV-associated gene shows little DE patterning.
```{r ERV explorer}

# ## ANNOTATE H SAPIENS
# mousetab <- read.table("/Users/apulvino/AmatoLab/QualPrelimAnalysis/mouse_anno.txt", sep = "\t", header = TRUE)
# res.framemouse <- data.frame(res)
# res.framemouse$Gene.stable.ID <- rownames(res.framemouse)
# 
# mouseres <- merge(mousetab, res.framemouse, by = "Gene.stable.ID")
# 
# ## use interpro annotation to search gag,pro,pol,erv,retroelement,etc keywords that will help you ID retroelements of interest between papio and human here
# 
# #read homo interpro tab
# mouseinterpro <- read.delim("/Users/apulvino/AmatoLab/QualPrelimAnalysis/mouse_interpro.txt", header = TRUE, sep = "\t")
# mouseres_final <- merge(mouseinterpro, mouseres, by = "Gene.stable.ID")
# 
# # subset homo gut wild expression results for just gag,pol,pro,retrovirus,and retroviral entries
# mouse_retro <- dplyr::filter(mouseres_final, grepl("retrovirus|Retrovirus|retroviral|Retroviral|retrovirus-like|Retrovirus-like|retrotransposon|Retrotransposon|Retro-transposon|retro-transposon|retrotransposon-derived|Retrotransposon-derived|retroelement|Retroelement",mouseres_final$Interpro.Description))
# 
# mouse_retro_uniq <- unique(mouse_retro)
# 
# ### ACADS
# plotCounts(dds, gene="ENSMUSG00000029545", intgroup = "Species")
# ### CDO1
# plotCounts(dds, gene="ENSMUSG00000033022", intgroup = "Species")
# plotCounts(dds, gene="ENSMUSG00000029630", intgroup = "Species")
# 
#plotCounts(dds_humac, gene="ENSMUSG00000022463", intgroup = "Species")
# plotCounts(dds, gene="ENSMUSG00000063856", intgroup = "Species")

```

