---
title: "Serum_metabolomics"
author: "Emma Christopher"
date: "2025-09-02"
output: html_document
---
\*\* Some of the data were not being tested with wilcox test but t test. Since metabolomics data are often not normally distributed, I went back and ensured all statistical tests performed were with wilcox.

We begin by loading the object from the preprocessing work done using tidyMass which culminated in generating a "mass_dataset" analysis object to use for data viz.

A preliminary set of PCAs on the object as a whole reveals that macaques may be a better outgroup than squirrel monkeys. It appears that humans microbial communities might also express a unique metabolic flexibility that squirrel monkeys may be similarly capable of. Of note, the last two PCAs are log scaled, and the first is not. Additionally, the first also depicts the loadings which indicate. I am still working on how to get compound name labels for the loadings over the annotation IDs so for now they remain as just the arrows.


##package loading

```{r, PCA of data, echo = FALSE, message = FALSE}
getwd()
set.seed(100)
knitr::opts_chunk$set(fig.width=10, fig.height=8) 
library(dplyr)
library(remotes)
#install.packages("furrr")
library(furrr)
#if (!require("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")
#BiocManager::install(version = "3.21")
library(BiocManager)
#if(!require(remotes)){
#install.packages("remotes")
#}
#remotes::install_gitlab("tidymass/tidymass")
library(tidymass)
library(massstat)
library(metpath)
library(massqc)
library(tidymass)
library(ggfortify)
library(vegan)
#install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
library(pairwiseAdonis)
#install.packages("factoextra")
library(factoextra)
library(ggplot2)
library(tidygraph)
library(RColorBrewer)
library(stringr)
library(webchem)
library(data.table)
library(KEGGREST)
#BiocManager::install("clusterProfiler", force = TRUE)
library(clusterProfiler)
library(readr)
## load the saved object created from tidyMassObjectGeneration.Rmd which includes some filtering and preprocessing before the actual analysis and data viz is done
### if you rerun the object generation script, you will find you pass a subtle warning/error message near the annotation step for the positive scan
### this is not a real issue -- it's just that we happen to not identify with our data any positive scans from alignment against the fiehn lab's database (only getting a couple hits on the negative scans for that one)
```

##Serum PCA
```{r, PCA of data, echo = FALSE, message = FALSE}
load("object2")

pca_object =
object %>%
  scale() %>%
  run_pca()
#pca_object$rotation


# PAPER 1
# tiff("qual2pca.tiff", units="in", width=8.5, height=5, res=300)

sqm.label = 8
hum.label = 8
mac.label = 8
axis.size = 3
axis.title = 20
title.size = 30

png("Serum_Metabolomics_PCA.jpeg", units="in", width=8, height=6, res=300)
object %>%
  `+`(1) %>%
  log %>%
  scale %>%
  massqc_pca(color_by = "group", line = FALSE) +
  scale_fill_manual(values = c(hum="#1b9e77",mac="#a6761d",sqm="#666666")) +
  scale_color_manual(values = c(hum="#1b9e77",mac="#a6761d",sqm="#666666")) +
  guides(
    fill = guide_legend(override.aes = list(colour = NA, stroke = 0)),
    color = guide_legend(override.aes = list(stroke = 0))
  )+ 
  geom_point(shape = 16, size = 5, alpha = 0.05) +
  theme(legend.key = element_rect(fill = "transparent", colour = "transparent")) +
  theme(axis.text = element_text(size = 20, face = "bold", colour = "black"),
        axis.line.x = element_line(size = axis.size, colour = "black"),
        axis.line.y = element_line(size = axis.size, colour = "black"),
        axis.title.x = element_text(size = axis.title, face = "bold", colour = "black"),
        axis.title.y = element_text(size = axis.title, face = "bold", colour = "black"),
        axis.ticks= element_blank(),
        plot.title = element_text(size = title.size, face = "bold", colour = "black", hjust = 0.5),
        panel.background=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.border = element_blank(),
        legend.position = "none",
        )+ 
  labs(title = "") +
  annotate("text", x = 0.55, y = 0.65, label = "sqm", color = "#666666", size = sqm.label, fontface = "bold") +
  annotate("text", x = -0.40, y = -0.65, label = "mac", color = "#a6761d", size = mac.label, fontface = "bold") +
  annotate("text", x = 0.65, y = -0.6, label = "hum", color = "#1b9e77", size = hum.label, fontface = "bold")
dev.off()

exp_data <- object@expression_data
exp_data$variable_id <- rownames(exp_data)
anno_data <- object@annotation_table
peak_table <- merge(anno_data, exp_data, by = "variable_id")
write_tsv(peak_table, "/Users/emmachristopher/Desktop/Northwestern/Anthony_metabolomics/Serum_Metabolomics/4nSerMetabol.tsv")


```


##PERMANOVA (not done)

```{r, 3-GROUP PERMANOVA }

library(phyloseq)
library(vegan)

#reading in phyloseq
phyloseq_object <- readRDS("psP.Rds")

#extracting sample data
metadata <- as(sample_data(phyloseq_object), "data.frame")
otu <- otu_table(phyloseq_object)
otu_df <- as.data.frame(otu)
distance_matrix <- distance(phyloseq_object, method = "bray")

#### remember to run these permanovas in this doc, just need to make sure you're first running the first block in the cecal_shotgun.Rmd
# ### for all three groups, determine significance of term "group" effect
dimnames <- c("mac01","mac03","mac04","mac06","mac07",
              "mac08","sqm09","sqm10","sqm11","sqm12",
              "sqm13","sqm14","sqm15","sqm16","sqm17",
              "sqm18", "hum19", "hum20","hum21","hum22","hum23","hum24","hum25","hum26")
#### ALL, 3-group pw comparison
 sample_distance <-
   as.matrix(dist_mass_dataset(x = object, margin = "sample", method = "manhattan"))
#rownames(sample_distance) <- dimnames
#colnames(sample_distance) <- dimnames
result <- metadata %>% filter(!sample %in% c("mac2","mac5"))
## need to remove samp 2 and samp 5 from sample_distance
mat_cleaned <- sample_distance[!grepl("Ben-Oli-20230331-02|Ben-Oli-20230331-05",rownames(sample_distance)),
                               !grepl("Ben-Oli-20230331-02|Ben-Oli-20230331-05", colnames(sample_distance)), drop = FALSE]
print("global effect of group permanova")
adonis2(mat_cleaned ~ group, data = result)



```

## SIRIUS classification (for database-free metabolite classification and FELLA enrichment)

```{r, sirius}

library(tidyverse)
library(stats)
library(ggplot2)
library(scales)
library(purrr)
############ ATTEMPTING TO MERGE WITH TIDYMASS THE SIRIUS OUTPUT FOR GROUPWISE ANALYSIS WITH RESULTS
############## THIS IS VERY CLOSE AND WOULD BE EASY TO FINISH
# # #download from SIRIUS run on quest to computer
df <- read_tsv("dualion_canopus_formula_summary_all.tsv")
canopus_formula <- df
# # 
# # # 1. Clean Canopus ID to isolate mappingFeatureId
canopus_clean <- canopus_formula %>%
   mutate(variable_id = str_remove(mappingFeatureId, ".*UNKNOWN_"))
# 
# # # 2. Merge with tidymass object using mappingFeatureId == alignedFeatureId
merged <- left_join(object@variable_info, canopus_clean, by = "variable_id")

# Your full dataset
df_bio <- read.delim("dualion_canopus_formula_summary_all.tsv",
                      sep = "\t", stringsAsFactors = FALSE, header = TRUE) 
df_bio_clean <- df_bio %>% filter(!pmap_lgl(., ~ any(c(...) == colnames(df_bio))))

df_free <- read.delim("dualion_canopus_formula_summary_all.tsv",
                     sep = "\t", stringsAsFactors = FALSE, header = TRUE)
df_free_clean <- df_free %>% filter(!pmap_lgl(., ~ any(c(...) == colnames(df_free))))

df_all <- read.delim("dualion_canopus_formula_summary_all.tsv",
                      sep = "\t", stringsAsFactors = FALSE, header = TRUE)
df_all_clean <- df_all %>% filter(!pmap_lgl(., ~ any(c(...) == colnames(df_all))))

# Pairs of (class_col, prob_col) to analyze and plot
classification_cols <- list(
  c("NPC.pathway", "NPC.pathway.Probability"),
  c("NPC.superclass", "NPC.superclass.Probability"),
  c("NPC.class", "NPC.class.Probability"),
  c("ClassyFire.superclass", "ClassyFire.superclass.probability"),
  c("ClassyFire.class", "ClassyFire.class.Probability"),
  c("ClassyFire.subclass", "ClassyFire.subclass.Probability"),
  c("ClassyFire.level.5", "ClassyFire.level.5.Probability"),
  c("ClassyFire.most.specific.class", "ClassyFire.most.specific.class.Probability")
)

library(dplyr)
library(ggplot2)

plot_class_enrichment <- function(df, class_col, prob_col, min_count = 5, conf_thresh = 0.9) {
  ##ensures prob col and class col are evaluated as numeric/string
  df <- df %>%
    mutate(
      !!sym(prob_col) := as.numeric(.data[[prob_col]]),
      !!sym(class_col) := as.character(.data[[class_col]])
    )
  ##filter missing vals group by class and compute median probability column
  summary_df <- df %>%
    filter(!is.na(.data[[class_col]]), !is.na(.data[[prob_col]])) %>%
    group_by(!!sym(class_col)) %>%
    summarise(
      n = n(),
      median_prob = median(.data[[prob_col]], na.rm = TRUE),
      .groups = "drop"
    ) %>%
    ##filter for compounds in common classes and high_conf = TRUE (if prob for compound class is greater than median)
    ##then we do log2 ratio to plot data in fold change style for relative enrichment measure
    filter(n >= min_count) %>%
    mutate(
      high_conf = median_prob > conf_thresh,
      centered_log2_count = log2(n / median(n))
    )
  ###plotting code, want to make sure that classes that are enriched
  ###e.g. above or below median threshold are colored blue/distinguished in some way
  plot <- ggplot(summary_df, aes(
    x = centered_log2_count,
    y = reorder(!!sym(class_col), centered_log2_count),
    fill = high_conf
  )) +
    geom_col() +
    geom_text(aes(label = paste0("n=", n)),
              hjust = ifelse(summary_df$centered_log2_count > 0, -0.1, 1.1),
              size = 3,
              color = "black") +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray40") +
    scale_fill_manual(values = c("TRUE" = "blue", "FALSE" = "grey70")) +
    labs(
      title = paste0("Class Enrichment for ", class_col),
      x = expression(log[2]*"(count / median count)"),
      y = class_col,
      fill = paste0("Median prob > ", conf_thresh)
    ) +
    theme_minimal(base_size = 14) +
    coord_cartesian(clip = "off") +
    theme(
      legend.position = "none",
      plot.title.position = "plot",
      plot.margin = margin(10, 40, 10, 10)
    )
  
  return(list(plot = plot, data = summary_df))
}

#get data for an  plotting 
results <- plot_class_enrichment(df_free_clean, "ClassyFire.superclass", "ClassyFire.superclass.probability")
head(results$data)

plots_free <- list()

for (cols in classification_cols) {
  class_col <- cols[1]
  prob_col <- cols[2]
  
  p <- plot_class_enrichment(df_free_clean, class_col, prob_col)
  print(p)  # Show plot in console
  
  # Save plots in a named list
  plots_free[[class_col]] <- p
}
setwd("/Users/emmachristopher/Desktop/Northwestern/Anthony_metabolomics/Serum_Metabolomics")
for (name in names(plots_free)) {
  this_plot <- plots_free[[name]][["plot"]]
  filename <- paste0("plot_free_", name, ".png")
  
  #message("Saving: ", filename)
  
  #ggsave(
   # filename = filename,
    #plot = this_plot,
    #width = 8,
    #height = 6,
    #dpi = 300
  #)
}

############### i feel like these are all kind of giving me the same basic result -- differentially abundant classes include carboxylic acids; fatty acids; amino acids .... going to go ahead and take the bio results (also happens to have the most observations/annotations) for a run of FELLA

```

###class enrichment cleaning
```{r, sirius}
#Class Enrichment for ClassyFire.class cleaning

results <- plot_class_enrichment(
  df_free_clean, 
  "ClassyFire.class", 
  "ClassyFire.class.Probability")
print(results$plot)

# Parameters
class_col <- "ClassyFire.class"
prob_col <- "ClassyFire.class.Probability"
min_count <- 5
conf_thresh <- 0.9

# Clean data
df_clean <- df_free_clean %>%
  mutate(
    class = as.character(.data[[class_col]]),
    prob = as.numeric(.data[[prob_col]])
  ) %>%
  filter(!is.na(class), !is.na(prob))

summary_df <- df_clean %>%
  group_by(class) %>%
  summarise(
    n = n(),
    median_prob = median(prob, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(n >= min_count) %>%
  mutate(
    high_conf = median_prob > conf_thresh,
    centered_log2_count = log2(n / median(n))
  )

#plot
class <- ggplot(summary_df, aes(
  x = centered_log2_count,
  y = reorder(class, centered_log2_count),
  fill = high_conf
)) +
  geom_col() +
  geom_text(aes(label = paste0("n=", n)),
            hjust = ifelse(summary_df$centered_log2_count > 0, -0.1, 1.1),
            size = 4,
            color = "black",
            fontface = "bold") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray40") +
  scale_fill_manual(values = c("TRUE" = "blue", "FALSE" = "grey70")) +
  labs(
    title = "",
    x = expression(bold(log[2]*"(count / median count)")),
    y = "",
    fill = paste0("Median prob > ", conf_thresh)
  ) +
  theme_minimal(base_size = 14) +
  coord_cartesian(clip = "off") +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.margin = margin(10, 40, 10, 10),
    axis.text.x = element_text(size = 14, face = "bold", colour = "black"),
    axis.text.y = element_text(size = 14, face = "bold", colour = "black", hjust = 1),
    plot.title = element_text(size = 20, face = "bold", colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
print(class)
# ggsave(
#     filename = "ClassyFire_Class_clean.jpeg",
#     plot = class,
#     width = 10,
#     height = 6,
#     dpi = 300,
#   )
```


```{r, sirius}
#Class Enrichment for ClassyFire.subclass cleaning

results <- plot_class_enrichment(
  df_free_clean, 
  "ClassyFire.subclass", 
  "ClassyFire.subclass.Probability")
print(results$plot)

# Parameters
class_col_sub <- "ClassyFire.subclass"
prob_col_sub <- "ClassyFire.subclass.Probability"
min_count <- 5
conf_thresh <- 0.9

#Clean data
df_clean <- df_free_clean %>%
  mutate(
    class = as.character(.data[[class_col_sub]]),
    prob = as.numeric(.data[[prob_col_sub]])
  ) %>%
  filter(!is.na(class), !is.na(prob))

# Summarize
summary_df <- df_clean %>%
  group_by(class) %>%
  summarise(
    n = n(),
    median_prob = median(prob, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(n >= min_count) %>%
  mutate(
    high_conf = median_prob > conf_thresh,
    centered_log2_count = log2(n / median(n))
  )

# Plot 
subclass <- ggplot(summary_df, aes(
  x = centered_log2_count,
  y = reorder(class, centered_log2_count),
  fill = high_conf
)) +
  geom_col() +
  geom_text(aes(label = paste0("n=", n)),
            hjust = ifelse(summary_df$centered_log2_count > 0, -0.1, 1.1),
            size = 4,
            color = "black",
            fontface = "bold") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray40") +
  scale_fill_manual(values = c("TRUE" = "blue", "FALSE" = "grey70")) +
  labs(
    title = "",
    x = expression(bold(log[2]*"(count / median count)")),
    y = "",
    fill = paste0("Median prob > ", conf_thresh)
  ) +
  theme_minimal(base_size = 14) +
  coord_cartesian(clip = "off") +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.margin = margin(10, 40, 10, 10),
    axis.text.x = element_text(size = 14, face = "bold", colour = "black"),
    axis.text.y = element_text(size = 14, face = "bold", colour = "black", hjust = 1),
    plot.title = element_text(size = 20, face = "bold", colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
print(subclass)
# ggsave(
#     filename = "ClassyFire_subclass_clean.jpeg",
#     plot = subclass,
#     width = 10,
#     height = 6,
#     dpi = 300,
  # )

```

##FELLA on run of SIRIUS

```{r, fig.width=12, fig.height=7, out.width='100%', fig.retina=2}

set.seed(123)

#BiocManager::install("FELLA")
library(FELLA)
library(readr)
library(igraph)
library(dplyr)
library(data.table)
library(ggraph)
library(tidygraph)
library(shadowtext)
library(stringr)
library(ggtext)


#Run this first
#keggraph <- buildGraphFromKEGGREST("mmu")

## only needs to be run once; might take a while the first time... expect to wait maybe 15-25 min
#buildDataFromGraph(keggdata.graph = keggraph, databaseDir = "fella_mmu", matrices = c("hypergeom", "diffusion","pagerank")
#                     , internalDir = TRUE)

sirius_kegg <- read_tsv("dualion_structure_identifications_all.tsv") 
confident_kegg_foreground <- sirius_kegg %>% 
  filter(links %like% "KEGG" & SiriusScoreNormalized >= 1) %>% 
  pull(links) %>%
  sub(".*KEGG:\\(([^)]+)\\).*", "KEGG:\\1", .) %>%
  gsub("KEGG:", "\\1", .)
kegg_background <- sirius_kegg %>%
  filter(links %like% "KEGG" #& SiriusScoreNormalized > 0.9
         ) %>%
  pull(links) %>%
  sub(".*KEGG:\\(([^)]+)\\).*", "KEGG:\\1", .)%>%
  gsub("KEGG:", "\\1", .)
fella.db <- loadKEGGdata(databaseDir = "fella_mmu", loadMatrix = c("hypergeom", "diffusion", "pagerank"),
                          internalDir = TRUE)
method <- "diffusion"
## need to use diffusion if you're running simulation/monte carlo approx!!!; mc did not work for this data/gave no results
#method <- "diffusion"
nlimit <- 250

fella.data <- defineCompounds(#compounds = confident_kegg_foreground,
                              compounds = unlist(strsplit(confident_kegg_foreground, "\\s+") %>% na.omit()),
                              #compoundsBackground = kegg_background,
                              compoundsBackground = unlist(strsplit(kegg_background, "\\s+")%>% na.omit()),
                              data = fella.db)
fella.analysis_mutual <- enrich(#compounds = confident_kegg_foreground, 
                                compounds = unlist(strsplit(confident_kegg_foreground, "\\s+")%>% na.omit()),
                                #compoundsBackground = kegg_background,
                                compoundsBackground = unlist(strsplit(kegg_background, "\\s+")%>% na.omit()),
                                method = method,
                                approx = "gamma",
                                data = fella.db,
                                )

results_sirius <- generateResultsTable(object = fella.analysis_mutual, data = fella.db, 
                                   method = method,
                                   nlimit = nlimit)
results_sirius$adj_pScore <- p.adjust(results_sirius$p.score, method = "BH")

enriched_subgraph <- generateResultsGraph(object = fella.analysis_mutual, data = fella.db, threshold = 0.05,
                                          method = method,
                                          nlimit = nlimit)

plotGraph(enriched_subgraph)

# Extract node data with significance
node_data <- results_sirius %>%
  filter(KEGG.id %in% V(enriched_subgraph)$name) %>%
  transmute(
   name = KEGG.id,
    p.score,
   log10_p_score = -log10(p.score),
    label_text = KEGG.name
 )

#pulling individual ID from node data
kegg_ids <- node_data$name
results <- list()
#setting parameters
batch_size <- 10
num_batches <- ceiling(length(kegg_ids) / batch_size)
#loop function for all IDs
for (i in 1:num_batches) {
  start_index <- ((i - 1) * batch_size) + 1
  end_index <- min(i * batch_size, length(kegg_ids))
  batch_ids <- kegg_ids[start_index:end_index]
  result <- keggGet(batch_ids)
  results <- c(results, result)
}
#extracting full IDs 
entries_df <- data.frame(
  ENTRY = sapply(results, function(x) x[["ENTRY"]]),
  NAME  = sapply(results, function(x) paste(x[["NAME"]], collapse = "; ")),
  stringsAsFactors = FALSE
)


View(entries_df)
list(entries_df$ENTRY[23:65])

entries_df$ENTRY[23:65] <- gsub("^EC\\s+", "", entries_df$ENTRY[23:65])

list(entries_df$ENTRY[23:65])

colnames(entries_df)[colnames(entries_df) == "ENTRY"] <- "name"
colnames(entries_df)[colnames(entries_df) == "NAME"] <- "FULL_NAME"

node_data_1 <- merge(node_data, entries_df, by = "name")
node_data_1 <- node_data_1[, c("name", "FULL_NAME", "p.score", "log10_p_score")]

colnames(node_data_1)[colnames(node_data_1) == "FULL_NAME"] <- "label_text"

##################################

##individually edit longer names to appear on 2 lines to prevent overlap

node_data_1 <- node_data_1 %>%
  mutate(label_text = if_else(
    label_text == "",
    "",
    label_text
  ))

g_tidy <- as_tbl_graph(enriched_subgraph) %>%
  left_join(node_data_1, by = "name") %>%
  mutate(
    #remove everything that comes after the first occurrence of 2 ;;
    label_text = str_remove(label_text, "\\s*;;.*$"),
    #remove everything that comes after the occurrence of a dash
    label_text = str_remove(label_text, "\\s*- .*$"),
    label_text = ifelse(is.na(label_text), name, label_text),
    degree = centrality_pagerank(directed = FALSE),
    highlight = degree >= quantile(degree, 0.7, na.rm = TRUE),
    label_to_show = ifelse(highlight, label_text, NA)
  ) %>%
  activate(edges) %>%
  mutate(edge_highlight = .N()$highlight[from] & .N()$highlight[to]
         )
g <- ggraph(g_tidy, layout = "stress") +
  geom_edge_link(color = "gray90", edge_width = 1.5) +
  geom_node_point(color = "gray90", size = 4) +
  geom_edge_link(aes(filter = edge_highlight), color = "dodgerblue2", edge_width = 1.5) +
  geom_node_point(aes(filter = highlight, size = log10_p_score), color = "dodgerblue3", size = 4,
                  ) +
  geom_node_text(aes(filter = highlight, label = label_to_show),
                 color = "black", fontface = "bold",
                 repel = TRUE, max.overlaps = 50, size = 6.5,
                 box.padding = 1.5, point.padding = 1, segment.size = 0.4) +
  scale_size_continuous(
    name = expression(-log[10](italic(p)*"-score")),
    range = c(4, 14)) +
  theme_void(base_size = 20) +  # boost all text elements
  labs(title = "", subtitle = "", size = "-log10(p-score)") +
  theme(legend.position = "none",
        plot.title = element_text(size = 26, face = "bold", hjust = 0.5),
        plot.subtitle = element_text(size = 20),
        panel.background = element_rect(fill = "white", color = NA),
        plot.background = element_rect(fill = "white", color = NA)
  )

#ggsave("enrichedpathwaynetwork_serum_clean.jpeg", plot = g, width = 100, height = 60, units = "cm", dpi = 600, bg = "white")

print(g)


#hist(results_mutual%>% filter(Entry.type== "pathway") %>% pull(p.score), breaks = 100, main = "Diffusion P-Score Distribution")

# lets plot those pathways
TEP <- results_sirius %>%
  filter(p.score < 0.05 & Entry.type == "pathway") %>%
  slice_min(order_by = p.score, n = 20) %>%
  mutate(KEGG.name = str_remove(KEGG.name, "\\s-s")) %>%
  ggplot(aes(x = reorder(KEGG.name, -p.score), y = -log10(p.score))) +
  geom_col(fill = "blue") +
  coord_flip() +
  labs(
    title = "",
    x = "",
    y = expression(bold(-log[10]~"(P-score)"))
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold", size = 30),
    axis.title.x = element_text(face = "bold", size = 20),
    axis.title.y = element_text(face = "bold", size = 20),
    axis.text = element_text(face = "bold", size = 18, colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
    )
print(TEP)
# ggsave(
#     filename = "Top_enriched_metabolic_pathways_all.jpeg",
#     plot = TEP,
#     width = 8,
#     height = 6,
#     dpi = 300,)





```


```{r, preparing objects for viz, echo = FALSE}

# extract individual groups needed for both case studies
sqm_id =
  object %>%
  activate_mass_dataset(what = "sample_info") %>%
  filter(group == "sqm") %>%
  pull(sample_id)

hum_id = 
  object %>%
  activate_mass_dataset(what = "sample_info") %>%
  filter(group == "hum") %>%
  pull(sample_id)

mac_id = 
  object %>%
  activate_mass_dataset(what = "sample_info") %>%
  filter(group == "mac") %>%
  pull(sample_id)


```

## Case Study 1: squirrel monkeys vs macaques

After preprocessing (see tidyMassObjectGeneration.Rmd) by the Case Study #1 design, we are left with 6 control macaques and 18 total individuals in the combined human-squirrel monkey group. Perhaps the uneven sampling will bias our observations here. To check, we'll perform other case studies in the analysis later which compare single groups against one another.

```{r, preparing the sqmVmac object, echo = FALSE}

################ Creating the object
### CASE 2: sqm vs. mac outgroup
obj_sqmac <- 
  mutate_fc(object = object,
            control_sample_id = sqm_id, 
            case_sample_id = mac_id,
            mean_median = "median")
## 4 outgroup mac samples.
## 18 humans and macs case samples.
groups_to_keep <- c("sqm", "mac")
obj_sqmac <- obj_sqmac %>%
  activate_mass_dataset("sample_info") %>%
  filter(group %in% groups_to_keep)


```

The wilcoxon rank test was used with benjamini hochberg p-value correction to generate the heatmap. This heatmap for Case Study 1 shows 19 differentially expressed metabolites which have benjamini-hochberg, p-adj values of less than 0.05 This heatmap of the 19 differentially expressed metabolites according to Case Study 1 design will be used for the remainder of the Case Study 1 analysis. Interesingly, there is much more variable expression of spermine, spermidine, and biochanin A across the human-sqm group compared to the macaques which have comparatively, very uniform lack of these polyamine metabolites.

###Heatmap
```{r}
#set.seed(100)
### CASE 2: humsqm vs. mac outgroup
### Calculate p-values
object_sqmac <- 
  mutate_p_value(
    object = obj_sqmac,
    control_sample_id = sqm_id,
    case_sample_id = mac_id,
    method = "wilcox.test",
    p_adjust_methods = "BH"
    )
diff_metabolites.sm.c <-
  object_sqmac %>% 
  activate_mass_dataset(what = "variable_info") %>% 
  extract_variable_info() %>%
  filter(!is.na(Compound.name) & p_value_adjust < 0.1) #%>%
  #filter(na_freq < 0.2 | na_freq_2 < 0.2) %>% 
  #filter(Total.score > 0.9 & fc < 0.5 | fc > 1)
#write_tsv(diff_metabolites.sm.c,"/Users/emmachristopher/Desktop/Northwestern/Anthony_metabolomics/Cecal_Metabolomics/diff_metabolites_sm_c.tsv")

#BiocManager::install('biodbChebi')
library(biodbChebi)
mybiodb <- biodb::newInst()
chebi <- mybiodb$getFactory()$createConn('chebi')
entry <- chebi$getEntry('CHEBI:16113')
chebi$getSearchCategories()

# We download and read in chebi.obo ontology file
#https://obofoundry.org/ontology/chebi.html
library(webchem)
library(httr)
library(jsonlite)
library(purrr)
library(furrr)
getChebi_Ont <- function(chebi_ids) {
  results <- list()
  for (ID in seq_along(chebi_ids)) {
    response <- GET(paste0("https://www.ebi.ac.uk/chebi/beta/api/public/compound/", 
                           URLencode(chebi_ids[[ID]],
                            reserved = TRUE)
                  ),
                  accept("application/json"))
    if (status_code(response) == 200) {
      results[[chebi_ids[[ID]]]] <- httr::content(response, as = "parsed", encoding = "UTF-8")
    } else {
      warning(sprintf("Failed to fetch %s (status %d)", chebi_ids[[ID]], status_code(response)))
      results[[chebi_ids[[ID]]]] <- NULL
    }
  }
  return(results)
}
chebi_results.sm <- getChebi_Ont(diff_metabolites.sm.c$Lab.ID)
# recursive find for any of the terms at any level nested in list of listed ont output
chebi_bio.sm <- keep(chebi_results.sm, ~
    !any(grepl(
    paste(c("dye", "synthetic", "artificial", "non-natural",
            "non-endogenous", "contaminant", "preservative", "disinfectant", "industrial", "agriculture", "insecticide"  ,"triazole", "acetanilide", "hydrochloride", "benzamide", 
  "phenacetin", "prodrug", "API", "pharmaceutical", "inhibitor", 
  "agonist", "antagonist", "PEG", "pyrrolidine", "piperidine", 
  "aniline", "toxicant", "xenobiotic", "organophosphate"),
          collapse = "|"), .x, ignore.case = TRUE)) 
     
)

Natdiff_metabolites.sm.c <- diff_metabolites.sm.c %>%
  filter(Lab.ID %in% names(chebi_bio.sm) |
    str_detect(Lab.ID, "shen|HILIC|NO|HMDB|hmdb|fiehn|^C[0-9]+"))
readr::write_csv(Natdiff_metabolites.sm.c,
                file = "/Users/emmachristopher/Desktop/Northwestern/Anthony_metabolomics/Cecal_Metabolomics/Natdiff_sqmac_metabolites.csv")
Natdiff_metabolites.sm.c <- read.csv("/Users/emmachristopher/Desktop/Northwestern/Anthony_metabolomics/Cecal_Metabolomics/Natdiff_sqmac_metabolites.csv")

metabo_heatmap.sm <-
  Natdiff_metabolites.sm.c %>%
   filter(p_value < 0.05) %>%
   filter(p_value_adjust < 0.1) %>%
  filter(fc <= 0.05 | fc >= 5) %>%
  filter(na_freq < 0.2 | na_freq_2 < 0.2) %>% 
  filter(Total.score > 0.999 #| SS > 0.7
         )

heat_object <- 
object_sqmac %>% 
  activate_mass_dataset(what = "sample_info") %>% 
  filter(sample_id %in% c(sqm_id, mac_id)) %>% 
  activate_mass_dataset(what = "variable_info") %>% 
  filter(variable_id %in% metabo_heatmap.sm$variable_id) %>% 
  `+`(1) %>% 
  log(2) %>% 
  scale()
library(ComplexHeatmap)
# h1 <- HeatmapAnnotation(
#   `Donor Group` = extract_sample_info(temp_object)$group,
#   col = list(`Donor Group` = c("mac"="#a6761d","sqm"="#1b9e77","hum"="#666666")
#   )
# )
h1 <- HeatmapAnnotation(
  `Donor Group` = anno_simple(extract_sample_info(heat_object)$group, pch = c(str_extract(extract_sample_info(heat_object)$group,"^[a-z]")),
                              col = c("mac"="#a6761d","sqm"="#666666","hum"="#1b9e77"),
                              pt_gp = gpar(col = "white"))
  )


        


## had to take a screenshot of the tiff to get it into google slides
# PAPER 1

heat_object@variable_info = extract_variable_info(heat_object) %>%
  mutate(Compound.name = 
           ifelse(Compound.name == "Acetaminophen",
                  paste(variable_id, mz, rt, sep = "_"),
                  Compound.name)
         )
# h1 <- HeatmapAnnotation(
#   `Donor Group` = extract_sample_info(temp_object)$group,
#   col = list(`Donor Group` = c("mac"="#a6761d","sqm"="#1b9e77","hum"="#666666")
#   )
# )temp_object <- update_mass_dataset(temp_object)
# PAPER 1
#png("qual2metaboheatmap_cec.png", units="in", width=8.5, height=5, res=300)
massstat::Heatmap(matrix = heat_object, 
                  name = "z-score",
                  #color_space = "Dark2",
                  row_names_gp = gpar(cex = 0.4, fontsize = 15),
                  column_names_gp = gpar(cex = 0.2, fontsize = 30),
                  column_names_rot = 45,
                  top_annotation = h1,
                  row_labels = heat_object@variable_info$Compound.name,
                  border = TRUE,
                  #column_labels = rep("",24),
                  show_column_dend = FALSE,
                  show_row_dend = FALSE
                  #column_labels = temp_object@sample_info$subject_id
                  ) 

#dev.off()

```

###Volcano and PERMANOVA
```{r, volcano plot for macVsqm, echo = FALSE, message = FALSE}

#### VOLCANO 
temp_object <- heat_object
options(ggrepel.max.overlaps = Inf)
library(ggrepel)
temp_object@variable_info$Compound.name.1 <- temp_object@variable_info$Compound.name
#tiff("qual2metabvolc.tiff", units="in", width=8.5, height=8.5, res=300)
volcano_plot(object = temp_object,
             add_text = TRUE,
             text_from = "Compound.name.1",
             fc_up_cutoff = 6,
             fc_down_cutoff = 0.1,
             p_value_cutoff = 0.05,
             
             )
#dev.off(
print(diff_metabolites.sm.c[,c(6,7,8,9)])

### SQMAC PERMANOVA
dimnames <- c("mac01","mac03","mac04","mac06","mac07",
              "mac08","sqm09","sqm10","sqm11","sqm12",
              "sqm13","sqm14","sqm15","sqm16","sqm17",
              "sqm18")
 sample_distance <-
   as.matrix(dist_mass_dataset(x = temp_object, margin = "sample", method = "manhattan"))
#rownames(sample_distance) <- dimnames
#colnames(sample_distance) <- dimnames
result <- metadata %>% filter(!sample %in% c("mac2","mac5") & !group %in% "hum")
mat_cleaned <- sample_distance[
  !grepl("Ben-Oli-20230331-02|Ben-Oli-20230331-05|Ben-Oli-20230331-19|Ben-Oli-20230331-2[0-6]", rownames(sample_distance)),
  !grepl("Ben-Oli-20230331-02|Ben-Oli-20230331-05|Ben-Oli-20230331-19|Ben-Oli-20230331-2[0-6]", colnames(sample_distance)),
  drop = FALSE
]
print("squirrel monkey- vs. macaque-donor")
#adonis2(mat_cleaned ~ group, data = result)


```

###Web
```{r, echo = FALSE, message = FALSE}

library(ggraph)
library(tidygraph)
graph_data_sqmac <-
  convert_mass_dataset2graph(
    object = temp_object,
    margin = "variable",
    cor_method = "spearman",
    p_adjust_cutoff = 0.01,
    p_value_cutoff = 0.01,
    pos_cor_cutoff = 0.75,
    neg_cor_cutoff = -0.75
  ) %>% 
  mutate(Degree = centrality_degree(mode = 'all'))
#rm(list = "temp_object")
#gc()
#install.packages("extrafont")
#install.packages("shadowtext")
library(extrafont)
library(shadowtext)
library(ggraph)
loadfonts()
plot <- 
ggraph(graph = graph_data_sqmac, layout = "kk") +
  geom_edge_fan(aes(color = correlation,
                    width = -log(p_value, 10)),
                show.legend = TRUE) +
  geom_node_point(aes(size = Degree)) +
  shadowtext::geom_shadowtext(aes(x = x, y = y,
                                  label = Compound.name.1
                                  ),
                              bg.colour = "white",
                              colour = "black")+
  theme_graph() +
  scale_edge_color_gradient2(low = "darkblue", mid = "white", high = "red")
plot

```
###KEGG
```{r, kegg, echo = FALSE, message = FALSE}

### what's the difference bw pagerank and diffusion for projection you ask?
### you have to increase nlimit by like 100 and then you get more than one nodes
### the few you get include taurine and hypotaurine metabolism, with page rank you get 
### bile secretion and taurine metabolism-- processes that actually make sense together 
### incidentally we also pick up SCFAs in these projections... something we have more refined empirical evidence for in previous analysis targeted metabolomics data published last year
#### hypergeom is too sparse
nlimit <- 250
method <- "pagerank"

setwd("/Users/emmachristopher/Desktop/Northwestern/Anthony_metabolomics/Cecal_Metabolomics")
library(FELLA)
#keggraph <- buildGraphFromKEGGREST("mmu")
## only needs to be run once; might take a while the first time... expect to wait maybe 15-25 min
#buildDataFromGraph(keggdata.graph = keggraph, databaseDir = "fella_mmu", matrices = c("hypergeom", "diffusion", "pagerank"), internalDir = TRUE)
fella.db <- loadKEGGdata(databaseDir = "fella_mmu", loadMatrix = c("hypergeom", "diffusion", "pagerank"),
                         internalDir = TRUE)
fella.data <- defineCompounds(compounds = Natdiff_metabolites.sm.c$KEGG.ID,
                              compoundsBackground = diff_metabolites.sm.c$KEGG.ID,
                              data = fella.db)
fella.analysis_sm <- enrich(compounds = Natdiff_metabolites.sm.c$KEGG.ID,
                            compoundsBackground = diff_metabolites.sm.c$KEGG.ID,
                            method = FELLA::listMethods(),
                            #method = "diffusion",
                            data = fella.db)
results_sm <- generateResultsTable(object = fella.analysis_sm, data = fella.db, 
                                   method = method, nlimit = nlimit)
results_sm$adj_pScore <- p.adjust(results_sm$p.score, method = "BH")

enriched_subgraph <- generateResultsGraph(object = fella.analysis_sm, data = fella.db,threshold = 0.05,
                                          method = method, nlimit = nlimit)

plotGraph(enriched_subgraph)


```

## Case Study 2: humans vs macaques

```{r, preparing the sqmVmac object, echo = FALSE}

################ Creating the object
### CASE 2: sqm vs. mac outgroup
obj_humac <- 
  mutate_fc(object = object,
            control_sample_id = hum_id, 
            case_sample_id = mac_id,
            mean_median = "median")
## 4 outgroup mac samples.
## 18 humans and macs case samples.
groups_to_keep <- c("hum", "mac")
obj_humac <- obj_humac %>%
  activate_mass_dataset("sample_info") %>%
  filter(group %in% groups_to_keep)


```

###Heatmap
```{r, heatmap for sqmVmac, echo = FALSE}

#set.seed(100)
### CASE 2: hummac vs. mac outgroup
### Calculate p-values
object_humac <- 
  mutate_p_value(
    object = obj_humac,
    control_sample_id = hum_id,
    case_sample_id = mac_id,
    method = "wilcox.test",
    p_adjust_methods = "BH"
    )
diff_metabolites.hm.c <-
  object_humac %>% 
  activate_mass_dataset(what = "variable_info") %>% 
  extract_variable_info() %>%
  filter(!is.na(Compound.name) & p_value_adjust < 0.1) 

#diff_HMcsNames <- get_cid(diff_metabolites.hm.c$Compound.name ,from = "name", domain = "compound")

chebi_results.hm <- getChebi_Ont(diff_metabolites.hm.c$Lab.ID)
# recursive find for any of the terms at any level nested in list of listed ont output
chebi_bio.hm <- keep(chebi_results.hm, ~
    !any(grepl(
    paste(c("dye", "synthetic", "artificial", "non-natural",
            "non-endogenous", "contaminant", "preservative", "disinfectant", "industrial", "agriculture", "insecticide"  ,"triazole", "acetanilide", "hydrochloride", "benzamide", 
  "phenacetin", "prodrug", "API", "pharmaceutical", "inhibitor", 
  "agonist", "antagonist", "PEG", "pyrrolidine", "piperidine", 
  "aniline", "toxicant", "xenobiotic", "organophosphate"),
          collapse = "|"), .x, ignore.case = TRUE)) 
     
)

Natdiff_metabolites.hm.c <- diff_metabolites.hm.c %>%
  filter(Lab.ID %in% names(chebi_bio.hm) |
    str_detect(Lab.ID, "shen|HILIC|NO|HMDB|hmdb|fiehn|^C[0-9]+"))
readr::write_csv(Natdiff_metabolites.hm.c,
                file = "/Users/emmachristopher/Desktop/Northwestern/Anthony_metabolomics/Cecal_Metabolomics/diff_humac_metabolites.csv")
Natdiff_metabolites.hm.c <- read.csv("/Users/emmachristopher/Desktop/Northwestern/Anthony_metabolomics/Cecal_Metabolomics/diff_humac_metabolites.csv")

metabo_heatmap.hm <-
  Natdiff_metabolites.hm.c %>%
   filter(p_value < 0.05) %>%
   filter(p_value_adjust < 0.1) %>%
  filter(fc <= 0.05 | fc >= 3) %>%
  filter(na_freq < 0.2 | na_freq_2 < 0.2) %>% 
  filter(Total.score > 0.999 #| SS > 0.7
         )

temp_object <- 
object_humac %>% 
  activate_mass_dataset(what = "sample_info") %>% 
  filter(sample_id %in% c(hum_id, mac_id)) %>% 
  activate_mass_dataset(what = "variable_info") %>% 
  filter(variable_id %in% metabo_heatmap.hm$variable_id) %>% 
  `+`(1) %>% 
  log(2) %>% 
  scale()
library(ComplexHeatmap)

h1 <- HeatmapAnnotation(
  `Donor Group` = anno_simple(extract_sample_info(temp_object)$group, pch = c(str_extract(extract_sample_info(temp_object)$group,"^[a-z]")),
                              col = c("mac"="#a6761d","sqm"="#666666","hum"="#1b9e77"),
                              pt_gp = gpar(col = "white"))
  )

temp_object@variable_info = extract_variable_info(temp_object) %>%
  mutate(Compound.name = 
           ifelse(Compound.name == "Acetaminophen",
                  paste(variable_id, mz, rt, sep = "_"),
                  Compound.name)
         )

# PAPER 1
#png("qual2metaboheatmap_cec.png", units="in", width=8.5, height=5, res=300)
massstat::Heatmap(matrix = temp_object, 
                  name = "z-score",
                  #color_space = "Dark2",
                  row_names_gp = gpar(cex = 0.4, fontsize = 15),
                  column_names_gp = gpar(cex = 0.2, fontsize = 30),
                  column_names_rot = 45,
                  top_annotation = h1,
                  row_labels = temp_object@variable_info$Compound.name,
                  border = TRUE,
                  #column_labels = rep("",24),
                  show_column_dend = FALSE,
                  show_row_dend = FALSE
                  #column_labels = temp_object@sample_info$subject_id
                  ) 

#dev.off()

```

###Volcano and PERMANOVA
```{r, volcano plot for humVmac, echo = FALSE, message = FALSE}

options(ggrepel.max.overlaps = Inf)
library(ggrepel)
temp_object@variable_info$Compound.name.1 <- temp_object@variable_info$Compound.name
#tiff("qual2metabvolc.tiff", units="in", width=8.5, height=8.5, res=300)
volcano_plot(object = temp_object,
             add_text = TRUE,
             text_from = "Compound.name.1",
             p_value_cutoff = 0.05
             )
#dev.off()
#readr::write_csv(Natdiff_metabolites.hm.c, 
#                 file = "/Users/apulvino/cecal_metabolomics/stat_analysis/diff_humac_metabolites.csv")
print(Natdiff_metabolites.hm.c[,c(7,8,9,10)])



####.   HUMAC PERMANOVA
dimnames <- c("mac01","mac03","mac04","mac06","mac07",
              "mac08", "hum19","hum20","hum21","hum22",
              "hum23","hum24","hum25","hum26")
 sample_distance <-
   as.matrix(dist_mass_dataset(x = temp_object, margin = "sample", method = "manhattan"))
# rownames(sample_distance) <- dimnames
# colnames(sample_distance) <- dimnames
result <- metadata %>% filter(!sample %in% c("mac2","mac5") & !group %in% "sqm")
mat_cleaned <- sample_distance[
  !grepl("Ben-Oli-20230331-02|Ben-Oli-20230331-05|Ben-Oli-20230331-09|Ben-Oli-20230331-1[0-8]", rownames(sample_distance)),
  !grepl("Ben-Oli-20230331-02|Ben-Oli-20230331-05|Ben-Oli-20230331-09|Ben-Oli-20230331-1[0-8]", colnames(sample_distance)),
  drop = FALSE
]
print("human- vs. macaque-donor")
#adonis2(mat_cleaned ~ group, data = result)


```

###Web
```{r, echo = FALSE, message = FALSE}

library(ggraph)
library(tidygraph)
graph_data_humac <-
  convert_mass_dataset2graph(
    object = temp_object,
    margin = "variable",
    cor_method = "spearman",
    p_adjust_cutoff = 0.05,
    p_value_cutoff = 0.05,
    pos_cor_cutoff = 0.5,
    neg_cor_cutoff = -0.5
  ) %>% 
  mutate(Degree = centrality_degree(mode = 'all'))
#rm(list = "temp_object")
#gc()
#install.packages("extrafont")
#install.packages("shadowtext")
library(extrafont)
library(shadowtext)
library(ggraph)
loadfonts()
plot <- 
ggraph(graph = graph_data_humac, layout = "kk") +
  geom_edge_fan(aes(color = correlation,
                    width = -log(p_value, 10)),
                show.legend = TRUE) +
  geom_node_point(aes(size = Degree)) +
  shadowtext::geom_shadowtext(aes(x = x, y = y,
                                  label = Compound.name.1
                                  ),
                              bg.colour = "white",
                              colour = "black")+
  theme_graph() +
  scale_edge_color_gradient2(low = "darkblue", mid = "white", high = "red")
plot

```

###KEGG
```{r, kegg, echo = FALSE, message = FALSE}

setwd("/Users/emmachristopher/Desktop/Northwestern/Anthony_metabolomics/Cecal_Metabolomics")
library(FELLA)
# fella_hsa or fella_mmu can be used :) 
## mmu end up working much better for obvious reasons
#keggraph <- buildGraphFromKEGGREST("mmu")
## only needs to be run once; might take a while the first time... expect to wait maybe 15-25 min
#buildDataFromGraph(keggdata.graph = keggraph, databaseDir = "fella_mmu", matrices = c("hypergeom", "diffusion", "pagerank"), internalDir = TRUE)
# fella.db <- loadKEGGdata(databaseDir = "fella_mmu", loadMatrix = c("hypergeom", "diffusion", "pagerank"),
#                          internalDir = TRUE)

fella.data <- defineCompounds(compounds = Natdiff_metabolites.hm.c$KEGG.ID,
                              compoundsBackground = diff_metabolites.hm.c$KEGG.ID,
                              data = fella.db)
fella.analysis_hm <- enrich(compounds = Natdiff_metabolites.hm.c$KEGG.ID,
                            compoundsBackground = diff_metabolites.hm.c$KEGG.ID,
                            method = FELLA::listMethods(),
                            #method = "diffusion",
                            data = fella.db)
results_hm <- generateResultsTable(object = fella.analysis_hm, data = fella.db, 
                                   method = method, nlimit = nlimit)
results_hm$adj_pScore <- p.adjust(results_hm$p.score, method = "BH")

enriched_subgraph <- generateResultsGraph(object = fella.analysis_hm, data = fella.db,threshold = 0.05,
                                          method = method,nlimit = nlimit)

plotGraph(enriched_subgraph)


```

## tidymass/Mutual compound run of sirius

```{r}

## determining mutually differentially abundant pathways/compounds
####.     KEGG path enrichment
################   PAPER 1   KEGG path enrichment    ###########################
#keggraph <- buildGraphFromKEGGREST("mmu")
## only needs to be run once; might take a while the first time... expect to wait maybe 15-25 min
#buildDataFromGraph(keggdata.graph = keggraph, databaseDir = "fella_mmu", matrices = c("hypergeom", "diffusion", "pagerank"), internalDir = TRUE)

# fella.db <- loadKEGGdata(databaseDir = "fella_mmu", loadMatrix = c("hypergeom", "diffusion", "pagerank"),
#                          internalDir = TRUE)
method <- "pagerank"
#method <- "diffusion"
nlimit <- 250
fella.data <- defineCompounds(compounds = intersect(Natdiff_metabolites.sm.c$KEGG.ID,
                                              Natdiff_metabolites.hm.c$KEGG.ID),
                          compoundsBackground = c(diff_metabolites.hm.c$KEGG.ID,
                                                 diff_metabolites.sm.c$KEGG.ID),
                              data = fella.db)
fella.analysis_mutual <- enrich(compounds = intersect(Natdiff_metabolites.sm.c$KEGG.ID,
                                              Natdiff_metabolites.hm.c$KEGG.ID), 
                              compoundsBackground = c(diff_metabolites.hm.c$KEGG.ID,
                                                 diff_metabolites.sm.c$KEGG.ID),
                            #method = FELLA::listMethods(),
                            #method = "pagerank",
                            method = method,
                            approx = "gamma",
                            data = fella.db)

results_mutual <- generateResultsTable(object = fella.analysis_mutual, data = fella.db, 
                                   method = method,
                                   nlimit = nlimit)
results_mutual$adj_pScore <- p.adjust(results_mutual$p.score, method = "BH")


enriched_subgraph <- generateResultsGraph(object = fella.analysis_mutual, data = fella.db,
                                          threshold = 0.05,
                                          method = method,
                                          nlimit = nlimit)

plotGraph(enriched_subgraph)

# Extract node data with significance
node_data_tidymass <- results_mutual %>%
  filter(KEGG.id %in% V(enriched_subgraph)$name) %>%
  transmute(
    name = KEGG.id,
    p.score,
    log10_p_score = -log10(p.score),
    label_text = KEGG.name
  )


#get all the full names for IDs

#to pull single value from node_data

library(KEGGREST)
#View(keggGet("1.11.2.2"))

#looping for all 118 values
kegg_ids_tm <- node_data_tidymass$name
results_tm <- list()

batch_size <- 10
num_batches <- ceiling(length(kegg_ids_tm) / batch_size)

for (i in 1:num_batches) {
  start_index <- ((i - 1) * batch_size) + 1
  end_index <- min(i * batch_size, length(kegg_ids_tm))
  batch_ids <- kegg_ids_tm[start_index:end_index]
  result_tm <- keggGet(batch_ids)
  results_tm <- c(results_tm, result_tm)
}

View(results_tm)

#extract entry and name from results file

entries_df_tm <- data.frame(
  ENTRY = sapply(results_tm, function(x) x[["ENTRY"]]),
  NAME  = sapply(results_tm, function(x) paste(x[["NAME"]], collapse = "; ")),
  stringsAsFactors = FALSE
)

View(entries_df_tm)
list(entries_df_tm$ENTRY[36:51])
entries_df_tm$ENTRY[entries_df_tm$ENTRY == "EC 1.11.2.2"] <- "1.11.2.2"
entries_df_tm$ENTRY[entries_df_tm$ENTRY == "EC 1.13.11.53"] <- "1.13.11.53"
entries_df_tm$ENTRY[entries_df_tm$ENTRY == "EC 1.14.14.139" ] <- "1.14.14.139"
entries_df_tm$ENTRY[entries_df_tm$ENTRY == "EC 1.8.1.9"] <- "1.8.1.9"
entries_df_tm$ENTRY[entries_df_tm$ENTRY == "EC 2.4.1.17"] <- "2.4.1.17"
entries_df_tm$ENTRY[entries_df_tm$ENTRY == "EC 2.7.1.159"] <-  "2.7.1.159"  
entries_df_tm$ENTRY[entries_df_tm$ENTRY == "EC 2.8.2.14"] <- "2.8.2.14"
entries_df_tm$ENTRY[entries_df_tm$ENTRY == "EC 2.8.2.2"] <- "2.8.2.2"
entries_df_tm$ENTRY[entries_df_tm$ENTRY == "EC 3.1.2.27"] <- "3.1.2.27"
entries_df_tm$ENTRY[entries_df_tm$ENTRY == "EC 3.1.3.56"] <- "3.1.3.56"
entries_df_tm$ENTRY[entries_df_tm$ENTRY == "EC 3.1.3.57" ] <- "3.1.3.57"
entries_df_tm$ENTRY[entries_df_tm$ENTRY == "EC 3.1.3.66" ] <- "3.1.3.66"  
entries_df_tm$ENTRY[entries_df_tm$ENTRY == "EC 3.1.6.2" ] <- "3.1.6.2" 
entries_df_tm$ENTRY[entries_df_tm$ENTRY == "EC 4.4.1.1" ] <- "4.4.1.1" 
entries_df_tm$ENTRY[entries_df_tm$ENTRY == "EC 4.4.1.13" ] <- "4.4.1.13" 
entries_df_tm$ENTRY[entries_df_tm$ENTRY == "EC 6.2.1.7" ] <- "6.2.1.7" 
list(entries_df_tm$ENTRY[36:51])
  
colnames(entries_df_tm)[colnames(entries_df_tm) == "ENTRY"] <- "ID"
colnames(entries_df_tm)[colnames(entries_df_tm) == "NAME"] <- "FULL_NAME"
colnames(node_data_tidymass)[colnames(node_data_tidymass) == "name"] <- "ID"

node_data_2 <- merge(node_data_tidymass, entries_df_tm, by = "ID")
node_data_2 <- node_data_2[, c("ID", "FULL_NAME", "p.score", "log10_p_score")]

colnames(node_data_2)[colnames(node_data_2) == "ID"] <- "name"
colnames(node_data_2)[colnames(node_data_2) == "FULL_NAME"] <- "label_text"
##################################

# g_tidy <- g_tidy %>%
#   mutate(label_to_show = case_when(
#     label_to_show == "1,2-dihydroxy-4-methylcyclohexa-3,5-diene-1-carboxylate:NAD+ oxidoreductase (decarboxylating)" ~ "1,2-dihydroxy-4-methylcyclohexa-3,\n5-diene-1-carboxylate:NAD+ oxidoreductase (decarboxylating)",
#     label_to_show == "	
# 1,6-dihydroxy-5-methylcyclohexa-2,4-dienecarboxylate:NAD+ oxidoreductase (decarboxylating)" ~ "	
# 1,6-dihydroxy-5-methylcyclohexa-2,4-dienecarboxylate:\nNAD+ oxidoreductase (decarboxylating)",
# label_to_show == "(6E)-8-methylnon-6-enoyl-CoA:vanillylamine 8-methylnon-6-enoyltranferase" ~ "	
# (6E)-8-methylnon-6-enoyl-CoA:vanillylamine\n8-methylnon-6-enoyltranferase",
#     TRUE ~ label_to_show


# g_tidy <- as_tbl_graph(enriched_subgraph) %>%
#   left_join(node_data_2, by = "name") %>%
#   mutate(
#     label_text = ifelse(is.na(label_text), name, label_text),
#     degree = centrality_pagerank(),
#     highlight = degree >= quantile(degree, 0.8, na.rm = TRUE),
#     label_to_show = ifelse(highlight, label_text, NA)
#   ) %>%
#   activate(edges) %>%
#   mutate(edge_highlight = .N()$highlight[from] & .N()$highlight[to]
#          )


g_tidy <- as_tbl_graph(enriched_subgraph) %>%
  left_join(node_data_2, by = "name") %>%
  mutate(
    label_text = ifelse(is.na(label_text), name, label_text),
    #remove everything that comes after the first occurrence of 2 ;;
    label_text = str_remove(label_text, "\\s*;;.*$"),
    #remove everything that comes after the occurrence of a dash
    label_text = str_remove(label_text, "\\s*- .*$"),
    degree = centrality_pagerank(),
    highlight = degree >= quantile(degree, 0.8, na.rm = TRUE),
    label_to_show = ifelse(highlight, label_text, NA)
  ) %>%
  activate(edges) %>%
  mutate(edge_highlight = .N()$highlight[from] & .N()$highlight[to])


# t <- ggraph(g_tidy, layout = "stress") +
#   geom_edge_link(color = "gray90", edge_width = 0.5) +
#   geom_node_point(color = "gray90", size = 5) +
#   geom_edge_link(aes(filter = edge_highlight), color = "#e6ab02", edge_width = 2) +
#   geom_node_point(aes(filter = highlight, size = log10_p_score), color = "#e6ab02") +
#   geom_node_text(
#     aes(filter = highlight, label = label_to_show),
#     color = "black",
#     fontface = "bold",
#     repel = TRUE,
#     max.overlaps = 50,         # show all possible labels
#     size = 3
#   ,                   # bump up text size
#     box.padding = 1.5,
#     point.padding = 1,
#     segment.size = 0.4
#   ) +
#   scale_size_continuous(
#     name = expression(-log[10](italic(p)*"-score")),
#     range = c(4, 14)            # scale node size larger
#   ) +
#   theme_void(base_size = 20) +  # boost all text elements
#   labs(
#     title = "TidyMass-Annotated Enriched Pathway Network",
#     subtitle = "",
#     size = "-log10(p-score)"
#   ) +
#   theme(
#     legend.position = "none",
#     plot.title = element_text(size = 26, face = "bold", hjust = 0.5),
#     plot.subtitle = element_text(size = 20)
#   )

t <- ggraph(g_tidy, layout = "stress") +
  # Base edges and nodes
  geom_edge_link(color = "gray90", edge_width = 0.5) +
  geom_node_point(color = "gray90", size = 5) +
  
  # Highlighted edges and nodes
  geom_edge_link(
    aes(filter = edge_highlight),
    color = "#e6ab02",
    edge_width = 2
  ) +
  geom_node_point(
    aes(filter = highlight, size = log10_p_score),
    color = "#e6ab02"
  ) +
  
  # Node labels
  geom_node_text(
    aes(filter = highlight, label = label_to_show),
    color = "black",
    fontface = "bold",
    repel = TRUE,
    max.overlaps = 50,               # show all possible labels
    size = 7,                        # bump up text size
    box.padding = 1.5,               # padding around text
    point.padding = 1,              # distance from nodes
    segment.size = 0.4
  ) +
  
  # Size scale
  scale_size_continuous(
    name = expression(-log[10](italic(p)*"-score")),
    range = c(4, 14)                 # scale node size larger
  ) +
  
  # Theme and labels
  theme_void(base_size = 20) +       # boost all text elements
  labs(
    title = "",
    subtitle = "",
    size = "-log10(p-score)"
  ) +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 30, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 20)
  )


print(t)
ggsave("/Users/emmachristopher/Desktop/Northwestern/Anthony_metabolomics/Cecal_Metabolomics/tidyanno_network_wide_NT.jpeg", plot = t,
       width = 90, height = 50, dpi = 600, units = "cm", bg = "white")

# hist(results_mutual%>% filter(Entry.type== "pathway") %>% pull(p.score), breaks = 100, main = "Diffusion P-Score Distribution")


# lets plot those pathways
TEP_2 <- results_mutual %>%
  filter(p.score < 0.05 & Entry.type == "pathway") %>%
  mutate(KEGG.name = str_remove(KEGG.name, "\\s*-.*$")) %>%
  slice_min(order_by = p.score, n = 20) %>%
  ggplot(aes(x = reorder(KEGG.name, -p.score), y = -log10(p.score))) +
  geom_col(fill = "#E6AB02FF") +
  coord_flip() +
  labs(
    title = "",
    x = "",
    y = expression(bold(-log[10]~"(P-score)"))
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(face = "bold", size = 30),
    axis.title.x = element_text(face = "bold", size = 15),
    axis.title.y = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold", size = 18, colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
    )
print(TEP_2)
ggsave(
    filename = "Top_enriched_metabolic_pathways_tm_NT.jpeg",
    plot = TEP_2,
    width = 8,
    height = 6,
    dpi = 300)
```
